### 2012-12-31 paper/2012/SWJ-PanLex/paper.tex
Added the paper here for now

\documentclass[sw]{iosart2c}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{times}
\usepackage{natbib}
\usepackage{amsmath}
\usepackage{dcolumn}
\usepackage{graphicx}
\usepackage{url}
\usepackage{xspace}
\usepackage{color}
\usepackage{eurosym}
\usepackage{todonotes}
\usepackage[breaklinks]{hyperref}
\def\sectionautorefname{Section}
\def\subsectionautorefname{Section}
\usepackage{multirow}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{array}
\usepackage{textcomp}
\newcolumntype{R}{>{\raggedleft\arraybackslash}X}
\newcolumntype{L}{>{\raggedright\arraybackslash}X}
\newcolumntype{C}{>{\centering\arraybackslash}X}

\newcommand{\TODO}[1]{{\color{red}{\textbf{TODO: {#1}}\xspace}}}
\newcommand{\CONSIDER}[1]{{\color{blue}{\textbf{CONSIDER: {#1}}\xspace}}}
\newcolumntype{d}[1]{D{.}{.}{#1}}

\newcommand{\claus}[1]{\todo[inline]{[Claus: #1]}}
%\newcommand{\textrdf}[1]{\texttt{#1}}
\newcommand{\textrdf}[1]{\emph{#1}}

\usepackage{listings}
%\usepackage{verbatim}
\lstset{%
    numberbychapter=false,
    numbers=left,
    numberstyle=\tiny,
    basicstyle=\footnotesize\ttfamily,
    tabsize=2,
    framexleftmargin=2pt,
    captionpos=b,
    frame=single,
    breaklines=true
}
\lstdefinestyle{rdf}{numberblanklines=true, morekeywords={}}
\lstdefinestyle{sparql}{basicstyle=\ttfamily\scriptsize,numberblanklines=true,
    morekeywords={SELECT,OPTIONAL,FROM,DISTINCT,a,WHERE,FILTER,GROUP,ORDER,LIMIT,BY,IN,AS},
    emph={r,pub,aairObject,verb,person,bday,s,p,o},emphstyle=\textit
}
\lstdefinestyle{turtle}{basicstyle=\ttfamily\tiny ,numberblanklines=true,
    morekeywords={a, @prefix},
    morecomment=[s][\textrm]{<}{>},
    morecomment=[s][\textit]{"}{"},
}

% Default \todo{} to inline mode
\newcommand*{\origtodo}{}
\let\origtodo\todo
\renewcommand*{\todo}{\origtodo[inline]}

\firstpage{1} \lastpage{5} \volume{1} \pubyear{2012}
\begin{document}
\begin{frontmatter}

%\pretitle{Pretitle}

\title{PanLex}
\runningtitle{PanLex}
\todo{find a real title}

\review{Name Surname, University, Country}{Name Surname, University, Country}{Name Surname, University, Country}

\author[A]{\fnms{Patrick} \snm{Westphal}},
\author[A]{\fnms{Claus} \snm{Stadler}},
%\runningauthor{M. Martin et al.}
\address[A]{University of Leipzig, Institute of Computer Science, AKSW, Postfach 100920, D-04009 Leipzig, Germany\\
E-mail: \{lastname\}@informatik.uni-leipzig.de}
\todo{the e-mail address pattern won't work for both of us}

\begin{abstract}
At present, there are approximately 6,900 actively spoken languages in the world.
However, some experts claim, that the process of globalization may eventually lead
to a world loosing this linguistic diversity.
The vision of the PanLex project is to save these languages, especially low density ones,
by allowing them to be a part of the Information Age by making them applicable via translations.
For this reason, PanLex gathers and integrates information from several
linguistict sources, such as monolingual dictionaries, bilingual dictionaries, multilingual dictionaries,
glossaries, standards and thesauri.
In this paper we describe how we transformed this data to RDF, interlinked it with
lexvo and DBpedia and published it as Linked Data and SPARQL.
\end{abstract}

\begin{keyword}
Multi-lingual Linked Open Data, PanLex, Lexcial Resource, RDF, RDB2RDF, Sparqlify
\end{keyword}
\end{frontmatter}

\section{Introduction}
% \begin{itemize}
% \item introduce PanLex project
% \item @author: Jonathan
% \item Motivation: About approximately 6900 active languages (\todo{source seems to be http://www.amazon.com/Ethnologue-Languages-M-Paul-Lewis/dp/1556712162}), Existing translation only between relatively few language pairs (dictionaries).
% \item Transitive word-by-word translations 
% * high density and low density languages
% * decline in linguistic diversity
% * These combined forces have led to predictions that something between half and 90%
%   of the world’s living languages will die within the next century (Woodbury 2006; UNESCO 2003: 2).
%   Weaker forms of these forces appear to be shrinking the use of medium-density languages in science,
%   diplomacy, business, and other domains (Phillipson 2008).
% * When languages die, they argue, the world loses
%   (1) irreplaceable knowledge of history, medicine, nature, and productive methods encoded in languages’
% * extraction + integration
% \end{itemize}
At present, there are about 7,000 languages in the world with approximately 6,900 of them being actively spoken\footnote{see \url{http://www.sil.org/iso639-3/download.asp} for a list of registered languages}.
Nonetheless, some experts claim, that the process of globalization may lead to a monolingual world loosing this linguistic diversity~\cite{endang_lang}.
Since there are a lot of low density languages that were not taken over from generation to generation anymore, we could irreversibly loose parts of human knowledge.
Some predictions state that this decline may lead to a loss of half to 90\% of human's languages~\cite{endang_lang}.

The vision of the PanLex project is to save all these thousands of languages, especially those low-density ones that are threatened by extinction.
PanLex tries to provide means for panlingual translations from any language of the world into any other.
But it is not feasible to just store all the translations of one word into any other language since this would result in trillions of translations.
So the PanLex project additionally uses artificial intelligence techniques to infer translations.
As a consequence, not just mainstream languages like English, but even these low-density languages may be used for a global communication.
% Since it is not feasible to just store all the translations of one word into any other language, which would result in trillions of translations, a system is needed that allows translations by inference using artificial intelligence.

Currently there is a growing community trying to combine linguistic knowledge with Semantic Web technologies and so build a web of linguistic data, also known as the \emph{Linguistic Linked Open Data (LLOD) cloud}.
Originally the PanLex project provides different web services\footnote{\url{http://panlex.org/try/}} and APIs\footnote{\url{http://panlex.org/tech/doc/api/}} to access its data.
Since the project shares the idea of open access we tried to make PanLex a part of the LLOD cloud.

In \autoref{sec:triplificate} we will introduce the PanLex dataset, present our PanLex RDF vocabulary and explain how we transformed the one into the other.
The following section is about the actual dataset publishing and how we linked to other datasets of the LLOD cloud.
Finally we discuss related work, conclude our approach and give some hints to future work.

\section{Triplification of the Raw Data}
\label{sec:triplificate}
In this section we first provide an analysis of the original PanLex data.
Subsequently, we will explain the steps taken to publish the data as RDF.
The vocabulary introduced for the PanLex dataset will be described and afterwards we will show how the original data was converted to RDF and published via a SPARQL endpoint.
The interlinking of the created RDF data with other datasets will be discussed in the following section.

\subsection{Analysis of the Original Dataset}
\label{sec:analysis}
At the core, PanLex is a \emph{database}, served by a PostgreSQL database server, that ``represents \emph{assertions} about the \emph{meanings} of \emph{expressions}''\footnote{\url{http://panlex.org/tech/doc/design/panlex-db-design.pdf}}.
The database contains about 20 million meanings and 19 million expressions extracted from about 4.000 sources.

Assertions are factual claims attributed to a certain source and editor.
They do not represent absolute truths and as such it is valid that different editors make contradictory assertions on different sources.

Expressions are lexemes (dictionary entries) and are related to their respective languages.
Furthermore, expressions have meanings. All expressions sharing the same meaning are either translations or synonyms.

The lexical database of the PanLex project is based on different sources like monolingual dictionaries,
bilingual dictionaries, multilingual dictionaries, glossaries, standards, thesauri and so on.
The list of used sources can also be found on the web\footnote{\url{http://panlex.org/tech/plrefs.shtml}}.

The actual data derived from these sources comprises single- and multi-word expressions and meanings assigned to them.
The most important entities and their relations are depicted in \autoref{fig:db-schema} and are explained in more the detail in the following.

%This denotation process is also reflected in the database design shown below.
%To carry out the goal of a multi-lingual information source a suitable database design is needed.
\begin{figure}
\centering
\includegraphics[width=0.4\textwidth]{images/schema.png}
\caption{The PanLex database schema}
\label{fig:db-schema}
\end{figure}

\begin{itemize}
  \item \emph{Expressions} are lemmas. For example, \emph{``go''} is a valid expression, whereas \emph{``went''} is not.
  Expressions are always given in a language variety and can only be given once per language variety.
  They were extracted from the above mentioned mono- or bilingual dictionaries or from other sources.
  \item \emph{Languages} in PanLex are identified using ISO 639-3\footnote{\url{http://www.sil.org/iso639-3/codes.asp}} individual and macrolanguage codes, ISO 639-2\footnote{\url{http://www.loc.gov/standards/iso639-2/}} collective codes and ISO 639-5\footnote{\url{http://www.loc.gov/standards/iso639-5/}} codes.
  \item \emph {Language Varieties} within a language are further distinguished by custom names.
  For example, ``eng'' is the ISO 639-3 code for the English language.
  Panlex defines ``simple english'' and ``british english'' as varieties of it.
  These can be given in the native writing system.
  So, in contrast to mechanisms like IETF BCP 47\footnote{\url{http://tools.ietf.org/html/bcp47}} there is no need for a transcription. %  and one can use variety names like ``{زبان اوستایی}''.
  Additionally these variety names do not have to be registered like the IETF BCP 47 language subtags\footnote{\url{http://www.iana.org/assignments/language-subtag-registry}}.
  \item \emph{Meanings} of expressions are stated by users based on their
  interpretation of an information source.
  Information sources can be dictionaries, thesauri but also personal knowledge. 
  The pair comprised of the user and the information source is referred to as a meaning's \emph{approver}.
  Users may add informal definitions to meanings in any language variety.
  \item \emph{Definitions} are optional descriptions of a meaning.
  They are given in a certain language variety.
  A description of the verb \emph{browse} for example could be \emph{``move or surf through various files on a computer, the Internet, etc.''}.
  \item \emph{Denotations} relate expressions to possible \emph{meanings}.
  So homonyms are represented by one expression being connected to different meanings.
  Denotations may be enriched with parts of speech tags adopted from the
  Open Lexicon Interchange Format (OLIF) standard\footnote{\url{http://www.olif.net/}}.
  For example, \emph{fall} can be a verb or a noun for autumn.
  Furthermore, it is possible to generically tag denotations with meta data
  using key-value pairs.
  For instance, an English expression \emph{pig} when synonymous with
  \emph{police officer} could be annotated with \emph{pragmatics=vulgar}.
  \item \emph{Users} are related to their defined language varieties and edited approvers.
  \item \emph{Licenses} are also considered by the PanLex project.
  At present there are ten different license categories an approver can get into.
  They are \emph{public domain}, \emph{Creative Commons}, \emph{request} (meaning that one has to ask the author of the source at hand), \emph{GNU General Public License}, \emph{GNU Lesser General Public License}, \emph{GNU Free Documentation License}, \emph{MIT License}, \emph{copyright} (stating that there is a certain copyright holder), \emph{other} and \emph{unknown}.
\end{itemize}

An overview of the number of instances per entity in the PanLex database is given in~\autoref{fig:plx-entity-counts}.
\begin{table}
\centering
\begin{scriptsize}
\begin{tabular}{lr}
\toprule
Entity & Instances \\
%\midrule
Expressions & 18,580,594 \\ % SELECT COUNT(ex) FROM ex;
%\midrule
Languages   & 7,839 \\ % SELECT COUNT(lc) FROM lc; -- all language codes (ISO 639-3, ISO 639-2, ISO 639-5) known in the PanLex db
% FIXME: there should not be more languages than language varieties
%\midrule
Language Varieties & 7,248 \\ % SELECT COUNT(lv) FROM lv;
%\midrule
Meanings    & 20,023,427 \\ % SELECT COUNT(mn) FROM mn;
%\midrule
Definitions & 2,522,605 \\ % SELECT COUNT(df) FROM df;
%\midrule
Denotations & 50,803,243 \\ % SELECT COUNT(dn) FROM dn;
%\midrule
Users       & 7 \\ % SELECT COUNT(us) FROM us;
%\midrule
Licenses    & 10 \\ % SELECT COUNT(li) FROM apli;
\bottomrule
\end{tabular}
\end{scriptsize}
\caption{Number of entities in the PanLex database}
\label{fig:plx-entity-counts}
\end{table}

%\todo{Possibly also: approver editors, language variety editors, exemplar/approved characters, approver varieties}

%The same expression may be mentioned in multiple information sources.
%However, an expression's possible meanings depend on the user's interpretion of them in context of such a source.
%For this reason, every meaning depends on an approver.
%\emph{Denotations} capture the mappings between expressions and their possible meanings.

There are two noteworthy limitations of the database schema:
Approver combines a user with an information source, however the \emph{information source} is not modelled as a distinct entity.
Users are only associated with language variety and approver entities. However,
it is not possible to track which denotations were added by users.
\todo{What are the consequences? }

%These sources are fundamental for the PanLex database design and are aggregated in the approver table.
%According to these approvers there are special meanings that are represented by expressions.
%%Meanings are always related to an approver, which captures how a user interprets an expression information source.

%So, if there are two expressions, they can either be synonyms (if they are given in the same language) or translations of each other.
%\\
%Expressions are lexemes that can be seen as entries of a lexicon or dictionary of a given language.
%They are represented in their lemma or dictionary form.

%These expressions can be given in different language varieties allowing a meaning preserving translation between these varieties.
%An expression is given in a language variety.
%Such a variety can be a standard variety, a dialect, a script-based variety a controlled technical variety etc.
%Examples would be English, Simple English, British English and American English.\\

%  Ghidorah, the Three-Headed Monster | Ghidorah
%  describe or draw attention to (a product, service, or event) in a public medium in order to promote sales or attendance | advertise
%  move or surf through various files on a computer, the Internet, etc. | browse

\subsection{The PanLex vocabulary}
\label{sec:vocabulary}
The schema described above was adopted to create RDF triple structures.
Since the relational schema is a good starting point most of the relations were taken over introducing a PanLex RDF vocabulary.
Therefore the namespace \texttt{<http://panlex.org/plx/>} is used, abbreviated by the \texttt{plx} prefix.

The structure of \emph{approvers} is nearly the same as in the database schema.
All approvers share the \texttt{plx:Approver} class.
The characteristics of an approver are described using mainly triples with literal objects.
These are for example \texttt{plx:registrationDate} pointing to an \texttt{xsd:date} literal, \texttt{dc:title} to assign the title of a source, \texttt{dc:creator} to give an \texttt{xsd:string} containing the author's name etc.
All properties defined for \texttt{plx:Approver} instances are shown in \autoref{tbl:vocabulary}.

Besides the approvers there are also \emph{languages} and \emph{language varieties} taken over into the RDF vocabulary.
They are of the classes \texttt{plx:Language} and \texttt{plx:LanguageVariety} respectively.
Language varieties have a name given by the \texttt{rdfs:label} property and belong to a certain language referred to via \texttt{plx:languageVarietyOf}.
For those languages that have an ISO 639-3 language code, this code is referenced via the \texttt{plx:iso639-3Code} property.
The equivalent ISO 639-1 code can be retrieved using the \texttt{plx:iso639-1Code} property.

\emph{ISO 639-1} and \emph{ISO 639-3 codes} are of the classes \texttt{plx:Iso639-1Code} and \texttt{plx:Iso639-3Code} respectively.
These entities are mainly used for interlinking to the Lexvo dataset\footnote{\url{http://lexvo.org/ontology}}.
See \autoref{sec:linking} for more details.

As in the original PanLex database there are also \emph{expressions} used in the RDF dataset.
These resources have an \texttt{rdfs:label} containing the textual representation of the considered expression.
Additionally there is a normalized version of the label, reachable via the \texttt{plx:degradedText} property.
Expression entities also have an assigned language variety (\texttt{plx:languageVariety}).
It refers to the language variety the textual representation is written in.

The RDF analogon of the PanLex \emph{meaning} is the \texttt{plx:Meaning}.
Entities of this class may have an identifier assigned with the \texttt{plx:identifier} property pointing to an \texttt{xsd:string} literal.
This meaning identifier represents ids given to the considered meaning in other data sources and may be used to build links to the meaning as represented on these sites.
Meanings may also have \emph{definitions}, entities of the \texttt{plx:Definition} class, giving a textual representation (\texttt{rdfs:label}) with more than just a lexeme in a certain language variety (\texttt{plx:languageVariety}).

Following the semantics of the PanLex database, meanings and expressions are linked via \emph{denotations}.
These are entities of the \texttt{plx:Denotation} class pointing to meanings and expressions via the \texttt{plx:denotationMeaning} and \texttt{plx:denotationExpression} property respectively.
Denotations may also have a word class assigned to them.
This can be achieved with the denotation's \texttt{plx:wordClass} property pointing to a \texttt{plx:WordClass} entity.

Since the PanLex project compiled its database by extracting data from different sources, the licenses of these sources were considered, too.
At present, we support the different license categories given in the database by creating resources of the \texttt{plx:License} class.
In the RDF dataset these have an additional \texttt{rdfs:label} property giving the names or descriptions of the licenses or license categories.
\begin{table}
\begin{tabular}{ll}
Class & Properties \\
\toprule
\texttt{plx:Approver} & \texttt{plx:registrationDate} \\
                      & \texttt{rdfs:label} \\
                      & \texttt{foaf:homepage} \\
                      & \texttt{plx:isbn} \\
                      & \texttt{dc:creator} \\
                      & \texttt{dc:title} \\
                      & \texttt{dc:publisher} \\
                      & \texttt{plx:yearOfPublication} \\
                      & \texttt{plx:quality} \\
                      & \texttt{plx:license} \\
\midrule
\texttt{plx:Language}
                      & \texttt{plx:iso639-3Code} \\
                      & \texttt{plx:iso639-1Code} \\
\midrule
\texttt{plx:LanguageVariety}
                      & \texttt{plx:languageVarietyOf} \\
                      & \texttt{rdfs:label} \\
\midrule
\texttt{plx:Iso639-1Code} & \\
\midrule
\texttt{plx:Iso639-3Code} & \\
\midrule
\texttt{plx:Expression}
                      & \texttt{plx:languageVariety} \\
                      & \texttt{rdfs:label} \\
                      & \texttt{plx:degradedText} \\
\midrule
\texttt{plx:Meaning}
                      & \texttt{plx:approver} \\
                      & \texttt{plx:meaningDefinition} \\
                      & \texttt{plx:identifier} \\
\midrule
\texttt{plx:Definition}
                      & \texttt{plx:languageVariety} \\
                      & \texttt{rdfs:label} \\
\midrule
\texttt{plx:Denotation}
                      & \texttt{plx:denotationMeaning} \\
                      & \texttt{plx:denotationExpression} \\
                      & \texttt{plx:wordClass} \\
\midrule
\texttt{plx:WordClass}  
                      & \texttt{rdfs:label} \\
\midrule
\texttt{plx:License}  & \texttt{rdfs:label} \\
\bottomrule
\end{tabular}
\caption{Classes and properties used in the PanLex RDF vocabulary}
\label{tbl:vocabulary}
\end{table}

An overview of the PanLex vocabulary is shown in \autoref{fig:vocabulary}.%
\begin{figure}
\includegraphics[width=0.47\textwidth]{images/pdf/ontology.pdf}
\caption{Overview of the PanLex RDF vocabulary}
\label{fig:vocabulary}
\end{figure}
There, any \texttt{rdf:type} triples as well as facts assigning ISO 639-1 and ISO 639-3 labels are omitted for clarity.

\subsection{Transformation workflow}
\label{sec:conversion}
\todo{how often in PanLex updated?}
The size of the PanLex database (including indexes) is approximately 18 GB,
which is big enough to make recurrent RDF conversion cumbersome:
A full RDF conversion comes bloats in the dataset size and data duplication.   
Using conventional hardware, it takes quite long
to convert all the data, which usually makes
testing and debugging and fixing issues in the modelling or data conversion of the
data to become very time consuming.
As the PanLex data already resides in a relation database, the use of a virtual
RDB 2 RDF mapping solution is a natural choice. The \emph{Sparqlify Mapping Language} thereby
offers a very easy to use mapping language.
Essentially, these mappings consist of three clauses:
\begin{itemize}
  \item \emph{From} Specifies the logical SQL table (i.e. table, view or query) underlying the Sparqlify-ML view.
  \item \emph{With} Defines SPARQL variables by means of expressions over relational columns that yield RDF terms.
  \item \emph{Construct} The template (i.e. the set of triple patterns) to be generated in this view, based on the SPARQL variable definitions.
\end{itemize}
~\autoref{fig:ex:sparqlify-ml} shows an example of a Sparqlify-ML view for the
languages in PanLex: From each row of the \emph{i1} table three resources are
bound to the variable names \emph{?lang}, \emph{?iso3} and \emph{?lexvo3}.
Resources for \emph{?lang} become typed as a \emph{Language}in the PanLex and the \emph{schema.org} namespace.
\todo{Maybe we should link to lexvo directly.}
This view based approach also demonstrates that changing the vocubulary or
adding support for now ones do no require an ETL process, and can therefore be done with little efforts.
%PanLex's languages  iso 639-3 code 
%typed with plx:Language
%resources are created for
%the languages in PanLex, linked with \emph{lexvo}, and

%how meanings are mapped to their definition text:
%Two resources are generated, namely \texttt{?mn} and \texttt{?mt}, which are typed with \texttt{plx:Meaning} and \texttt{plx:MeaningText} respectively.
%Note that we could not simply use \texttt{rdfs:label} for associating the definition text to the meaning resource directly, because
%the ISO 639-3 language code and PanLex language varieties do not directly map to valid RDF language tags.\todo{cite the standard} 
%Instead, we retained the variety information by attaching it to the meaning text resource.
%The with clause specifies the creation of URIs in the panlex namespace based on the respective entities database IDs.
%Furthermore, it defines the definition text to become a plain literal.

\begin{figure}
\centering
\begin{lstlisting}
Create View i1 As
  Construct {
    ?lang a plx:Language ;
          a <http://schema.org/Language> ;
          plx:iso639-3Code ?iso3 .

    ?iso3 a plx:Iso639-3Code ;
          owl:sameAs ?lexvo3 .
  }
  With
    ?lang = uri(concat(plx:language, '/', ?iso3))
    ?iso3 = uri(concat(plx:iso639-3, '/', ?iso3))
    ?lexvo3 = uri(concat('http://lexvo.org/id/iso639-3/', ?iso3))
  From
    i1
\end{lstlisting}
\caption{An excerpt of a Sparqlify-ML view definition for PanLex's languages. This
example also demonstrates how ``is-a relations'' to schema.org and links to lexvo are established.}
\label{fig:ex:sparqlify-ml}
\end{figure}

\section{Linking}
\label{sec:linking}
In this section we first describe how we interlinked PanLex's language codes
with Lexvo\footnote{\url{http://www.lexvo.org/}} and the expressions with DBpedia\footnote{\url{http://dbpedia.org}}.
Afterwards we give an overview of how the dataset and the related resources are published.

Since Lexvo covers all ISO 639-3 language codes, interlinking could be done with
the Sparqlify-ML view in Figure~\ref{fig:ex:sparqlify-ml}.
For DBpedia we were interested in creating \emph{valid} and
thus \emph{dereferenceable} links.
Therefore, we iterated the \emph{titles} datasets\footnote{\url{http://wiki.dbpedia.org/Downloads38} - Note that more languages are available in the raw download section at \url{http://downloads.dbpedia.org/3.8/}},
which map non-localized DBpedia URIs to their page titles in the respective language.
For each language version we normalized the labels in the same way as being done in PanLex, i.e.
Unicode NFKD\footnote{\url{http://unicode.org/reports/tr15/}} normalization and
removal of punctuation characters are applied.
The DBpedia resources are then mapped to the PanLex expression that are equal to the resource's normalized label in the respective language.
Table ~\ref{fig:plx-dbp-link-counts} summarizes the number of links obtained.
\begin{table}
\centering
\begin{scriptsize}
\begin{tabular}{llll}
\toprule
Language & 639-1 & 639-3 & Links \\
\midrule
English    & en & eng & 1,415,241 \\
German     & de & deu &   224,146 \\
French     & fr & fra &   187,364 \\
Italian    & it & ita &   147,485 \\
Spanish    & sp & spa &   117,056 \\
Portuguese & pt & por &   112,266 \\
Polish     & pl & pol &   110,974 \\
Russian    & ru & rus &    68,040 \\
Czech      & cs & ces &    28,767 \\
Catalan    & ca & cat &    27,779 \\
Korean     & ko & kor &    24,912 \\
Turkish    & tr & tur &    22,258 \\
Bulgarian  & bg & bul &    19,431 \\
Hungarian  & hu & hun &    18,203 \\
Redo Slovene & sl & slv &  4,636 \\
Greek      & el & ell &  1,112 \\
\bottomrule
\end{tabular}
\end{scriptsize}
\caption{Number of DBpedia links per language}
\label{fig:plx-dbp-link-counts}
\end{table}

% // (i1)
% // ISO 639-3 language name links to lexvo.org
% Create View i1 As
%     Construct {
%         ?iso1 a plx:Iso639-1Code.
%         ?iso3 a plx:Iso639-3Code.
%         ?iso1 owl:sameAs ?lexvo1.
%         ?iso3 owl:sameAs ?lexvo3.
%         ?lang a plx:Language.
%         ?lang a lvont:Language.
%         ?lang a <http://schema.org/Language>.
%         ?lang plx:iso639-1Code ?iso1.
%         ?lang plx:iso639-3Code ?iso3.
%     }
%     With
%         ?iso1 = uri(concat(plx:iso639-1, '/', ?iso1))
%         ?iso3 = uri(concat(plx:iso639-3, '/', ?iso3))
%         ?lexvo1 = uri(concat('http://lexvo.org/id/iso639-1/', ?iso1))
%         ?lexvo3 = uri(concat('http://lexvo.org/id/iso639-3/', ?iso3))
%         ?lang = uri(concat(plx:language, '/', ?iso3))
%     From
%         i1
%

\section{Publishing}
\label{sec:publishing}
PanLex itself already provides direct access to the PostgreSQL database for scientists,
several user interfaces\footnote{\url{http://panlex.org/try/}} and an
API\footnote{\url{http://panlex.org/tech/doc/api/}}.
%The usage of these interfaces is not restricted in order to make the provided lexical
%translations as accessible as possible.
With our RDF conversion work, we complement these interfaces with
the Linked Data\footnote{\url{http://panlex.org/plx/TODO}} and the SPARQL
endpoint\footnote{\url{http://panlex.org/sparql} and \url{http://panlex.org/snorql}}.
An overview is shown in ~\autoref{fig:panlex-architecture}.
Our code for interlinking is
hosted on GitHub\footnote{\url{https://github.com/AKSW/PanLex-2-RDF}}.
\todo{Where to host the linksets?}

\begin{figure}
\centering
\includegraphics[width=0.4\textwidth]{images/PanLex-SW-Architecture}
\caption{PanLex architecture}
\label{fig:panlex-architecture}
\end{figure}

\todo{Fix the image: It should also show some sources which panlex uses.}

\section{Dataset Benefits and Usage Scenarios}
\label{sec:usage}
There are general benefits of RDF conversions,
such as the paradim shift towards thinking about what one intends to express
with the data, making this meaning explicit using ontologies,
enabling of interlinking, easing data integration tasks and uniform
data access via Linked Data and SPARQL.

With the RDF version of PanLex, a large lexical resource participates in the Web of Data,
making it available to a broader community. Users can easily explore the PanLex data using
the services provided by SNORQL and Pubby. This also increases the chance of development of
interesting mesh ups.
For example, the TeraDict translation service\footnote{\url{http://panlex.org/cgi-bin/apidemo-2.cgi}}
could be easily realized using simple SPARQL query as shown in
\todo{Create the query: translations in a given target language of an expression}. 
Anothe future usage scenario is to link PanLex to Wortschatz in order to 
improve the recall of word-by-word translations by considering co-occurences.

\section{Related Work}
\label{sec:related}
PanLex is an integration project of many existing lexical resources with
the goal of eventually covering all languages in the world, especially
ones having low numbers in their speakers.
As such, the extraction of information from linguistic sources, as well as
techniques for automatically inferring translations are
relevant work discussed in \cite{panlex_probtrans}.

An important initiative is the
\emph{Global Wordnet Association}\footnote{\url{http://www.globalwordnet.org/}},
which offers a platform for sharing word nets and defines several goals.
These include setting forth standards for uniformly representing wordnets
of different languages and establishing an universal index of meaning.

%It draws from the EuroWordNet~\cite{eurowordnet} and the  
%Manually engineered dictionaries %such as EuroWordNet [38]
% andcurrently mainly used for research purposes\todo{correct?}.
%A list of the resources used in panlex is compiled on the website\footnote{\url{http://panlex.org/tech/plrefs.shtml}}.
%In this section we discuss work related to the extraction and inference
%approaches used in PanLex, and work on the Semantic Web.

%- There has been considerable research on methods to acquire translation lexicons from either
%MRDs (Neff and McCord, 1990; Helmreich et
%al., 1993; Copestake et al., 1994) or from par-
%allel text (Gale and Church, 1991; Fung, 1995;
%Melamed, 1997; Franz et al., 2001), but this has
%generally been limited to a small number of lan-
%guages.
%- Manually engineered dictionaries such as
%EuroWordNet (Vossen, 1998) (whats the name of the new project?)
%- TRANS GRAPH algorithm (Etzioni etal., 2007).

At the level of the Semantic Web, there is currently a trend in publishing linguistic resources
as Linked Data\footnote{\url{http://www.w3.org/DesignIssues/LinkedData.html}} and via SPARQL{\footnote{\url{http://www.w3.org/TR/sparql11-query/}}}.
These efforts are referred to as the \emph{Linguistic Linked Open Data Cloud} (LLOD)\footnote{\url{http://linguistics.okfn.org/resources/llod/}}. 
%Some of these datasets already make use of quasi-standard linguistic Semantic Web Ontolgies, such as  
Thereby, several (quasi-)standard ontologies have been developed for covering different aspects of linguistic resources.
Examples include the \emph{Ontologies of Linguistic Annotation} (OLiA)~\cite{olia2010},
the \emph{Lexicon Model for Ontologies} (lemon)~\cite{lemon2011} for modeling lexicon and machine-readable dictionaries,
\emph{POWLA} for modeling linguistic corpora\cite{powla2012}
and the \emph{Natural Language Processing Interchange Format} (NIF)\footnote{\url{http://nlp2rdf.org/nif-1-0}}.

%\begin{itemize}
%\item @author: Claus, Patrick
%\item Research related to PanLex itself \url{http://panlex.org/research/}
%\item Panlingual Translator: Word by word translation + back translation for verification, broken English (no grammar) ranked by
%Do we have a comparison to wiktionary?

%Difference to dict.cc\footnote{\url{http://dict.cc}}, leo.org\footnote{\url{http://leo.org}}, Wiktionary\footnote{\url{http://wiktionary.org}} and Wikipedia (DBpedia)

%\item

%\item Research in the same field as PanLex
%\end{itemize}

\section{Conclusions and Future Work}
\label{sec:conclusion}
In this paper we describe the PanLex database and its conversion to RDF.
Based on our URI and vocabulary design, we created appropriate view definitions
for the virtual RDB to RDF solution \emph{Sparqlify} which carries out
the actualy RDF transformation.
Furthermore, we interlinked the languages in PanLex with lexvo,
and created a total of \todo {number} derefenceable
links to DBpedia for expressions in 16 languages.
%We detailed the developed vocabulary and

In this work we also identified some weaknesses in
the orginal database design,
which we intend to overcome in the future:
The orginal dataset currently does not cleanly model information sources and approvers
as distinct entities.\todo{We should still map it like that with Sparqlify - resuse ap for the source}
Also, while it is possible to determine
a meaning's source by its approver, this is not possible for the denotations.
Retaining this information seems beneficial for translation approches,
as this enables for example the exclusion of a poor quality or for a certain task
inappropriate source.

%\todo{Is there some quasi standard translation vocab we could re-use?}
%\todo{lang codes for minor languages}

\bibliographystyle{abbrv}
\bibliography{bibliography,../../bib/aksw}
\end{document}

### 2013-01-01 paper/2012/SWJ-PanLex/paper.tex
Added Jonathan's suggestions

@@
 \begin{abstract}
-At present, there are approximately 6,900 actively spoken languages in the world.
-However, some experts claim, that the process of globalization may eventually lead
-to a world loosing this linguistic diversity.
-The vision of the PanLex project is to save these languages, especially low density ones,
-by allowing them to be a part of the Information Age by making them applicable via translations.
-For this reason, PanLex gathers and integrates information from several
-linguistict sources, such as monolingual dictionaries, bilingual dictionaries, multilingual dictionaries,
+At present, there are approximately 7,000 living languages in the world.
+However, some experts claim that the process of globalization may eventually lead
+to the world losing this linguistic diversity.
+The vision of the PanLex project is to save these languages, especially low-density ones,
+by allowing them to be intertranslatable and thus to be a part of the Information Age.
+For this reason, PanLex gathers and integrates information from thousands of
+linguistic resources, such as monolingual dictionaries, bilingual dictionaries, multilingual dictionaries,
 glossaries, standards and thesauri.
@@
 % * These combined forces have led to predictions that something between half and 90%
-%   of the world’s living languages will die within the next century (Woodbury 2006; UNESCO 2003: 2).
+%   of the world's living languages will die within the next century (Woodbury 2006; UNESCO 2003: 2).
 %   Weaker forms of these forces appear to be shrinking the use of medium-density languages in science,
@@
 % * When languages die, they argue, the world loses
-%   (1) irreplaceable knowledge of history, medicine, nature, and productive methods encoded in languages’
+%   (1) irreplaceable knowledge of history, medicine, nature, and productive methods encoded in languages'
 % * extraction + integration
 % \end{itemize}
-At present, there are about 7,000 languages in the world with approximately 6,900 of them being actively spoken\footnote{see \url{http://www.sil.org/iso639-3/download.asp} for a list of registered languages}.
-Nonetheless, some experts claim, that the process of globalization may lead to a monolingual world loosing this linguistic diversity~\cite{endang_lang}.
-Since there are a lot of low density languages that were not taken over from generation to generation anymore, we could irreversibly loose parts of human knowledge.
-Some predictions state that this decline may lead to a loss of half to 90\% of human's languages~\cite{endang_lang}.
+At present, there are about 7,000 living languages in the world\footnote{See \url{http://www.sil.org/iso639-3/download.asp} for a list of registered languages}.
+Nonetheless, some experts claim that processes such as nation-state consolidation and globalization are producing language attrition so rapidly that up to 90\% of all languages alive today will be extinct within a century~\cite{endang_lang}.
+Theorists of biolinguistic diversity~\cite{nettle} argue that the loss of
+language diversity, the loss of human biological knowledge, and the loss of
+species diversity are mutually supportive and thus that language preservation
+and revitalization are essential to the preservation of biological diversity.
+The vision of the PanLex project is to save all these thousands of languages,
+especially those low-density ones that are threatened by extinction,
+by supporting their use in global communication. This requires panlingual
+translation: translation from any language of the world into any other.
+One of the crucial components of panlingual translation of discourses is
+panlingual lexical translation. PanLex is designed to support that component.
+It documents the known lexical translations (translations of lexemes) among all
+languages.
 
-The vision of the PanLex project is to save all these thousands of languages, especially those low-density ones that are threatened by extinction.
-PanLex tries to provide means for panlingual translations from any language of the world into any other.
-But it is not feasible to just store all the translations of one word into any other language since this would result in trillions of translations.
-So the PanLex project additionally uses artificial intelligence techniques to infer translations.
-As a consequence, not just mainstream languages like English, but even these low-density languages may be used for a global communication.
-% Since it is not feasible to just store all the translations of one word into any other language, which would result in trillions of translations, a system is needed that allows translations by inference using artificial intelligence.
+Although a list of all the translations of all words into all other languages
+would be large (amounting to trillions of translations), there are more
+serious obstacles. The vast majority of these translations are unknown.
+Those that are known are not universally agreed on. And lexical polysemy
+and ambiguity make the question ``What is the translation of lexeme X into language Y?''
+underspecified for practical purposes.
+So the PanLex project is designed to permit the application of artificial
+intelligence techniques to infer and select translations.
 
-Currently there is a growing community trying to combine linguistic knowledge with Semantic Web technologies and so build a web of linguistic data, also known as the \emph{Linguistic Linked Open Data (LLOD) cloud}.
-Originally the PanLex project provides different web services\footnote{\url{http://panlex.org/try/}} and APIs\footnote{\url{http://panlex.org/tech/doc/api/}} to access its data.
-Since the project shares the idea of open access we tried to make PanLex a part of the LLOD cloud.
+Currently there is a growing community trying to combine linguistic knowledge
+with Semantic Web technologies and thereby build a Web of Linguistic Data,
+also known as the \emph{Linguistic Linked Open Data (LLOD) cloud}.
+The PanLex project is still in its data-acquisition phase, but it has provided a
+few web services\footnote{\url{http://panlex.org/try/}} and
+APIs\footnote{\url{http://panlex.org/tech/doc/api/}} to access its data.
+Since the project shares the idea of open access, we have undertaken the project
+of making PanLex a part of the LLOD cloud.
 
-In \autoref{sec:triplificate} we will introduce the PanLex dataset, present our PanLex RDF vocabulary and explain how we transformed the one into the other.
-The following section is about the actual dataset publishing and how we linked to other datasets of the LLOD cloud.
+In \autoref{sec:triplificate} we will introduce the PanLex dataset, present our
+PanLex RDF vocabulary and explain how we transformed the one into the other.
+The following section is about the actual dataset publishing and how we linked
+to other datasets of the LLOD cloud.
 Finally we discuss related work, conclude our approach and give some hints to future work.
@@
 In this section we first provide an analysis of the original PanLex data.
-Subsequently, we will explain the steps taken to publish the data as RDF.
-The vocabulary introduced for the PanLex dataset will be described and afterwards we will show how the original data was converted to RDF and published via a SPARQL endpoint.
-The interlinking of the created RDF data with other datasets will be discussed in the following section.
+Subsequently, we introduce our URI and vocabulary design and we explain 
+the steps taken to publish the data as RDF.
+%The vocabulary 
+%afterwards we show how the original data was converted to RDF and published
+%via a SPARQL endpoint.
+%The interlinking of the created RDF data with other datasets is discussed in the
+%following section.
 
@@
 At the core, PanLex is a \emph{database}, served by a PostgreSQL database server, that ``represents \emph{assertions} about the \emph{meanings} of \emph{expressions}''\footnote{\url{http://panlex.org/tech/doc/design/panlex-db-design.pdf}}.
-The database contains about 20 million meanings and 19 million expressions extracted from about 4.000 sources.
+As of now, the database contains about 20 million meanings and 19 million expressions extracted from about 2,000 sources.
 
 Assertions are factual claims attributed to a certain source and editor.
-They do not represent absolute truths and as such it is valid that different editors make contradictory assertions on different sources.
+They do not represent absolute truths and as such it is valid that different editors make contradictory assertions based on the same sources.
 
@@
 The lexical database of the PanLex project is based on different sources like monolingual dictionaries,
-bilingual dictionaries, multilingual dictionaries, glossaries, standards, thesauri and so on.
+bilingual dictionaries, multilingual dictionaries, glossaries, standards, and thesauri.
 The list of used sources can also be found on the web\footnote{\url{http://panlex.org/tech/plrefs.shtml}}.
@@
   \item \emph{Languages} in PanLex are identified using ISO 639-3\footnote{\url{http://www.sil.org/iso639-3/codes.asp}} individual and macrolanguage codes, ISO 639-2\footnote{\url{http://www.loc.gov/standards/iso639-2/}} collective codes and ISO 639-5\footnote{\url{http://www.loc.gov/standards/iso639-5/}} codes.
-  \item \emph {Language Varieties} within a language are further distinguished by custom names.
+  \item \emph {Language Varieties} within a language are further distinguished by arbitrary unique integer identifiers.
   For example, ``eng'' is the ISO 639-3 code for the English language.
-  Panlex defines ``simple english'' and ``british english'' as varieties of it.
-  These can be given in the native writing system.
+  Panlex defines various varieties of it, including ``English'' (eng-000), ``Simple English'' (eng-001) and ``British English'' (eng-005).
+  Their labels, when possible, are autonyms, written in the native writing system.
   So, in contrast to mechanisms like IETF BCP 47\footnote{\url{http://tools.ietf.org/html/bcp47}} there is no need for a transcription. %  and one can use variety names like ``{زبان اوستایی}''.
-  Additionally these variety names do not have to be registered like the IETF BCP 47 language subtags\footnote{\url{http://www.iana.org/assignments/language-subtag-registry}}.
-  \item \emph{Meanings} of expressions are stated by users based on their
+  Additionally these variety identifiers and labels do not have to be registered like the IETF BCP 47 language subtags\footnote{\url{http://www.iana.org/assignments/language-subtag-registry}}.
+  \item \emph{Meanings} of expressions are distinguished by editors based on their
   interpretation of an information source.
-  Information sources can be dictionaries, thesauri but also personal knowledge. 
-  The pair comprised of the user and the information source is referred to as a meaning's \emph{approver}.
-  Users may add informal definitions to meanings in any language variety.
+  Information sources can be dictionaries, thesauri, etc., and also personal knowledge. 
+  The pair comprised of the editor and the information source is referred to as a meaning's \emph{approver}.
+  Editors may add informal definitions to meanings in any language variety.
   \item \emph{Definitions} are optional descriptions of a meaning.
   They are given in a certain language variety.
-  A description of the verb \emph{browse} for example could be \emph{``move or surf through various files on a computer, the Internet, etc.''}.
-  \item \emph{Denotations} relate expressions to possible \emph{meanings}.
+  A description of the verb \emph{browse} for example could be \emph{``move or surf through various files on a computer, the Internet, etc.''}, marked as a definition in the ``English'' language variety.
+  \item \emph{Denotations} relate expressions to \emph{meanings}.
   So homonyms are represented by one expression being connected to different meanings.
-  Denotations may be enriched with parts of speech tags adopted from the
+  Denotations may be enriched with part-of-speech tags selected from a closed list based on the
   Open Lexicon Interchange Format (OLIF) standard\footnote{\url{http://www.olif.net/}}.
   For example, \emph{fall} can be a verb or a noun for autumn.
-  Furthermore, it is possible to generically tag denotations with meta data
+  Furthermore, it is possible to generically tag denotations with metadata
   using key-value pairs.
@@
   \emph{police officer} could be annotated with \emph{pragmatics=vulgar}.
-  \item \emph{Users} are related to their defined language varieties and edited approvers.
+  \item \emph{Users} have editorial privileges over the language varieties and the approvers that they define.
   \item \emph{Licenses} are also considered by the PanLex project.
-  At present there are ten different license categories an approver can get into.
-  They are \emph{public domain}, \emph{Creative Commons}, \emph{request} (meaning that one has to ask the author of the source at hand), \emph{GNU General Public License}, \emph{GNU Lesser General Public License}, \emph{GNU Free Documentation License}, \emph{MIT License}, \emph{copyright} (stating that there is a certain copyright holder), \emph{other} and \emph{unknown}.
+  At present there are ten different license categories an approver can be annotated with.
+  They are \emph{public domain}, \emph{Creative Commons}, \emph{request} (meaning that one has to ask the author of the resource), \emph{GNU General Public License}, \emph{GNU Lesser General Public License}, \emph{GNU Free Documentation License}, \emph{MIT License}, \emph{copyright} (stating that there is a certain copyright holder), \emph{other} and \emph{unknown}.
 \end{itemize}
 
-An overview of the number of instances per entity in the PanLex database is given in~\autoref{fig:plx-entity-counts}.
+An overview of the number of instances per entity in the current PanLex database is given in~\autoref{fig:plx-entity-counts}.
 \begin{table}
@@
 Approver combines a user with an information source, however the \emph{information source} is not modelled as a distinct entity.
-Users are only associated with language variety and approver entities. However,
-it is not possible to track which denotations were added by users.
+Users are only associated with language variety and approver entities.
+However, whenever a given approver has more than one user with editorial privileges, it is not possible to track which denotations were added by which of the approver's authorized users.
 \todo{What are the consequences? }
@@
 
-The RDF analogon of the PanLex \emph{meaning} is the \texttt{plx:Meaning}.
+The RDF analog of the PanLex \emph{meaning} is the \texttt{plx:Meaning}.
 Entities of this class may have an identifier assigned with the \texttt{plx:identifier} property pointing to an \texttt{xsd:string} literal.
@@
 which is big enough to make recurrent RDF conversion cumbersome:
-A full RDF conversion comes bloats in the dataset size and data duplication.   
-Using conventional hardware, it takes quite long
-to convert all the data, which usually makes
+A full RDF conversion bloats the dataset and results in data duplication.   
+Using conventional hardware, it takes impractically long
+to convert all the data. This makes
 testing and debugging and fixing issues in the modelling or data conversion of the
-data to become very time consuming.
-As the PanLex data already resides in a relation database, the use of a virtual
-RDB 2 RDF mapping solution is a natural choice. The \emph{Sparqlify Mapping Language} thereby
-offers a very easy to use mapping language.
+data very time-consuming.
+As the PanLex data already resides in a relational database, the use of a virtual
+RDB 2 RDF mapping solution is a natural choice. The \emph{Sparqlify Mapping Language}
+offers a very easy-to-use mapping language.
 Essentially, these mappings consist of three clauses:
@@
 \todo{Maybe we should link to lexvo directly.}
-This view based approach also demonstrates that changing the vocubulary or
-adding support for now ones do no require an ETL process, and can therefore be done with little efforts.
+This view-based approach also demonstrates that changing the vocubulary or
+adding support for new ones does not require an ETL process, and can therefore be done with little effort.
 %PanLex's languages  iso 639-3 code 
@@
 which map non-localized DBpedia URIs to their page titles in the respective language.
-For each language version we normalized the labels in the same way as being done in PanLex, i.e.
+For each language version we normalized the labels: 
 Unicode NFKD\footnote{\url{http://unicode.org/reports/tr15/}} normalization and
-removal of punctuation characters are applied.
-The DBpedia resources are then mapped to the PanLex expression that are equal to the resource's normalized label in the respective language.
+removal of punctuation characters were applied.
+% PanLex now uses NFC, not NFKD, for character normalization and does not remove punctuation.
+The DBpedia resources were then mapped to the PanLex expression that was equal to the resource's normalized label in the respective language.
 Table ~\ref{fig:plx-dbp-link-counts} summarizes the number of links obtained.
@@
 the services provided by SNORQL and Pubby. This also increases the chance of development of
-interesting mesh ups.
+interesting mashups.
 For example, the TeraDict translation service\footnote{\url{http://panlex.org/cgi-bin/apidemo-2.cgi}}
@@
 the goal of eventually covering all languages in the world, especially
-ones having low numbers in their speakers.
-As such, the extraction of information from linguistic sources, as well as
-techniques for automatically inferring translations are
+ones having few speakers.
+The extraction of information from linguistic sources, and
+techniques for automatically inferring translations, are
 relevant work discussed in \cite{panlex_probtrans}.
@@
 \emph{Global Wordnet Association}\footnote{\url{http://www.globalwordnet.org/}},
-which offers a platform for sharing word nets and defines several goals.
+which offers a platform for sharing wordnets and defines several goals.
 These include setting forth standards for uniformly representing wordnets
-of different languages and establishing an universal index of meaning.
+of different languages and establishing a universal index of meaning.
 
@@
 %Some of these datasets already make use of quasi-standard linguistic Semantic Web Ontolgies, such as  
-Thereby, several (quasi-)standard ontologies have been developed for covering different aspects of linguistic resources.
+Several (quasi-)standard ontologies have been developed for covering different aspects of linguistic resources.
 Examples include the \emph{Ontologies of Linguistic Annotation} (OLiA)~\cite{olia2010},
@@
 for the virtual RDB to RDF solution \emph{Sparqlify} which carries out
-the actualy RDF transformation.
+the actual RDF transformation.
 Furthermore, we interlinked the languages in PanLex with lexvo,
@@
 as distinct entities.\todo{We should still map it like that with Sparqlify - resuse ap for the source}
-Also, while it is possible to determine
-a meaning's source by its approver, this is not possible for the denotations.
+Also, while it is possible to determine each meaning's and each denotation's approver, an approver can have multiple users with editorial privileges, so it is not possible to track each contributed, modified, and deleted meaning and denotation by editor.
 Retaining this information seems beneficial for translation approches,
-as this enables for example the exclusion of a poor quality or for a certain task
-inappropriate source.
+as this enables for example the attribution of qualities and relevances to particular editors.
 

### 2013-01-01 paper/2012/SWJ-PanLex/clean-latex.sh

@@ -1,0 +1,3 @@
+#!/bin/bash
+rm *.toc *.out *.aux *.bbl *.blg *.log *.lof *.lot *.eps
+rm paper.pdf

### 2013-01-01 paper/2012/SWJ-PanLex/paper.tex
Reworking the vocabulary design section

@@
 %\newcommand{\textrdf}[1]{\texttt{#1}}
-\newcommand{\textrdf}[1]{\emph{#1}}
+\newcommand{\vocab}[1]{\emph{#1}}
+
+\usepackage[scaled]{beramono}
+\newcommand\Small{\fontsize{9}{9.2}\selectfont}
+%\newcommand*\LSTfont{\Small\ttfamily\SetTracking{encoding=*}{-60}\lsstyle}
 
 \usepackage{listings}
+%\lstset{language=Python}
+%\lstset{%
+% morekeywords={Create,View,As,Construct,With,From}
+%}%
+\lstset{emph={%  
+    Create,View,As,Construct,With,From%
+    },emphstyle={\color{blue}\bfseries}%
+}%
+
 %\usepackage{verbatim}
@@
 Nonetheless, some experts claim that processes such as nation-state consolidation and globalization are producing language attrition so rapidly that up to 90\% of all languages alive today will be extinct within a century~\cite{endang_lang}.
-Theorists of biolinguistic diversity~\cite{nettle} argue that the loss of
+Theorists of biolinguistic diversity argue that the loss of
 language diversity, the loss of human biological knowledge, and the loss of
 species diversity are mutually supportive and thus that language preservation
-and revitalization are essential to the preservation of biological diversity.
+and revitalization are essential to the preservation of biological diversity~\cite{nettle}.
 The vision of the PanLex project is to save all these thousands of languages,
@@
   Expressions are always given in a language variety and can only be given once per language variety.
-  They were extracted from the above mentioned mono- or bilingual dictionaries or from other sources.
+  They were extracted from the afore mentioned mono- or bilingual dictionaries or from other sources.
   \item \emph{Languages} in PanLex are identified using ISO 639-3\footnote{\url{http://www.sil.org/iso639-3/codes.asp}} individual and macrolanguage codes, ISO 639-2\footnote{\url{http://www.loc.gov/standards/iso639-2/}} collective codes and ISO 639-5\footnote{\url{http://www.loc.gov/standards/iso639-5/}} codes.
@@
   interpretation of an information source.
-  Information sources can be dictionaries, thesauri, etc., and also personal knowledge. 
+%  Information sources can be dictionaries, thesauri, etc., and also personal knowledge. 
   The pair comprised of the editor and the information source is referred to as a meaning's \emph{approver}.
@@
   A description of the verb \emph{browse} for example could be \emph{``move or surf through various files on a computer, the Internet, etc.''}, marked as a definition in the ``English'' language variety.
-  \item \emph{Denotations} relate expressions to \emph{meanings}.
+  \item \emph{Denotations} relate expressions to meanings.
   So homonyms are represented by one expression being connected to different meanings.
@@
 \label{sec:vocabulary}
-The schema described above was adopted to create RDF triple structures.
-Since the relational schema is a good starting point most of the relations were taken over introducing a PanLex RDF vocabulary.
-Therefore the namespace \texttt{<http://panlex.org/plx/>} is used, abbreviated by the \texttt{plx} prefix.
+%The schema described above was adopted to create RDF triple structures.
+%is a good starting point most of the relations were taken over introducing a PanLex RDF vocabulary.
+The entities and relations of the schema described in the previous section
+serve as the base for the development of the PanLex RDF vocabulary.
+In general, all PanLex RDF resources reside
+in the namespace \texttt{<http://panlex.org/plx/>}, abbreviated with \texttt{plx}.
+An example of the resulting ontology is depicted in \autoref{tbl:vocabulary} and
+summarized as follows:
+Unless otherwise noted, the URIs of instances of PanLex classes follow the pattern
+\emph{plx:\{className\}/\{id\}}, where \{className\} is spelled in lower camel case
+and the \{id\} is the primary key of the corresponding database table.
 
-The structure of \emph{approvers} is nearly the same as in the database schema.
-All approvers share the \texttt{plx:Approver} class.
-The characteristics of an approver are described using mainly triples with literal objects.
-These are for example \texttt{plx:registrationDate} pointing to an \texttt{xsd:date} literal, \texttt{dc:title} to assign the title of a source, \texttt{dc:creator} to give an \texttt{xsd:string} containing the author's name etc.
-All properties defined for \texttt{plx:Approver} instances are shown in \autoref{tbl:vocabulary}.
+%As in the original PanLex database there are also \emph{expressions} used in the RDF dataset.
+Expressions are modeled as intances of \vocab{plx:Expression}.
+Their textual representations and normalized versions thereof become the 
+values of the properties \texttt{rdfs:label} and \texttt{plx:degradedText}, respectively.
+Their corresponding language variety is stated
+using \texttt{plx:languageVariety}.
 
-Besides the approvers there are also \emph{languages} and \emph{language varieties} taken over into the RDF vocabulary.
-They are of the classes \texttt{plx:Language} and \texttt{plx:LanguageVariety} respectively.
-Language varieties have a name given by the \texttt{rdfs:label} property and belong to a certain language referred to via \texttt{plx:languageVarietyOf}.
-For those languages that have an ISO 639-3 language code, this code is referenced via the \texttt{plx:iso639-3Code} property.
-The equivalent ISO 639-1 code can be retrieved using the \texttt{plx:iso639-1Code} property.
+%These resources have an \texttt{rdfs:label} containing the textual representation of the considered expression.
+%Additionally there is a normalized version of the label, reachable via the \texttt{plx:degradedText} property.
+%Expression entities also have an assigned language variety (\texttt{plx:languageVariety}).
+%It refers to the language variety the textual representation is written in.
 
-\emph{ISO 639-1} and \emph{ISO 639-3 codes} are of the classes \texttt{plx:Iso639-1Code} and \texttt{plx:Iso639-3Code} respectively.
-These entities are mainly used for interlinking to the Lexvo dataset\footnote{\url{http://lexvo.org/ontology}}.
-See \autoref{sec:linking} for more details.
+For Language and Language Varieties the classes
+texttt{plx:Language} and \texttt{plx:LanguageVariety} are introduced.
 
-As in the original PanLex database there are also \emph{expressions} used in the RDF dataset.
-These resources have an \texttt{rdfs:label} containing the textual representation of the considered expression.
-Additionally there is a normalized version of the label, reachable via the \texttt{plx:degradedText} property.
-Expression entities also have an assigned language variety (\texttt{plx:languageVariety}).
-It refers to the language variety the textual representation is written in.
+\emph{ISO 639-1} and \emph{ISO 639-3 codes} are of the classes
+\texttt{plx:Iso639-1Code} and \texttt{plx:Iso639-3Code} respectively.
 
@@
 
-Since the PanLex project compiled its database by extracting data from different sources, the licenses of these sources were considered, too.
+All approvers share the \texttt{plx:Approver} class. The characteristics of an approver are described using mainly triples with literal objects.
+These are for example \texttt{plx:registrationDate} pointing to an \texttt{xsd:date} literal,
+\texttt{dc:title} to assign the title of a source,
+\texttt{dc:creator} to give an \texttt{xsd:string} containing the author's name.
+
+%These entities are mainly used for interlinking to the Lexvo dataset\footnote{\url{http://lexvo.org/ontology}}.
+%See \autoref{sec:linking} for more details.
+
+%Besides the approvers there are also \emph{languages} and \emph{language varieties} taken over into the RDF vocabulary.
+%They are of the classes \texttt{plx:Language} and \texttt{plx:LanguageVariety} respectively.
+%Language varieties have a name given by the \texttt{rdfs:label} property and belong to a certain language referred to via \texttt{plx:languageVarietyOf}.
+%For those languages that have an ISO 639-3 language code, this code is referenced via the \texttt{plx:iso639-3Code} property.
+%The equivalent ISO 639-1 code can be retrieved using the \texttt{plx:iso639-1Code} property.
+
+%The structure of \emph{approvers} is nearly the same as in the database schema.
+% etc.
+%All properties defined for \texttt{plx:Approver} instances are shown in \autoref{tbl:vocabulary}.
+
+%Besides the approvers there are also \emph{languages} and \emph{language varieties} taken over into the RDF vocabulary.
+%They are of the classes \texttt{plx:Language} and \texttt{plx:LanguageVariety} respectively.
+%Language varieties have a name given by the \texttt{rdfs:label} property and belong to a certain language referred to via \texttt{plx:languageVarietyOf}.
+%For those languages that have an ISO 639-3 language code, this code is referenced via the \texttt{plx:iso639-3Code} property.
+%The equivalent ISO 639-1 code can be retrieved using the \texttt{plx:iso639-1Code} property.
+
+Since the PanLex project compiled its database by extracting data from different sources, the licenses of these sources were also considered.
 At present, we support the different license categories given in the database by creating resources of the \texttt{plx:License} class.
-In the RDF dataset these have an additional \texttt{rdfs:label} property giving the names or descriptions of the licenses or license categories.
+%with appropriate descriptions in their \texttt{rdfs:label}.
+%In the RDF dataset these have an additional \texttt{rdfs:label} property giving the names or descriptions 
 \begin{table}
+\begin{scriptsize}
 \begin{tabular}{ll}
@@
 \end{tabular}
-\caption{Classes and properties used in the PanLex RDF vocabulary}
+\end{scriptsize}
+\caption{Classes and properties used in the PanLex RDF vocabulary. Note that all \texttt{rdf:type} and \texttt{rdfs:label} properties have been omitted for brevity.}
 \label{tbl:vocabulary}
@@
 \end{figure}
-There, any \texttt{rdf:type} triples as well as facts assigning ISO 639-1 and ISO 639-3 labels are omitted for clarity.
 
@@
 \begin{itemize}
-  \item \emph{From} Specifies the logical SQL table (i.e. table, view or query) underlying the Sparqlify-ML view.
-  \item \emph{With} Defines SPARQL variables by means of expressions over relational columns that yield RDF terms.
-  \item \emph{Construct} The template (i.e. the set of triple patterns) to be generated in this view, based on the SPARQL variable definitions.
+  \item \emph{From:} Specifies the logical SQL table (i.e. table, view or query) underlying the Sparqlify-ML view.
+  \item \emph{With:} Defines SPARQL variables by means of expressions over relational columns that yield RDF terms.
+  \item \emph{Construct:} The template (i.e. the set of triple patterns) to be generated in this view, based on the SPARQL variable definitions.
 \end{itemize}
@@
 \centering
-\begin{lstlisting}
+\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
 Create View i1 As

### 2013-01-01 paper/2012/SWJ-PanLex/paper.tex
Some more work

@@
 
-\title{PanLex}
-\runningtitle{PanLex}
-\todo{find a real title}
+\title{PanLex-RDF}
+\runningtitle{Countering language attrition with the Web of Data}
+%\todo{find a real title}
 
@@
 \author[A]{\fnms{Patrick} \snm{Westphal}},
-\author[A]{\fnms{Claus} \snm{Stadler}},
+\author[B]{\fnms{Claus} \snm{Stadler}},
+\author[C]{\fnms{Jonathan} \snm{Pool}},
+
 %\runningauthor{M. Martin et al.}
+
 \address[A]{University of Leipzig, Institute of Computer Science, AKSW, Postfach 100920, D-04009 Leipzig, Germany\\
 E-mail: \{lastname\}@informatik.uni-leipzig.de}
-\todo{the e-mail address pattern won't work for both of us}
+\address[B]{University of Leipzig}
+\address[C]{TODO}
+%\todo{the e-mail address pattern won't work for both of us}
 
@@
 Therefore, we iterated the \emph{titles} datasets\footnote{\url{http://wiki.dbpedia.org/Downloads38} - Note that more languages are available in the raw download section at \url{http://downloads.dbpedia.org/3.8/}},
-which map non-localized DBpedia URIs to their page titles in the respective language.
+which map (non-localized) DBpedia URIs to their page titles in the respective language.
 For each language version we normalized the labels: 
@@
 Hungarian  & hu & hun &    18,203 \\
-Redo Slovene & sl & slv &  4,636 \\
+Slovene    & sl & slv &    11,981 \\
 Greek      & el & ell &     1,112 \\
+\midrule
+Total      &    &     & 2,537,015 \\
 \bottomrule
@@
 With our RDF conversion work, we complement these interfaces with
-the Linked Data\footnote{\url{http://panlex.org/plx/TODO}} and the SPARQL
+the Linked Data
+%\footnote{\url{http://panlex.org/plx/TODO}}
+and the SPARQL
 endpoint\footnote{\url{http://panlex.org/sparql} and \url{http://panlex.org/snorql}}.
@@
 Furthermore, we interlinked the languages in PanLex with lexvo,
-and created a total of \todo {number} derefenceable
-links to DBpedia for expressions in 16 languages.
+and created about 2.5 million links to DBpedia for expressions in 16 languages.
 %We detailed the developed vocabulary and
@@
 The orginal dataset currently does not cleanly model information sources and approvers
-as distinct entities.\todo{We should still map it like that with Sparqlify - resuse ap for the source}
+as distinct entities.
 Also, while it is possible to determine each meaning's and each denotation's approver, an approver can have multiple users with editorial privileges, so it is not possible to track each contributed, modified, and deleted meaning and denotation by editor.

### 2013-01-01 src/main/java/org/aksw/panlex/interlinking/dbpedia/Interlink_DBpedia.java

@@ -122,7 +122,7 @@ public class Interlink_DBpedia {
 		
 		Map<String, String> datasetToLang = MapReader.readFile(new File("src/main/resources/datasets.tsv"));
 		List<LangDataset> fileList = new ArrayList<LangDataset>();
-		String basePath = "var/cache/Datasets/DBpedia/3.8/";
+		String basePath = "var/cache/datasets/dbpedia/3.8/";
 		for(Entry<String, String> entry : datasetToLang.entrySet()) {
 			String name = entry.getKey();
 			String lang = entry.getValue();

### 2013-01-01 paper/2012/SWJ-PanLex/paper.tex
Almost done

@@
 \usepackage{array}
+
+%\usepackage{authblk}
+
 \usepackage{textcomp}
@@
 
-\author[A]{\fnms{Patrick} \snm{Westphal}},
-\author[B]{\fnms{Claus} \snm{Stadler}},
-\author[C]{\fnms{Jonathan} \snm{Pool}},
+%\author[1]{\fnms{Patrick} \snm{Westphal}},
+%\author[1]{\fnms{Claus} \snm{Stadler}},
+%\author[2]{\fnms{Jonathan} \snm{Pool}}
+%\address[1]{University of Leipzig, Institute of Computer Science, AKSW}
+%\address[2]{TODO}
+
+%\address[1]{University of Leipzig, Institute of Computer Science, AKSW, Postfach 100920, D-04009 Leipzig, Germany}
+
+%\author[A]{Patrick Westphal\thanks{A.A@university.edu}},
+%\author[A]{Claus Stadler\thanks{cstadler@informatik.uni-leipzig.de}},
+%\author[B]{Jonathan Pool\thanks{todo}}
+%\address[A]{University of Leipzig, Institute of Computer Science, AKSW, Postfach 100920, D-04009 Leipzig, Germany\\ E-mail: \{lastname\}@informatik.uni-leipzig.de}
+%\address[A]{University of Leipzig, Institute of Computer Science, AKSW, Postfach 100920, D-04009 Leipzig, Germany}
+%\address[B]{TODO}
+
+%\author[A]{\fnms{Patrick} \snm{Westphal}}
+%\author[B]{\fnms{Claus} \snm{Stadler}}
 
@@
 
-\address[A]{University of Leipzig, Institute of Computer Science, AKSW, Postfach 100920, D-04009 Leipzig, Germany\\
-E-mail: \{lastname\}@informatik.uni-leipzig.de}
-\address[B]{University of Leipzig}
-\address[C]{TODO}
+%\address[B]{University of Leipzig}
+%\address[C]{TODO}
 %\todo{the e-mail address pattern won't work for both of us}
@@
 %\emph{Denotations} capture the mappings between expressions and their possible meanings.  
-
-There are two noteworthy limitations of the database schema:
-Approver combines a user with an information source, however the \emph{information source} is not modelled as a distinct entity.
-Users are only associated with language variety and approver entities.
-However, whenever a given approver has more than one user with editorial privileges, it is not possible to track which denotations were added by which of the approver's authorized users.
-\todo{What are the consequences? }
+Note that \emph{approver} combines a user with an \emph{information source},
+however the \emph{information source} is not modelled as a distinct entity.
+Also, the information of whether two meanings with different approvers
+are still equal is not being captured.
 
@@
 %As in the original PanLex database there are also \emph{expressions} used in the RDF dataset.
-Expressions are modeled as intances of \vocab{plx:Expression}.
+\begin{itemize}
+\item Expressions are modeled as intances of \vocab{plx:Expression}.
 Their textual representations and normalized versions thereof become the 
@@
 
+\item For language and language varieties the classes
+\texttt{plx:Language} and \texttt{plx:LanguageVariety} are introduced.
+\emph{ISO 639-1} and \emph{ISO 639-3 codes} are of the classes
+\texttt{plx:Iso639-1Code} and \texttt{plx:Iso639-3Code} respectively.
+
+\item The RDF analog of the PanLex \emph{meaning} is the \texttt{plx:Meaning}.
+Entities of this class may have an identifier assigned with the
+\texttt{plx:identifier} property pointing to an \texttt{xsd:string} literal.
+This meaning identifier represents ids given to the considered meaning in other
+data sources and may at a later stage be used to build links to the meaning
+as represented on these sites.
+Meanings may also have \emph{definitions}, entities of the
+\texttt{plx:Definition} class, giving a textual representation
+(\texttt{rdfs:label}) with more than just a lexeme in a certain
+language variety (\texttt{plx:languageVariety}).
+
 %These resources have an \texttt{rdfs:label} containing the textual representation of the considered expression.
@@
 %It refers to the language variety the textual representation is written in.
-
-For Language and Language Varieties the classes
-texttt{plx:Language} and \texttt{plx:LanguageVariety} are introduced.
-
-\emph{ISO 639-1} and \emph{ISO 639-3 codes} are of the classes
-\texttt{plx:Iso639-1Code} and \texttt{plx:Iso639-3Code} respectively.
-
-The RDF analog of the PanLex \emph{meaning} is the \texttt{plx:Meaning}.
-Entities of this class may have an identifier assigned with the \texttt{plx:identifier} property pointing to an \texttt{xsd:string} literal.
-This meaning identifier represents ids given to the considered meaning in other data sources and may be used to build links to the meaning as represented on these sites.
-Meanings may also have \emph{definitions}, entities of the \texttt{plx:Definition} class, giving a textual representation (\texttt{rdfs:label}) with more than just a lexeme in a certain language variety (\texttt{plx:languageVariety}).
-
-Following the semantics of the PanLex database, meanings and expressions are linked via \emph{denotations}.
+\item Following the semantics of the PanLex database, meanings and expressions are linked via \emph{denotations}.
 These are entities of the \texttt{plx:Denotation} class pointing to meanings and expressions via the \texttt{plx:denotationMeaning} and \texttt{plx:denotationExpression} property respectively.
@@
 This can be achieved with the denotation's \texttt{plx:wordClass} property pointing to a \texttt{plx:WordClass} entity.
-
-All approvers share the \texttt{plx:Approver} class. The characteristics of an approver are described using mainly triples with literal objects.
+\item All approvers share the \texttt{plx:Approver} class. The characteristics of an approver are described using mainly triples with literal objects.
 These are for example \texttt{plx:registrationDate} pointing to an \texttt{xsd:date} literal,
@@
 \texttt{dc:creator} to give an \texttt{xsd:string} containing the author's name.
+\end{itemize}
 
@@
 
-An overview of the PanLex vocabulary is shown in \autoref{fig:vocabulary}.%
+%An overview of the PanLex vocabulary is shown in \autoref{fig:vocabulary}.%
 \begin{figure}
@@
 \label{sec:conversion}
-\todo{how often in PanLex updated?}
-The size of the PanLex database (including indexes) is approximately 18 GB,
-which is big enough to make recurrent RDF conversion cumbersome:
-A full RDF conversion bloats the dataset and results in data duplication.   
+New sources are added to PanLex on almost a daily basis.   
+The size of the PanLex database (including indexes) is currently approximately
+18 GB, which is big enough to make recurrent RDF conversion cumbersome:
+The RDF files of a full conversion of a dataset of this size are very huge.
 Using conventional hardware, it takes impractically long
@@
 languages in PanLex: From each row of the \emph{i1} table three resources are
+created from the \emph{iso3} column and 
 bound to the variable names \emph{?lang}, \emph{?iso3} and \emph{?lexvo3}.
-Resources for \emph{?lang} become typed as a \emph{Language}in the PanLex and the \emph{schema.org} namespace.
-\todo{Maybe we should link to lexvo directly.}
+Resources for \emph{?lang} become typed as a \emph{Language} in the PanLex and the \emph{schema.org} namespace.
 This view-based approach also demonstrates that changing the vocubulary or
 adding support for new ones does not require an ETL process, and can therefore be done with little effort.
+%\todo{Maybe we should link to lexvo directly.}
 %PanLex's languages  iso 639-3 code 
@@
 hosted on GitHub\footnote{\url{https://github.com/AKSW/PanLex-2-RDF}}.
-\todo{Where to host the linksets?}
+The created linksets are hosted in the PanLex database and published together
+with the other data using Sparqlify.
 
@@
 For example, the TeraDict translation service\footnote{\url{http://panlex.org/cgi-bin/apidemo-2.cgi}}
-could be easily realized using simple SPARQL query as shown in
-\todo{Create the query: translations in a given target language of an expression}. 
+could be easily realized using simple SPARQL queries.
+% as shown in
+%\todo{Create the query: translations in a given target language of an expression}. 
 Anothe future usage scenario is to link PanLex to Wortschatz in order to 

### 2013-01-01 paper/2012/SWJ-PanLex/paper.tex
added approver count
My changes

@@
 
+%patrick.westphal@studserv.uni-leipzig.de
 %\author[1]{\fnms{Patrick} \snm{Westphal}},
@@
 In this paper we describe how we transformed this data to RDF, interlinked it with
-lexvo and DBpedia and published it as Linked Data and SPARQL.
+Lexvo and DBpedia and published it as Linked Data and SPARQL.
 \end{abstract}
@@
 \centering
-\includegraphics[width=0.4\textwidth]{images/PanLex-SW-Architecture}
+\includegraphics[width=0.4\textwidth]{images/PanLex-SW-Architecture-New}
 \caption{PanLex architecture}
@@
 \end{figure}
-
-\todo{Fix the image: It should also show some sources which panlex uses.}
 
@@
 the actual RDF transformation.
-Furthermore, we interlinked the languages in PanLex with lexvo,
+Furthermore, we interlinked the languages in PanLex with Lexvo,
 and created about 2.5 million links to DBpedia for expressions in 16 languages.

### 2013-01-02 paper/2012/SWJ-PanLex/paper.tex

@@
 %patrick.westphal@studserv.uni-leipzig.de
-%\author[1]{\fnms{Patrick} \snm{Westphal}},
-%\author[1]{\fnms{Claus} \snm{Stadler}},
-%\author[2]{\fnms{Jonathan} \snm{Pool}}
-%\address[1]{University of Leipzig, Institute of Computer Science, AKSW}
-%\address[2]{TODO}
+\author[1]{\fnms{Patrick} \snm{Westphal}},
+\author[2]{\fnms{Claus} \snm{Stadler}},
+\author[3]{\fnms{Jonathan} \snm{Pool}}
+\address[1]{University of Leipzig, patrick.westphal@studserv.uni-leipzig.de}
+\address[2]{University of Leipzig, cstadler@informatik.uni-leipzig.de}
+\address[3]{Long Now Foundation, San Francisco, pool@utilika.org}
+
+%\author[1]{Author A\thanks{A.A@university.edu}}
+%\author[1]{Author B\thanks{B.B@university.edu}}
+%\author[1]{Author C\thanks{C.C@university.edu}}
+%\author[2]{Author D\thanks{D.D@university.edu}}
+%\author[2]{Author E\thanks{E.E@university.edu}}
+%\affil[1]{Department of Computer Science, \LaTeX\ University}
+%\affil[2]{Department of Mechanical Engineering, \LaTeX\ University}
+
+%\address[1]{University of Leipzig, Institute of Computer Science, AKSW, patrick.westphal@studserv.uni-leipzig.de}
 
@@
 %\midrule
+Approvers & 3,905 \\ % SELECT COUNT(ex) FROM ex;
 Expressions & 18,580,594 \\ % SELECT COUNT(ex) FROM ex;
@@
 \texttt{plx:identifier} property pointing to an \texttt{xsd:string} literal.
-This meaning identifier represents ids given to the considered meaning in other
-data sources and may at a later stage be used to build links to the meaning
-as represented on these sites.
+%This meaning identifier represents ids given to the considered meaning in other
+%data sources and may at a later stage be used to build links to the meaning
+%as represented on these sites.
 Meanings may also have \emph{definitions}, entities of the

### 2013-01-02 paper/2012/SWJ-PanLex/paper.tex
Some changes at the end

@@
 however the \emph{information source} is not modelled as a distinct entity.
-Also, the information of whether two meanings with different approvers
-are still equal is not being captured.
+Also, the information of whether or not two meanings with different approvers
+are the same is not being captured.
 
@@
 18 GB, which is big enough to make recurrent RDF conversion cumbersome:
-The RDF files of a full conversion of a dataset of this size are very huge.
+The RDF file of a full conversion of a database is much larger than the
+database itself.
 Using conventional hardware, it takes impractically long
@@
 As the PanLex data already resides in a relational database, the use of a virtual
-RDB 2 RDF mapping solution is a natural choice. The \emph{Sparqlify Mapping Language}
-offers a very easy-to-use mapping language.
+RDB 2 RDF mapping solution is a natural choice.
+The \emph{Sparqlify system}\footnote{\url{https://github.com/AKSW/Sparqlify}} offers,
+besides an efficient query rewriting engine, also a very easy-to-use mapping
+language (ML).
 Essentially, these mappings consist of three clauses:
@@
 \label{sec:linking}
-In this section we first describe how we interlinked PanLex's language codes
-with Lexvo\footnote{\url{http://www.lexvo.org/}} and the expressions with DBpedia\footnote{\url{http://dbpedia.org}}.
-Afterwards we give an overview of how the dataset and the related resources are published.
+The Sparqlify view in the previous section ~\ref{fig:ex:sparqlify-ml} already
+established the interlinking of the PanLex languages with Lexvo.
+In this section we outline the interlinking with DBpedia.
 
-Since Lexvo covers all ISO 639-3 language codes, interlinking could be done with
-the Sparqlify-ML view in Figure~\ref{fig:ex:sparqlify-ml}.
+%In this section we first describe how we interlinked PanLex's language codes
+%with Lexvo\footnote{\url{http://www.lexvo.org/}} and the expressions with DBpedia\footnote{\url{http://dbpedia.org}}.
+%Afterwards we give an overview of how the dataset and the related resources are published.
+
+%Since Lexvo covers all ISO 639-3 language codes, interlinking could be done with
+%the Sparqlify-ML view in Figure~\ref{fig:ex:sparqlify-ml}.
 For DBpedia we were interested in creating \emph{valid} and
@@
 which map (non-localized) DBpedia URIs to their page titles in the respective language.
-For each language version we normalized the labels: 
+For each language version we normalized the labels by applying
 Unicode NFKD\footnote{\url{http://unicode.org/reports/tr15/}} normalization and
-removal of punctuation characters were applied.
+removal of punctuation characters.
 % PanLex now uses NFC, not NFKD, for character normalization and does not remove punctuation.
-The DBpedia resources were then mapped to the PanLex expression that was equal to the resource's normalized label in the respective language.
+Each DBpedia resource was then mapped to the PanLex expression that was equal to the resource's normalized label in the respective language.
 Table ~\ref{fig:plx-dbp-link-counts} summarizes the number of links obtained.
@@
 An overview is shown in ~\autoref{fig:panlex-architecture}.
-Our code for interlinking is
+Our Sparqlify-ML views and the code for interlinking are
 hosted on GitHub\footnote{\url{https://github.com/AKSW/PanLex-2-RDF}}.
@@
 the services provided by SNORQL and Pubby. This also increases the chance of development of
-interesting mashups.
+new interesting mashups.
 For example, the TeraDict translation service\footnote{\url{http://panlex.org/cgi-bin/apidemo-2.cgi}}
-could be easily realized using simple SPARQL queries.
+could now be easily realized using simple SPARQL queries.
 % as shown in
 %\todo{Create the query: translations in a given target language of an expression}. 
-Anothe future usage scenario is to link PanLex to Wortschatz in order to 
-improve the recall of word-by-word translations by considering co-occurences.
+Another future usage scenario is to link PanLex to Wortschatz with the aim of 
+improving the recall of word-by-word translations by considering the co-occurences.
 

### 2013-01-02 paper/2012/SWJ-PanLex/paper.tex
Title changed
made texttt environments small

--- 2013-01-02 paper/2012/SWJ-PanLex/paper.tex
@@
   Expressions are always given in a language variety and can only be given once per language variety.
-  They were extracted from the afore mentioned mono- or bilingual dictionaries or from other sources.
+  They were extracted from the afore-mentioned mono- or bilingual dictionaries or from other sources.
   \item \emph{Languages} in PanLex are identified using ISO 639-3\footnote{\url{http://www.sil.org/iso639-3/codes.asp}} individual and macrolanguage codes, ISO 639-2\footnote{\url{http://www.loc.gov/standards/iso639-2/}} collective codes and ISO 639-5\footnote{\url{http://www.loc.gov/standards/iso639-5/}} codes.
-  \item \emph {Language Varieties} within a language are further distinguished by arbitrary unique integer identifiers.
+  \item \emph {Language varieties} within a language are further distinguished by arbitrary unique integer identifiers.
   For example, ``eng'' is the ISO 639-3 code for the English language.
@@
 In general, all PanLex RDF resources reside
-in the namespace \texttt{<http://panlex.org/plx/>}, abbreviated with \texttt{plx}.
+in the namespace \texttt{\small <http://panlex.org/plx/>}, abbreviated with \texttt{\small plx}.
 An example of the resulting ontology is depicted in \autoref{tbl:vocabulary} and
@@
 Their textual representations and normalized versions thereof become the 
-values of the properties \texttt{rdfs:label} and \texttt{plx:degradedText}, respectively.
+values of the properties \texttt{\small rdfs:label} and \texttt{\small plx:degradedText}, respectively.
 Their corresponding language variety is stated
-using \texttt{plx:languageVariety}.
+using \texttt{\small plx:languageVariety}.
 
 \item For language and language varieties the classes
-\texttt{plx:Language} and \texttt{plx:LanguageVariety} are introduced.
+\texttt{\small plx:Language} and \texttt{\small plx:LanguageVariety} are introduced.
 \emph{ISO 639-1} and \emph{ISO 639-3 codes} are of the classes
-\texttt{plx:Iso639-1Code} and \texttt{plx:Iso639-3Code} respectively.
+\texttt{\small plx:Iso639-1Code} and \texttt{\small plx:Iso639-3Code} respectively.
 
-\item The RDF analog of the PanLex \emph{meaning} is the \texttt{plx:Meaning}.
+\item The RDF analog of the PanLex \emph{meaning} is the \texttt{\small plx:Meaning}.
 Entities of this class may have an identifier assigned with the
-\texttt{plx:identifier} property pointing to an \texttt{xsd:string} literal.
+\texttt{\small plx:identifier} property pointing to an \texttt{\small xsd:string} literal.
 %This meaning identifier represents ids given to the considered meaning in other
@@
 Meanings may also have \emph{definitions}, entities of the
-\texttt{plx:Definition} class, giving a textual representation
-(\texttt{rdfs:label}) with more than just a lexeme in a certain
-language variety (\texttt{plx:languageVariety}).
+\texttt{\small plx:Definition} class, giving a textual representation
+(\texttt{\small rdfs:label}) with more than just a lexeme in a certain
+language variety (\texttt{\small plx:languageVariety}).
 
@@
 \item Following the semantics of the PanLex database, meanings and expressions are linked via \emph{denotations}.
-These are entities of the \texttt{plx:Denotation} class pointing to meanings and expressions via the \texttt{plx:denotationMeaning} and \texttt{plx:denotationExpression} property respectively.
+These are entities of the \texttt{\small plx:Denotation} class pointing to meanings and expressions via the \texttt{\small plx:denotationMeaning} and \texttt{\small plx:denotationExpression} property respectively.
 Denotations may also have a word class assigned to them.
-This can be achieved with the denotation's \texttt{plx:wordClass} property pointing to a \texttt{plx:WordClass} entity.
-\item All approvers share the \texttt{plx:Approver} class. The characteristics of an approver are described using mainly triples with literal objects.
-These are for example \texttt{plx:registrationDate} pointing to an \texttt{xsd:date} literal,
-\texttt{dc:title} to assign the title of a source,
-\texttt{dc:creator} to give an \texttt{xsd:string} containing the author's name.
+This can be achieved with the denotation's \texttt{\small plx:wordClass} property pointing to a \texttt{\small plx:WordClass} entity.
+\item All approvers share the \texttt{\small plx:Approver} class. The characteristics of an approver are described using mainly triples with literal objects.
+These are for example \texttt{\small plx:registrationDate} pointing to an \texttt{\small xsd:date} literal,
+\texttt{\small dc:title} to assign the title of a source,
+\texttt{\small dc:creator} to give an \texttt{\small xsd:string} containing the author's name.
 \end{itemize}
@@
 Since the PanLex project compiled its database by extracting data from different sources, the licenses of these sources were also considered.
-At present, we support the different license categories given in the database by creating resources of the \texttt{plx:License} class.
+At present, we support the different license categories given in the database by creating resources of the \texttt{\small plx:License} class.
 %with appropriate descriptions in their \texttt{rdfs:label}.
@@
 \end{scriptsize}
-\caption{Classes and properties used in the PanLex RDF vocabulary. Note that all \texttt{rdf:type} and \texttt{rdfs:label} properties have been omitted for brevity.}
+\caption{Classes and properties used in the PanLex RDF vocabulary. Note that all \texttt{\small rdf:type} and \texttt{\small rdfs:label} properties have been omitted for brevity.}
 \label{tbl:vocabulary}

### 2013-01-02 paper/2012/SWJ-PanLex/paper.tex
Cosmetics

@@
 
-\title{PanLex-RDF}
-\runningtitle{Countering language attrition with the Web of Data}
+\title{Countering language attrition with PanLex and the Web of Data}
+\runningtitle{Countering language attrition with PanLex and the Web of Data}
 %\todo{find a real title}
@@
 \caption{An excerpt of a Sparqlify-ML view definition for PanLex's languages. This
-example also demonstrates how ``is-a relations'' to schema.org and links to lexvo are established.}
+example also demonstrates how ``is-a relations'' to schema.org and links to Lexvo are established.}
 \label{fig:ex:sparqlify-ml}
@@
 \begin{scriptsize}
-\begin{tabular}{llll}
+\begin{tabular}{lllr}
 \toprule

### 2013-01-02 paper/2012/SWJ-PanLex/paper.tex
added manual linebreaks, minor enhancements

@@
 in the namespace \texttt{\small <http://panlex.org/plx/>}, abbreviated with \texttt{\small plx}.
-An example of the resulting ontology is depicted in \autoref{tbl:vocabulary} and
+An example of the resulting ontology is depicted in \autoref{fig:vocabulary} and
 summarized as follows:
@@
 \begin{itemize}
-\item Expressions are modeled as intances of \vocab{plx:Expression}.
+\item Expressions are modeled as instances of \\ \texttt{\small plx:Expression}. % manual linebreak
 Their textual representations and normalized versions thereof become the 
-values of the properties \texttt{\small rdfs:label} and \texttt{\small plx:degradedText}, respectively.
-Their corresponding language variety is stated
-using \texttt{\small plx:languageVariety}.
+values of the properties \texttt{\small rdfs:label} and \\ \texttt{\small plx:degradedText}, respectively. % manual linebreak
+Their corresponding language variety is stated using \\ \texttt{\small plx:languageVariety}. % manual linebreak
 
@@
 \emph{ISO 639-1} and \emph{ISO 639-3 codes} are of the classes
-\texttt{\small plx:Iso639-1Code} and \texttt{\small plx:Iso639-3Code} respectively.
+\texttt{\small plx:Iso639-1Code} and \\ \texttt{\small plx:Iso639-3Code} respectively. % manual linebreak
 
@@
 \item Following the semantics of the PanLex database, meanings and expressions are linked via \emph{denotations}.
-These are entities of the \texttt{\small plx:Denotation} class pointing to meanings and expressions via the \texttt{\small plx:denotationMeaning} and \texttt{\small plx:denotationExpression} property respectively.
+These are entities of the \texttt{\small plx:Denotation} class pointing to meanings and expressions via the \texttt{\small plx:denotationMeaning} and \\ \texttt{\small plx:denotationExpression} property respectively.
+% manual linebreak
 Denotations may also have a word class assigned to them.
@@
 \end{scriptsize}
-\caption{Classes and properties used in the PanLex RDF vocabulary. Note that all \texttt{\small rdf:type} and \texttt{\small rdfs:label} properties have been omitted for brevity.}
+\caption{Classes and properties used in the PanLex RDF vocabulary. Note that all \texttt{rdf:type} and \texttt{rdfs:label} properties have been omitted for brevity.}
 \label{tbl:vocabulary}
@@
 \caption{An excerpt of a Sparqlify-ML view definition for PanLex's languages. This
-example also demonstrates how ``is-a relations'' to schema.org and links to Lexvo are established.}
+example also demonstrates how ``is-a relations'' to schema.org and links to lexvo are established.}
 \label{fig:ex:sparqlify-ml}
@@
 \begin{scriptsize}
-\begin{tabular}{lllr}
+\begin{tabular}{llll}
 \toprule

### 2013-01-02 paper/2012/SWJ-PanLex/paper.tex
Hm, possible conflict

--- 2013-01-02 paper/2012/SWJ-PanLex/paper.tex
@@
 bilingual dictionaries, multilingual dictionaries, glossaries, standards, and thesauri.
-The list of used sources can also be found on the web\footnote{\url{http://panlex.org/tech/plrefs.shtml}}.
+The list of used sources can also be found on the Web\footnote{\url{http://panlex.org/tech/plrefs.shtml}}.
 
@@
 \begin{itemize}
-\item Expressions are modeled as intances of \vocab{plx:Expression}.
-Their textual representations and normalized versions thereof become the 
+\item Expressions are modeled as intances of the class \texttt{\small plx:Expression}.
+Their original and normalized textual representations become the 
 values of the properties \texttt{\small rdfs:label} and \texttt{\small plx:degradedText}, respectively.

### 2013-01-02 paper/2012/SWJ-PanLex/paper.tex
minor enhancements in linking section

--- 2013-01-02 paper/2012/SWJ-PanLex/paper.tex
@@
 \label{sec:linking}
-The Sparqlify view in the previous section ~\ref{fig:ex:sparqlify-ml} already
+The Sparqlify view in the previous section (\autoref{fig:ex:sparqlify-ml}) already
 established the interlinking of the PanLex languages with Lexvo.
@@
 Each DBpedia resource was then mapped to the PanLex expression that was equal to the resource's normalized label in the respective language.
-Table ~\ref{fig:plx-dbp-link-counts} summarizes the number of links obtained.
+\autoref{fig:plx-dbp-link-counts} summarizes the number of links obtained.
 \begin{table}

### 2013-01-02 paper/2012/SWJ-PanLex/paper.tex

--- 2013-01-02 paper/2012/SWJ-PanLex/paper.tex
@@
 \texttt{\small plx:Language} and \texttt{\small plx:LanguageVariety} are introduced.
-\emph{ISO 639-1} and \emph{ISO 639-3 codes} are of the classes
-\texttt{\small plx:Iso639-1Code} and \texttt{\small plx:Iso639-3Code} respectively.
+\emph{ISO 639-1} and \emph{ISO 639-3 codes} become instances of the classes
+\texttt{\small plx:Iso639-1Code} and \texttt{\small plx:Iso639-3Code}. % manual linebreak
 
@@
 \texttt{\small plx:Definition} class, giving a textual representation
-(\texttt{\small rdfs:label}) with more than just a lexeme in a certain
+(\texttt{\small rdfs:label})
+%with more than just a lexeme
+in a certain
 language variety (\texttt{\small plx:languageVariety}).
@@
 \item Following the semantics of the PanLex database, meanings and expressions are linked via \emph{denotations}.
-These are entities of the \texttt{\small plx:Denotation} class pointing to meanings and expressions via the \texttt{\small plx:denotationMeaning} and \texttt{\small plx:denotationExpression} property respectively.
+These are entities of the \texttt{\small plx:Denotation} class pointing
+to meanings and expressions via the properties \texttt{\small plx:denotationMeaning}
+and \\ \texttt{\small plx:denotationExpression}. % manual linebreak
 Denotations may also have a word class assigned to them.

### 2013-01-02 paper/2012/SWJ-PanLex/paper.tex

@@
 making it available to a broader community. Users can easily explore the PanLex data using
-the services provided by SNORQL and Pubby. This also increases the chance of development of
+the services provided by SNORQL\footnote{\url{https://github.com/kurtjx/SNORQL}} and Pubby\footnote{\url{http://wifo5-03.informatik.uni-mannheim.de/pubby/}}.
+This also increases the chance of development of
 new interesting mashups.
@@
 %\todo{Create the query: translations in a given target language of an expression}. 
-Another future usage scenario is to link PanLex to Wortschatz with the aim of 
+Another future usage scenario is to link PanLex to Wortschatz\footnote{\url{http://wortschatz.uni-leipzig.de/}} with the aim of 
 improving the recall of word-by-word translations by considering the co-occurences.
@@
 \label{sec:related}
-PanLex is an integration project of many existing lexical resources with
-the goal of eventually covering all languages in the world, especially
-ones having few speakers.
+PanLex is an integration project of many existing lexical resources. % with
+% the goal of eventually covering all languages in the world, especially
+% ones having few speakers.
 The extraction of information from linguistic sources, and
@@
 of different languages and establishing a universal index of meaning.
-
+%
 %It draws from the EuroWordNet~\cite{eurowordnet} and the  
@@
 %approaches used in PanLex, and work on the Semantic Web. 
-
+%
 %- There has been considerable research on methods to acquire translation lexicons from either
@@
 %- TRANS GRAPH algorithm (Etzioni etal., 2007).
-
+%
 At the level of the Semantic Web, there is currently a trend in publishing linguistic resources
-as Linked Data\footnote{\url{http://www.w3.org/DesignIssues/LinkedData.html}} and via SPARQL{\footnote{\url{http://www.w3.org/TR/sparql11-query/}}}.
-These efforts are referred to as the \emph{Linguistic Linked Open Data Cloud} (LLOD)\footnote{\url{http://linguistics.okfn.org/resources/llod/}}. 
+%as Linked Data\footnote{\url{http://www.w3.org/DesignIssues/LinkedData.html}} and via SPARQL{\footnote{\url{http://www.w3.org/TR/sparql11-query/}}}.
+as Linked Data\footnote{\url{http://www.w3.org/DesignIssues/LinkedData.html}} and via SPARQL.
+These efforts are referred to as the \emph{Linguistic Linked Open Data Cloud} (LLOD). 
 %Some of these datasets already make use of quasi-standard linguistic Semantic Web Ontolgies, such as  

### 2014-01-22 panlex-rdf-paper/2012/MLOD_PanLex/coverletter_v3.tex
Added the panlex paper to the repo

Again we thank the reviewers and editors for their constructive feedback on our submission.

Overall, we performed the following changes:
- Section 1 (Introduction): The description of PanLex was shortened. We emphasized the role of PanLex, as we see it, in the Web of Data.
- Section 2 (Triplification): We added a section on vocabulary reuse and added according depicitions. Added a table with approvers per license type.
- Section 4 (Publication): Improved descriptions of the published resources, especially that downloads are tagged with a timestamp of their creation. 
- Section 5 (Benefits and Usage Scenarios): Pointed to the web site for SPARQL query examples due to space reasons. Added cross language information retrieval and DBpedia spotlight (entity disambiguation) as potential applications.
- Section 6 (Related Work): Minor explanation of the relationship between wordnets and PanLex.
- Section 7 (Conclusions and Future Work): Improved descriptions of directions for future work, notably that descriptions of datasets within PanLex could be modeled with VoID.

### 2014-01-22 panlex-rdf-paper/2012/MLOD_PanLex/paper.tex
Added the panlex paper to the repo

--- 2013-01-02 paper/2012/SWJ-PanLex/paper.tex
@@
     numberstyle=\tiny,
-    basicstyle=\footnotesize\ttfamily,
+%    basicstyle=\footnotesize\ttfamily,
+    basicstyle=\ttfamily \tiny,
     tabsize=2,
@@
 %patrick.westphal@studserv.uni-leipzig.de
-\author[1]{\fnms{Patrick} \snm{Westphal}},
-\author[2]{\fnms{Claus} \snm{Stadler}},
-\author[3]{\fnms{Jonathan} \snm{Pool}}
-\address[1]{University of Leipzig, patrick.westphal@studserv.uni-leipzig.de}
-\address[2]{University of Leipzig, cstadler@informatik.uni-leipzig.de}
-\address[3]{Long Now Foundation, San Francisco, pool@utilika.org}
+\author[A]{\fnms{Patrick} \snm{Westphal}},
+\author[A]{\fnms{Claus} \snm{Stadler}},
+\author[B]{\fnms{Jonathan} \snm{Pool}}
+%\address[1]{University of Leipzig, patrick.westphal@studserv.uni-leipzig.de}
+\address[A]{University of Leipzig, \{pwestphal, cstadler\}@informatik.uni-leipzig.de}
+\address[B]{Long Now Foundation, San Francisco, pool@panlex.org}
 
@@
 to the world losing this linguistic diversity.
-The vision of the PanLex project is to save these languages, especially low-density ones,
-by allowing them to be intertranslatable and thus to be a part of the Information Age.
+The vision of the PanLex project is to help save these languages, especially
+low-density ones, by allowing them to be intertranslatable and thus to be a part of the Information Age.
 For this reason, PanLex gathers and integrates information from thousands of
@@
 glossaries, standards and thesauri.
-In this paper we describe how we transformed this data to RDF, interlinked it with
-Lexvo and DBpedia and published it as Linked Data and SPARQL.
+In this dataset description paper we detail how we transformed this data to
+RDF, interlinked it with Lexvo and DBpedia and published it as Linked Data and
+via SPARQL.
 \end{abstract}
@@
 \section{Introduction}
+\label{sec:intro}
 % \begin{itemize}
@@
 and revitalization are essential to the preservation of biological diversity~\cite{nettle}.
-The vision of the PanLex project is to save all these thousands of languages,
-especially those low-density ones that are threatened by extinction,
+The vision of the PanLex project is to help save these thousands of
+languages, especially those low-density ones that are threatened by extinction,
 by supporting their use in global communication. This requires panlingual
@@
 
-In \autoref{sec:triplificate} we will introduce the PanLex dataset, present our
+In \autoref{sec:triplificate} we introduce the PanLex dataset, present our
 PanLex RDF vocabulary and explain how we transformed the one into the other.
-The following section is about the actual dataset publishing and how we linked
-to other datasets of the LLOD cloud.
-Finally we discuss related work, conclude our approach and give some hints to future work.
+\autoref{sec:linking} is about how we linked to other datasets of the LLOD
+cloud, whereas \autoref{sec:publishing} is about the actual dataset publishing.
+Usage scenarios are given in \autoref{sec:usage}.
+In \autoref{sec:related} we discuss related work, and finally, in
+\autoref{sec:conclusion} we conclude our approach and give some hints to future
+work.
 
@@
 \label{sec:triplificate}
-In this section we first provide an analysis of the original PanLex data.
+In this section we first provide an analysis of the original PanLex dataset.
 Subsequently, we introduce our URI and vocabulary design and we explain 
@@
 \label{sec:analysis}
-At the core, PanLex is a \emph{database}, served by a PostgreSQL database server, that ``represents \emph{assertions} about the \emph{meanings} of \emph{expressions}''\footnote{\url{http://panlex.org/tech/doc/design/panlex-db-design.pdf}}.
+At the core of the PanLex \emph{project}, there is the PanLex
+\emph{database} which is created from the imports of thousands of lexical
+resources, such as monolingual dictionaries, bilingual dictionaries, multilingual dictionaries, glossaries, standards, and thesauri.
+The concrete list of used sources is available online\footnote{\url{http://panlex.org/tech/plrefs.shtml}}.
+The data derived from these sources comprises single- and multi-word expressions and meanings assigned to them.
+Conceptually, the PanLex database thus ``represents \emph{assertions} about the \emph{meanings} of \emph{expressions}''\footnote{\url{http://panlex.org/tech/doc/design/panlex-db-design.pdf}}.
 As of now, the database contains about 20 million meanings and 19 million expressions extracted from about 2,000 sources.
+The most important entities and their relations are depicted in \autoref{fig:db-schema} and are explained in more the detail in the following.
 
-Assertions are factual claims attributed to a certain source and editor.
-They do not represent absolute truths and as such it is valid that different editors make contradictory assertions based on the same sources.
+%At the core, PanLex is a \emph{database}, served by a PostgreSQL database
+%server, that ``represents \emph{assertions} about the \emph{meanings} of
+% \emph{expressions}''\footnote{\url{http://panlex.org/tech/doc/design/panlex-db-design.pdf}}.
+%As of now, the database contains about 20 million meanings and 19 million
+% expressions extracted from about 2,000 sources.
 
-Expressions are lexemes (dictionary entries) and are related to their respective languages.
-Furthermore, expressions have meanings. All expressions sharing the same meaning are either translations or synonyms.
+%Assertions are factual claims attributed to a certain source and editor.
+%They do not represent absolute truths and as such it is valid that different
+% editors make contradictory assertions based on the same sources.
 
-The lexical database of the PanLex project is based on different sources like monolingual dictionaries,
-bilingual dictionaries, multilingual dictionaries, glossaries, standards, and thesauri.
-The list of used sources can also be found on the Web\footnote{\url{http://panlex.org/tech/plrefs.shtml}}.
+%Expressions are lexemes (dictionary entries) and are related to their
+% respective languages.
+%Furthermore, expressions have meanings. All expressions sharing the same
+% meaning are either translations or synonyms.
 
-The actual data derived from these sources comprises single- and multi-word expressions and meanings assigned to them.
-The most important entities and their relations are depicted in \autoref{fig:db-schema} and are explained in more the detail in the following.
+%The lexical database of the PanLex project is based on different sources like
+% monolingual dictionaries, bilingual dictionaries, multilingual dictionaries, glossaries, standards, and thesauri.
+%The list of used sources can also be found on the
+% Web\footnote{\url{http://panlex.org/tech/plrefs.shtml}}.
+
+%The actual data derived from these sources comprises single- and multi-word
+% expressions and meanings assigned to them.
+%The most important entities and their relations are depicted in
+% \autoref{fig:db-schema} and are explained in more the detail in the following.
 
@@
 \begin{itemize}
-  \item \emph{Expressions} are lemmas. For example, \emph{``go''} is a valid expression, whereas \emph{``went''} is not.
+  %\item A meaning's \emph{approver} is a pair comprised of the editor and the considered information source.
+  \item The starting point of the data acquisition is the \emph{approver}
+  entity: An editor processes the content of a certain
+  mono- or multilingual source as mentioned above and adds it to the PanLex dataset.
+  This combination of a user and a source is referred to as an approver.
+
+  \item \emph{Expressions} are lemmas, i.e. dictionary entries. For example,
+  \emph{``go''} is a valid expression, whereas \emph{``went''} is not.
   Expressions are always given in a language variety and can only be given once per language variety.
-  They were extracted from the afore-mentioned mono- or bilingual dictionaries or from other sources.
+  %They were extracted from the afore-mentioned mono- or bilingual dictionaries
+  % or from other sources.
   \item \emph{Languages} in PanLex are identified using ISO 639-3\footnote{\url{http://www.sil.org/iso639-3/codes.asp}} individual and macrolanguage codes, ISO 639-2\footnote{\url{http://www.loc.gov/standards/iso639-2/}} collective codes and ISO 639-5\footnote{\url{http://www.loc.gov/standards/iso639-5/}} codes.
-  \item \emph {Language varieties} within a language are further distinguished by arbitrary unique integer identifiers.
-  For example, ``eng'' is the ISO 639-3 code for the English language.
+  \item \emph {Language varieties} allow one to make more fine grained
+  distinctions within a language. Their codes are composed of the language code
+  combined with a PanLex specific identifier.
+  For example, ``eng''
+  is the ISO 639-3 code for the English language.
   Panlex defines various varieties of it, including ``English'' (eng-000), ``Simple English'' (eng-001) and ``British English'' (eng-005).
@@
   So, in contrast to mechanisms like IETF BCP 47\footnote{\url{http://tools.ietf.org/html/bcp47}} there is no need for a transcription. %  and one can use variety names like ``{زبان اوستایی}''.
-  Additionally these variety identifiers and labels do not have to be registered like the IETF BCP 47 language subtags\footnote{\url{http://www.iana.org/assignments/language-subtag-registry}}.
-  \item \emph{Meanings} of expressions are distinguished by editors based on their
-  interpretation of an information source.
+  %Additionally these variety identifiers and labels do not have to be registered like the IETF BCP 47 language subtags\footnote{\url{http://www.iana.org/assignments/language-subtag-registry}}.
+  \item \emph{Meanings} in PanLex are entities of which each represents a unique
+  possible sense of an expression. 
+  %They are allocated as necessary by an editor based on his or her
+ %interpretation of expressions.
+  Meanings are assigned by editors based on their interpretation of expressions.
+  Usually this assignment is done on a per-source basis so that identical meanings
+  across multiple sources are not resolved.
+  %different
+  % meaning identifiers may refer to the same intended sense.
+  This means that if there is e.g. a translation of the fruit \emph{``apple''}
+  in an English to German dictionary and another translation from English to
+  French, these do not necessarily result in a single meaning entity linking to
+  all involved languages.
+  %the to corresponding expressions in English, German and French.
+  Instead, there could be \emph{``apple''} and \emph{``Apfel''} sharing one
+  meaning entity and \emph{``apple''} and \emph{``pomme''} sharing another.
+%  Editors may add informal definitions to meanings in any language variety.
 %  Information sources can be dictionaries, thesauri, etc., and also personal knowledge. 
-  The pair comprised of the editor and the information source is referred to as a meaning's \emph{approver}.
-  Editors may add informal definitions to meanings in any language variety.
+%  The pair comprised of the editor and the information source is referred to as a meaning's \emph{approver}.
   \item \emph{Definitions} are optional descriptions of a meaning.
@@
   A description of the verb \emph{browse} for example could be \emph{``move or surf through various files on a computer, the Internet, etc.''}, marked as a definition in the ``English'' language variety.
-  \item \emph{Denotations} relate expressions to meanings.
-  So homonyms are represented by one expression being connected to different meanings.
-  Denotations may be enriched with part-of-speech tags selected from a closed list based on the
-  Open Lexicon Interchange Format (OLIF) standard\footnote{\url{http://www.olif.net/}}.
+  \item \emph{Denotations} are entities that relate expressions to meanings and
+  may optionally carry annotations in form of sets of key value pairs. For
+  instance, an English expression \emph{pig}, when referring to
+  \emph{police officer}, could be annotated with \emph{pragmatics=vulgar}. 
+  Furthermore, denotations can be tagged with part-of-speech tags, such as word
+  classes, selected from a closed list based on the Open Lexicon Interchange Format (OLIF) standard\footnote{\url{http://www.olif.net/}}.
   For example, \emph{fall} can be a verb or a noun for autumn. 
-  Furthermore, it is possible to generically tag denotations with metadata
-  using key-value pairs.
-  For instance, an English expression \emph{pig} when synonymous with
-  \emph{police officer} could be annotated with \emph{pragmatics=vulgar}.
+  Homonyms are those expressions that are connected to multiple meanings.
   \item \emph{Users} have editorial privileges over the language varieties and the approvers that they define.
@@
 Note that \emph{approver} combines a user with an \emph{information source},
-however the \emph{information source} is not modelled as a distinct entity.
+however the \emph{information source} is not modeled as a distinct entity.
 Also, the information of whether or not two meanings with different approvers
@@
 In general, all PanLex RDF resources reside
-in the namespace \texttt{\small <http://panlex.org/plx/>}, abbreviated with \texttt{\small plx}.
+in the namespace \texttt{\small <http://ld.panlex.org/plx/>}, abbreviated with
+\texttt{\small plx}.
 An example of the resulting ontology is depicted in \autoref{fig:vocabulary} and
@@
 \begin{itemize}
-\item Expressions are modeled as intances of the class \texttt{\small plx:Expression}.
+\item Expressions are modeled as instances of the class \texttt{\small plx:Expression}.
 Their original and normalized textual representations become the 
@@
 \emph{ISO 639-1} and \emph{ISO 639-3 codes} become instances of the classes
-\texttt{\small plx:Iso639-1Code} and \texttt{\small plx:Iso639-3Code}. % manual linebreak
+\texttt{\small plx:Iso639-1Code} and \texttt{\small plx:Iso639-3Code}.
 
@@
 \item All approvers share the \texttt{\small plx:Approver} class. The characteristics of an approver are described using mainly triples with literal objects.
-These are for example \texttt{\small plx:registrationDate} pointing to an \texttt{\small xsd:date} literal,
+These are for example %\texttt{\small plx:registrationDate} pointing to an \texttt{\small xsd:date} literal,
 \texttt{\small dc:title} to assign the title of a source,
@@
 \begin{table}
-\begin{scriptsize}
+\begin{tiny}
 \begin{tabular}{ll}
@@
 \toprule
-\texttt{plx:Approver} & \texttt{plx:registrationDate} \\
-                      & \texttt{rdfs:label} \\
-                      & \texttt{foaf:homepage} \\
-                      & \texttt{plx:isbn} \\
-                      & \texttt{dc:creator} \\
-                      & \texttt{dc:title} \\
-                      & \texttt{dc:publisher} \\
-                      & \texttt{plx:yearOfPublication} \\
-                      & \texttt{plx:quality} \\
-                      & \texttt{plx:license} \\
+\texttt{plx:Approver} & \texttt{plx:registrationDate}, \texttt{rdfs:label}\\
+                      & \texttt{foaf:homepage}, \texttt{plx:license}, \texttt{dc:date}\\
+                      & \texttt{dc:creator}, \texttt{dc:title}, \texttt{plx:quality} \\
+                      & \texttt{dc:publisher}, \texttt{dbpedia-owl:isbn} \\
 \midrule
@@
 \texttt{plx:LanguageVariety}
-                      & \texttt{plx:languageVarietyOf} \\
-                      & \texttt{rdfs:label} \\
+                      & \texttt{plx:languageVarietyOf}, \texttt{rdfs:label} \\
 \midrule
@@
                       & \texttt{plx:languageVariety} \\
-                      & \texttt{rdfs:label} \\
-                      & \texttt{plx:degradedText} \\
+                      & \texttt{plx:degradedText}, \texttt{rdfs:label} \\
 \midrule
 \texttt{plx:Meaning}
-                      & \texttt{plx:approver} \\
+                      & \texttt{plx:approver}, \texttt{plx:identifier} \\
                       & \texttt{plx:meaningDefinition} \\
-                      & \texttt{plx:identifier} \\
 \midrule
 \texttt{plx:Definition}
-                      & \texttt{plx:languageVariety} \\
-                      & \texttt{rdfs:label} \\
+                      & \texttt{plx:languageVariety}, \texttt{rdfs:label} \\
 \midrule
@@
 \end{tabular}
-\end{scriptsize}
-\caption{Classes and properties used in the PanLex RDF vocabulary. Note that all \texttt{rdf:type} and \texttt{rdfs:label} properties have been omitted for brevity.}
+\end{tiny}
+\caption{Classes and properties used in the PanLex RDF vocabulary. Note that all \texttt{rdf:type} properties are omitted for brevity.}
 \label{tbl:vocabulary}
@@
 \begin{figure}
-\includegraphics[width=0.47\textwidth]{images/pdf/ontology.pdf}
+\includegraphics[width=0.46\textwidth]{images/pdf/ontology.pdf}
 \caption{Overview of the PanLex RDF vocabulary}
@@
 to convert all the data. This makes
-testing and debugging and fixing issues in the modelling or data conversion of the
+testing and debugging and fixing issues in the modeling or data conversion of the
 data very time-consuming.
 As the PanLex data already resides in a relational database, the use of a virtual
-RDB 2 RDF mapping solution is a natural choice.
+RDB2RDF\footnote{\url{http://www.w3.org/2001/sw/wiki/RDB2RDF}}  mapping solution is a natural choice.
 The \emph{Sparqlify system}\footnote{\url{https://github.com/AKSW/Sparqlify}} offers,
 besides an efficient query rewriting engine, also a very easy-to-use mapping
-language (ML).
+language, called \emph{Sparqlification Mapping Language} (SML).
 Essentially, these mappings consist of three clauses:
 \begin{itemize}
-  \item \emph{From:} Specifies the logical SQL table (i.e. table, view or query) underlying the Sparqlify-ML view.
+  \item \emph{From:} Specifies the logical SQL table (i.e. table, view or query) underlying the SML view.
   \item \emph{With:} Defines SPARQL variables by means of expressions over relational columns that yield RDF terms.
@@
 \end{itemize}
-~\autoref{fig:ex:sparqlify-ml} shows an example of a Sparqlify-ML view for the
+\autoref{fig:ex:sparqlify-ml} shows an example of an SML view for the
 languages in PanLex: From each row of the \emph{i1} table three resources are
@@
 Resources for \emph{?lang} become typed as a \emph{Language} in the PanLex and the \emph{schema.org} namespace.
-This view-based approach also demonstrates that changing the vocubulary or
-adding support for new ones does not require an ETL process, and can therefore be done with little effort.
+This view-based approach also demonstrates that changing the vocabulary or
+adding support for new ones does not require an extract transform load (ETL) process, and can therefore be done with little effort.
 %\todo{Maybe we should link to lexvo directly.}
@@
 \centering
-\begin{lstlisting}[basicstyle=\footnotesize\ttfamily]
+\begin{lstlisting}
 Create View i1 As
@@
   With
-    ?lang = uri(concat(plx:language, '/', ?iso3))
-    ?iso3 = uri(concat(plx:iso639-3, '/', ?iso3))
-    ?lexvo3 = uri(concat('http://lexvo.org/id/iso639-3/', ?iso3))
+    ?lang = uri(plx:language, '/', ?iso3)
+    ?iso3 = uri(plx:iso639-3, '/', ?iso3)
+    ?lexvo3 = uri('http://lexvo.org/id/iso639-3/', ?iso3)
   From
@@
 \end{lstlisting}
-\caption{An excerpt of a Sparqlify-ML view definition for PanLex's languages. This
-example also demonstrates how ``is-a relations'' to schema.org and links to Lexvo are established.}
+\caption{An excerpt of an SML view definition for PanLex's languages. This
+example also demonstrates how ``is-a'' relations to schema.org and links to Lexvo are established.}
 \label{fig:ex:sparqlify-ml}
@@
 \label{sec:linking}
-The Sparqlify view in the previous section (\autoref{fig:ex:sparqlify-ml}) already
+The SML view in the previous section (\autoref{fig:ex:sparqlify-ml}) already
 established the interlinking of the PanLex languages with Lexvo.
@@
 thus \emph{dereferenceable} links.
-Therefore, we iterated the \emph{titles} datasets\footnote{\url{http://wiki.dbpedia.org/Downloads38} - Note that more languages are available in the raw download section at \url{http://downloads.dbpedia.org/3.8/}},
+Therefore, we iterated the \emph{titles} datasets\footnote{\url{http://wiki.dbpedia.org/Downloads38}% - Note that more languages are available in the raw download section at \url{http://downloads.dbpedia.org/3.8/}},
+},
 which map (non-localized) DBpedia URIs to their page titles in the respective language.
@@
 \autoref{fig:plx-dbp-link-counts} summarizes the number of links obtained.
+
+In total, about 2.5 million links were obtained for approx. 20 million
+expressions. This relatively low coverage can be attributed to
+frequently appearing multi-word expressions that do not match the DBpedia titles well, and
+the fact that in this work we yet only considered DBpedia datasets for
+mainstream languages, whereas PanLex focuses on low-density ones.
+
+%The relatively low coverage of these links (about 2,500,000 links for about 20,000,000 expressions) can be attributed to the frequently appearing multi-word expressions that do not match well to titles and the fact that the considered DBpedia datasets are mainly given for mainstream languages but not for the low-density ones in focus of PanLex.
 \begin{table}
@@
 \label{sec:publishing}
-PanLex itself already provides direct access to the PostgreSQL database for scientists,
-several user interfaces\footnote{\url{http://panlex.org/try/}} and an
-API\footnote{\url{http://panlex.org/tech/doc/api/}}.
+% PanLex itself already provides direct access to the PostgreSQL database for scientists,
+% several user interfaces\footnote{\url{http://panlex.org/try/}} and an
+% API\footnote{\url{http://panlex.org/tech/doc/api/}}.
+As stated in~\autoref{sec:intro}, the PanLex project already provides several
+interfaces for data access.
 %The usage of these interfaces is not restricted in order to make the provided lexical
@@
 With our RDF conversion work, we complement these interfaces with
-the Linked Data
+Linked Data
 %\footnote{\url{http://panlex.org/plx/TODO}}
-and the SPARQL
-endpoint\footnote{\url{http://panlex.org/sparql} and \url{http://panlex.org/snorql}}.
+and an SPARQL
+endpoint\footnote{\url{http://ld.panlex.org/sparql} and \url{http://ld.panlex.org/snorql}}.
 An overview is shown in ~\autoref{fig:panlex-architecture}.
-Our Sparqlify-ML views and the code for interlinking are
+Our SML views and the code for interlinking are
 hosted on GitHub\footnote{\url{https://github.com/AKSW/PanLex-2-RDF}}.
@@
 with the other data using Sparqlify.
-
+Downloads and additional information are available on the Panlex-RDF
+project page\footnote{\url{http://ld.panlex.org}}.
 \begin{figure}
@@
 There are general benefits of RDF conversions,
-such as the paradim shift towards thinking about what one intends to express
+such as the paradigm shift towards thinking about what one intends to express
 with the data, making this meaning explicit using ontologies,
@@
 new interesting mashups.
-For example, the TeraDict translation service\footnote{\url{http://panlex.org/cgi-bin/apidemo-2.cgi}}
+For example, the TeraDict translation service\footnote{\url{http://panlex.org/teradict/?lg=eng}}
 could now be easily realized using simple SPARQL queries.
@@
 Another future usage scenario is to link PanLex to Wortschatz\footnote{\url{http://wortschatz.uni-leipzig.de/}} with the aim of 
-improving the recall of word-by-word translations by considering the co-occurences.
+improving the recall of word-by-word translations by considering their co-occurrences.
 
@@
 %as Linked Data\footnote{\url{http://www.w3.org/DesignIssues/LinkedData.html}} and via SPARQL{\footnote{\url{http://www.w3.org/TR/sparql11-query/}}}.
-as Linked Data\footnote{\url{http://www.w3.org/DesignIssues/LinkedData.html}} and via SPARQL.
+as Linked Data %\footnote{\url{http://www.w3.org/DesignIssues/LinkedData.html}}
+and via SPARQL.
 These efforts are referred to as the \emph{Linguistic Linked Open Data Cloud} (LLOD). 
@@
 Examples include the \emph{Ontologies of Linguistic Annotation} (OLiA)~\cite{olia2010},
-the \emph{Lexicon Model for Ontologies} (lemon)~\cite{lemon2011} for modeling lexicon and machine-readable dictionaries,
+the \emph{Lexicon Model for Ontologies} (lemon)~\cite{lemon2011} as well as \emph{General Ontology for Linguistic Description} (GOLD)~\cite{farr2003} for modeling lexicon and machine-readable dictionaries,
 \emph{POWLA} for modeling linguistic corpora\cite{powla2012}
@@
 \label{sec:conclusion}
-In this paper we describe the PanLex database and its conversion to RDF.
+In this dataset description we detailed the PanLex database and its conversion
+to RDF.
 Based on our URI and vocabulary design, we created appropriate view definitions
-for the virtual RDB to RDF solution \emph{Sparqlify} which carries out
+for the virtual RDB2RDF solution \emph{Sparqlify} which carries out
 the actual RDF transformation.
@@
 and created about 2.5 million links to DBpedia for expressions in 16 languages.
+Although not discussed explicitly due to space limitations, we
+also made first steps towards enhancing the RDB2RDF view definitions
+as to facilitate the integration of the PanLex data with the lemon and GOLD
+data models.
+
+In conclusion, we see this work as a valuable contribution to the vision of
+countering language attrition by making a vast amount of lexical
+resources available as Linked Open Data.
+
+% We see this work of making a vast amount of lexical
+% resources available as Linked Open Data as a valuable
+% contribution to the vision of countering language attrition.
+
 %We detailed the developed vocabulary and
 
-In this work we also identified some weaknesses in
-the orginal database design,
-which we intend to overcome in the future:
-The orginal dataset currently does not cleanly model information sources and approvers
+We identified some weaknesses in
+the original database design, which we intend to overcome in the future:
+The original dataset currently does not cleanly model information sources and approvers
 as distinct entities.
 Also, while it is possible to determine each meaning's and each denotation's approver, an approver can have multiple users with editorial privileges, so it is not possible to track each contributed, modified, and deleted meaning and denotation by editor.
-Retaining this information seems beneficial for translation approches,
-as this enables for example the attribution of qualities and relevances to particular editors.
+Retaining this information seems beneficial for translation approaches,
+as this enables for example the attribution of qualities and relevances to
+specific editors.
+
+%We made first steps towards the support of additional linguistic
+%ontology models: There
+%So apart from the PanLex data model, expression-meaning relations can also be
+%retrieved using the lemon model.
+%Another approach to reuse existing vocabularies is the integration of the
+% linguistic sign model of the GOLD standard.
 

### 2014-01-22 panlex-rdf-paper/2012/MLOD_PanLex/paper_v3.tex
Added the panlex paper to the repo

--- 2014-01-22 panlex-rdf-paper/2012/MLOD_PanLex/paper.tex
@@
 \usepackage[breaklinks]{hyperref}
+
 \def\sectionautorefname{Section}
 \def\subsectionautorefname{Section}
+
 \usepackage{multirow}
@@
 \usepackage{array}
+\usepackage{textcomp}
 
-%\usepackage{authblk}
-
-\usepackage{textcomp}
 \newcolumntype{R}{>{\raggedleft\arraybackslash}X}
@@
 \newcommand\Small{\fontsize{9}{9.2}\selectfont}
-%\newcommand*\LSTfont{\Small\ttfamily\SetTracking{encoding=*}{-60}\lsstyle}
 
@@
 }%
-
-%\usepackage{verbatim}
 \lstset{%
@@
 \begin{frontmatter} 
-
-%\pretitle{Pretitle}
-
 \title{Countering language attrition with PanLex and the Web of Data}
 \runningtitle{Countering language attrition with PanLex and the Web of Data}
-%\todo{find a real title}
 
@@
 
-%patrick.westphal@studserv.uni-leipzig.de
 \author[A]{\fnms{Patrick} \snm{Westphal}},
@@
 \author[B]{\fnms{Jonathan} \snm{Pool}}
-%\address[1]{University of Leipzig, patrick.westphal@studserv.uni-leipzig.de}
 \address[A]{University of Leipzig, \{pwestphal, cstadler\}@informatik.uni-leipzig.de}
@@
 
-%\author[1]{Author A\thanks{A.A@university.edu}}
-%\author[1]{Author B\thanks{B.B@university.edu}}
-%\author[1]{Author C\thanks{C.C@university.edu}}
-%\author[2]{Author D\thanks{D.D@university.edu}}
-%\author[2]{Author E\thanks{E.E@university.edu}}
-%\affil[1]{Department of Computer Science, \LaTeX\ University}
-%\affil[2]{Department of Mechanical Engineering, \LaTeX\ University}
-
-%\address[1]{University of Leipzig, Institute of Computer Science, AKSW, patrick.westphal@studserv.uni-leipzig.de}
-
-%\address[1]{University of Leipzig, Institute of Computer Science, AKSW, Postfach 100920, D-04009 Leipzig, Germany}
-
-%\author[A]{Patrick Westphal\thanks{A.A@university.edu}},
-%\author[A]{Claus Stadler\thanks{cstadler@informatik.uni-leipzig.de}},
-%\author[B]{Jonathan Pool\thanks{todo}}
-%\address[A]{University of Leipzig, Institute of Computer Science, AKSW, Postfach 100920, D-04009 Leipzig, Germany\\ E-mail: \{lastname\}@informatik.uni-leipzig.de}
-%\address[A]{University of Leipzig, Institute of Computer Science, AKSW, Postfach 100920, D-04009 Leipzig, Germany}
-%\address[B]{TODO}
-
-%\author[A]{\fnms{Patrick} \snm{Westphal}}
-%\author[B]{\fnms{Claus} \snm{Stadler}}
-
-%\runningauthor{M. Martin et al.}
-
-%\address[B]{University of Leipzig}
-%\address[C]{TODO}
-%\todo{the e-mail address pattern won't work for both of us}
-
 \begin{abstract}
 At present, there are approximately 7,000 living languages in the world.
-However, some experts claim that the process of globalization may eventually lead
-to the world losing this linguistic diversity.
-The vision of the PanLex project is to help save these languages, especially
-low-density ones, by allowing them to be intertranslatable and thus to be a part of the Information Age.
-For this reason, PanLex gathers and integrates information from thousands of
-linguistic resources, such as monolingual dictionaries, bilingual dictionaries, multilingual dictionaries,
-glossaries, standards and thesauri.
-In this dataset description paper we detail how we transformed this data to
-RDF, interlinked it with Lexvo and DBpedia and published it as Linked Data and
-via SPARQL.
+However, some experts claim that the process of globalization may eventually lead to the world losing this linguistic diversity.
+The vision of the PanLex project is to help save these languages, especially low-density ones, by allowing them to be intertranslatable and thus to be a part of the Information Age.
+Semantic Web technologies can support achieving this goal, for reasons such as
+their capabilities of flexibly representing, interlinking and reasoning with 
+data, in our case particularly linguistic resources and annotations.
+%Using Semantic Web technologies can support achieving this goal, especially due
+% to inference capabilities and the interlinking of the PanLex data with other data sources.
+Conversely, an RDF version of PanLex makes a significant contribution towards
+improving the coverage of the Linguistic Web of Data, as to the best of our
+knowledge there exists no large scale Linked Data data set for panlingual translation of
+non-mainstream languages.
+In this dataset description paper we detail how we transformed the data of the PanLex project to RDF,
+established conformance with the lemon and GOLD data models,
+interlinked it with Lexvo and DBpedia, and published it as Linked Data and via
+SPARQL.
 \end{abstract}
@@
 \label{sec:intro}
-% \begin{itemize}
-% \item introduce PanLex project
-% \item @author: Jonathan
-% \item Motivation: About approximately 6900 active languages (\todo{source seems to be http://www.amazon.com/Ethnologue-Languages-M-Paul-Lewis/dp/1556712162}), Existing translation only between relatively few language pairs (dictionaries).
-% \item Transitive word-by-word translations 
-% * high density and low density languages
-% * decline in linguistic diversity
-% * These combined forces have led to predictions that something between half and 90%
-%   of the world's living languages will die within the next century (Woodbury 2006; UNESCO 2003: 2).
-%   Weaker forms of these forces appear to be shrinking the use of medium-density languages in science,
-%   diplomacy, business, and other domains (Phillipson 2008).
-% * When languages die, they argue, the world loses
-%   (1) irreplaceable knowledge of history, medicine, nature, and productive methods encoded in languages'
-% * extraction + integration
-% \end{itemize}
 At present, there are about 7,000 living languages in the world\footnote{See \url{http://www.sil.org/iso639-3/download.asp} for a list of registered languages}.
 Nonetheless, some experts claim that processes such as nation-state consolidation and globalization are producing language attrition so rapidly that up to 90\% of all languages alive today will be extinct within a century~\cite{endang_lang}.
-Theorists of biolinguistic diversity argue that the loss of
-language diversity, the loss of human biological knowledge, and the loss of
-species diversity are mutually supportive and thus that language preservation
-and revitalization are essential to the preservation of biological diversity~\cite{nettle}.
-The vision of the PanLex project is to help save these thousands of
-languages, especially those low-density ones that are threatened by extinction,
-by supporting their use in global communication. This requires panlingual
-translation: translation from any language of the world into any other.
-One of the crucial components of panlingual translation of discourses is
-panlingual lexical translation. PanLex is designed to support that component.
-It documents the known lexical translations (translations of lexemes) among all
-languages.
+Theorists of biolinguistic diversity argue that the loss of language diversity, the loss of human biological knowledge, and the loss of species diversity are mutually supportive and thus that language preservation and revitalization are essential to the preservation of biological diversity~\cite{nettle}.
+The vision of the PanLex project is to help save these thousands of languages, especially those low-density ones that are threatened by extinction, by supporting their use in global communication.
+This requires panlingual translation: translation from any language of the world into any other.
+One of the crucial components of panlingual translation of discourses is panlingual lexical translation.
+PanLex is designed to support that component by making use of several thousand
+sources and artificial intelligence approaches.
+%It documents the known lexical translations (translations of lexemes) among all
+%languages.
 
-Although a list of all the translations of all words into all other languages
-would be large (amounting to trillions of translations), there are more
-serious obstacles. The vast majority of these translations are unknown.
-Those that are known are not universally agreed on. And lexical polysemy
-and ambiguity make the question ``What is the translation of lexeme X into language Y?''
-underspecified for practical purposes.
-So the PanLex project is designed to permit the application of artificial
-intelligence techniques to infer and select translations.
+Apart from using PanLex's data for inferring lexical translations, a
+further direction of work is to generate added value by enriching this data with
+knowledge from other (linguistic) sources.
+Grounding on the idea of a Semantic Web, a complete stack of technologies has
+been devised and standardized by the
+\emph{World Wide Web Consortium}\footnote{\url{http://www.w3.org/standards/semanticweb/}}
+supporting the definition of a machine readable and interpretable Linked Data
+network, which has led to emergence of the \emph{Web of Data}.
+%Technologies that support this aim are already available and in use in the so
+%called \emph{Web of Data}.
+Currently there is a growing community working on leveraging these Semantic Web
+technologies for linguistic knowledge and thereby building a Web of Linguistic
+Data, also known as the \emph{Linguistic Linked Open Data (LLOD) cloud}.
 
-Currently there is a growing community trying to combine linguistic knowledge
-with Semantic Web technologies and thereby build a Web of Linguistic Data,
-also known as the \emph{Linguistic Linked Open Data (LLOD) cloud}.
-The PanLex project is still in its data-acquisition phase, but it has provided a
-few web services\footnote{\url{http://panlex.org/try/}} and
-APIs\footnote{\url{http://panlex.org/tech/doc/api/}} to access its data.
-Since the project shares the idea of open access, we have undertaken the project
-of making PanLex a part of the LLOD cloud.
-
+Our steps to connect PanLex to this Linked Data network are as follows.
 In \autoref{sec:triplificate} we introduce the PanLex dataset, present our
-PanLex RDF vocabulary and explain how we transformed the one into the other.
-\autoref{sec:linking} is about how we linked to other datasets of the LLOD
-cloud, whereas \autoref{sec:publishing} is about the actual dataset publishing.
+PanLex RDF vocabulary, explain how we transformed the one into the other, and
+how we established conformance with additional data models.
+\autoref{sec:linking} is about how we linked to other datasets of the LLOD cloud,
+whereas \autoref{sec:publishing} is about the actual dataset publishing.
 Usage scenarios are given in \autoref{sec:usage}.
-In \autoref{sec:related} we discuss related work, and finally, in
-\autoref{sec:conclusion} we conclude our approach and give some hints to future
-work.
+In \autoref{sec:related} we discuss related work, and finally, in \autoref{sec:conclusion} we conclude our approach and give some hints to future work.
 
@@
 \label{sec:triplificate}
-In this section we first provide an analysis of the original PanLex dataset.
-Subsequently, we introduce our URI and vocabulary design and we explain 
-the steps taken to publish the data as RDF.
-%The vocabulary 
-%afterwards we show how the original data was converted to RDF and published
-%via a SPARQL endpoint.
-%The interlinking of the created RDF data with other datasets is discussed in the
-%following section.
+In this section, we first provide an analysis of the PanLex dataset.
+Subsequently, we introduce our URI and vocabulary design which closely
+resembles PanLex's original conceptual model.
+%, which was developed independently from other ontology engineering approaches.
+Afterwards, we briefly describe how we classified PanLex's
+instance data using additional data models. Finally, we explain the steps
+taken to transform the data to RDF.
+
+%As PanLex defines its own  conceptual model
 
@@
 \label{sec:analysis}
-At the core of the PanLex \emph{project}, there is the PanLex
-\emph{database} which is created from the imports of thousands of lexical
-resources, such as monolingual dictionaries, bilingual dictionaries, multilingual dictionaries, glossaries, standards, and thesauri.
+At the core of the PanLex \emph{project}, there is the PanLex \emph{database} which is created from the imports of thousands of lexical resources, such as mono- and multilingual dictionaries, glossaries, standards, and thesauri.
 The concrete list of used sources is available online\footnote{\url{http://panlex.org/tech/plrefs.shtml}}.
-The data derived from these sources comprises single- and multi-word expressions and meanings assigned to them.
+The data derived from these sources comprise single- and multi-word expressions and meanings assigned to them.
 Conceptually, the PanLex database thus ``represents \emph{assertions} about the \emph{meanings} of \emph{expressions}''\footnote{\url{http://panlex.org/tech/doc/design/panlex-db-design.pdf}}.
 As of now, the database contains about 20 million meanings and 19 million expressions extracted from about 2,000 sources.
-The most important entities and their relations are depicted in \autoref{fig:db-schema} and are explained in more the detail in the following.
+The most important entities and relations of PanLex's conceptual model are
+depicted in \autoref{fig:db-schema} and are explained in more the detail in the following.
 
-%At the core, PanLex is a \emph{database}, served by a PostgreSQL database
-%server, that ``represents \emph{assertions} about the \emph{meanings} of
-% \emph{expressions}''\footnote{\url{http://panlex.org/tech/doc/design/panlex-db-design.pdf}}.
-%As of now, the database contains about 20 million meanings and 19 million
-% expressions extracted from about 2,000 sources.
-
-%Assertions are factual claims attributed to a certain source and editor.
-%They do not represent absolute truths and as such it is valid that different
-% editors make contradictory assertions based on the same sources.
-
-%Expressions are lexemes (dictionary entries) and are related to their
-% respective languages.
-%Furthermore, expressions have meanings. All expressions sharing the same
-% meaning are either translations or synonyms.
-
-%The lexical database of the PanLex project is based on different sources like
-% monolingual dictionaries, bilingual dictionaries, multilingual dictionaries, glossaries, standards, and thesauri.
-%The list of used sources can also be found on the
-% Web\footnote{\url{http://panlex.org/tech/plrefs.shtml}}.
-
-%The actual data derived from these sources comprises single- and multi-word
-% expressions and meanings assigned to them.
-%The most important entities and their relations are depicted in
-% \autoref{fig:db-schema} and are explained in more the detail in the following.
-
-%This denotation process is also reflected in the database design shown below.
-%To carry out the goal of a multi-lingual information source a suitable database design is needed.
 \begin{figure}
@@
 \begin{itemize}
-  %\item A meaning's \emph{approver} is a pair comprised of the editor and the considered information source.
-  \item The starting point of the data acquisition is the \emph{approver}
-  entity: An editor processes the content of a certain
-  mono- or multilingual source as mentioned above and adds it to the PanLex dataset.
+  \item The starting point of the data acquisition is the \emph{approver} entity: An editor processes the content of a certain mono- or multilingual source as mentioned above and adds it to the PanLex dataset.
     This combination of a user and a source is referred to as an approver.
-
-  \item \emph{Expressions} are lemmas, i.e. dictionary entries. For example,
-  \emph{``go''} is a valid expression, whereas \emph{``went''} is not.
+  \item \emph{Expressions} are lemmas, i.e. dictionary entries.
+    For example, \emph{``go''} is a valid expression, whereas \emph{``went''} is not.
     Expressions are always given in a language variety and can only be given once per language variety.
-  %They were extracted from the afore-mentioned mono- or bilingual dictionaries
-  % or from other sources.
   \item \emph{Languages} in PanLex are identified using ISO 639-3\footnote{\url{http://www.sil.org/iso639-3/codes.asp}} individual and macrolanguage codes, ISO 639-2\footnote{\url{http://www.loc.gov/standards/iso639-2/}} collective codes and ISO 639-5\footnote{\url{http://www.loc.gov/standards/iso639-5/}} codes.
-  \item \emph {Language varieties} allow one to make more fine grained
-  distinctions within a language. Their codes are composed of the language code
-  combined with a PanLex specific identifier.
-  For example, ``eng''
-  is the ISO 639-3 code for the English language.
+  \item \emph {Language varieties} allow one to make more fine grained distinctions within a language.
+    Their codes are composed of the language code combined with a PanLex specific identifier.
+    For example, ``eng'' is the ISO 639-3 code for the English language.
     Panlex defines various varieties of it, including ``English'' (eng-000), ``Simple English'' (eng-001) and ``British English'' (eng-005).
     Their labels, when possible, are autonyms, written in the native writing system.
-  So, in contrast to mechanisms like IETF BCP 47\footnote{\url{http://tools.ietf.org/html/bcp47}} there is no need for a transcription. %  and one can use variety names like ``{زبان اوستایی}''.
-  %Additionally these variety identifiers and labels do not have to be registered like the IETF BCP 47 language subtags\footnote{\url{http://www.iana.org/assignments/language-subtag-registry}}.
-  \item \emph{Meanings} in PanLex are entities of which each represents a unique
-  possible sense of an expression. 
-  %They are allocated as necessary by an editor based on his or her
- %interpretation of expressions.
+    So, in contrast to mechanisms like IETF BCP 47\footnote{\url{http://tools.ietf.org/html/bcp47}} there is no need for a transcription.
+  \item \emph{Meanings} in PanLex are entities of which each represents a unique possible sense of an expression.
     Meanings are assigned by editors based on their interpretation of expressions.
-  Usually this assignment is done on a per-source basis so that identical meanings
-  across multiple sources are not resolved.
-  %different
-  % meaning identifiers may refer to the same intended sense.
-  This means that if there is e.g. a translation of the fruit \emph{``apple''}
-  in an English to German dictionary and another translation from English to
-  French, these do not necessarily result in a single meaning entity linking to
-  all involved languages.
-  %the to corresponding expressions in English, German and French.
-  Instead, there could be \emph{``apple''} and \emph{``Apfel''} sharing one
-  meaning entity and \emph{``apple''} and \emph{``pomme''} sharing another.
-%  Editors may add informal definitions to meanings in any language variety.
-%  Information sources can be dictionaries, thesauri, etc., and also personal knowledge. 
-%  The pair comprised of the editor and the information source is referred to as a meaning's \emph{approver}.
+    Usually this assignment is done on a per-source basis so that identical meanings across multiple sources are not resolved.
+    This means that if there is e.g.
+a translation of the fruit \emph{``apple''} in an English to German dictionary and another translation from English to French, these do not necessarily result in a single meaning entity linking to all involved languages.
+    Instead, there could be \emph{``apple''} and \emph{``Apfel''} sharing one meaning entity and \emph{``apple''} and \emph{``pomme''} sharing another.
   \item \emph{Definitions} are optional descriptions of a meaning.
@@
     A description of the verb \emph{browse} for example could be \emph{``move or surf through various files on a computer, the Internet, etc.''}, marked as a definition in the ``English'' language variety.
-  \item \emph{Denotations} are entities that relate expressions to meanings and
-  may optionally carry annotations in form of sets of key value pairs. For
-  instance, an English expression \emph{pig}, when referring to
-  \emph{police officer}, could be annotated with \emph{pragmatics=vulgar}. 
-  Furthermore, denotations can be tagged with part-of-speech tags, such as word
-  classes, selected from a closed list based on the Open Lexicon Interchange Format (OLIF) standard\footnote{\url{http://www.olif.net/}}.
-  For example, \emph{fall} can be a verb or a noun for autumn. 
-  Homonyms are those expressions that are connected to multiple meanings.
+  \item \emph{Denotations} are entities that relate expressions to meanings and may optionally carry annotations in form of sets of key value pairs.
+    For instance, an English expression \emph{pig}, when referring to \emph{police officer}, could be annotated with \emph{pragmatics=vulgar}.
+    Furthermore, denotations can be tagged with part-of-speech tags, such as word classes, selected from a closed list based on the Open Lexicon Interchange Format (OLIF) standard\footnote{\url{http://www.olif.net/}}.
+    %For example, \emph{fall} can be a verb or a noun for autumn.
+    %Homonyms are those expressions that are connected to multiple meanings.
   \item \emph{Users} have editorial privileges over the language varieties and the approvers that they define.
@@
     At present there are ten different license categories an approver can be annotated with.
-  They are \emph{public domain}, \emph{Creative Commons}, \emph{request} (meaning that one has to ask the author of the resource), \emph{GNU General Public License}, \emph{GNU Lesser General Public License}, \emph{GNU Free Documentation License}, \emph{MIT License}, \emph{copyright} (stating that there is a certain copyright holder), \emph{other} and \emph{unknown}.
+    They are \emph{public domain}, \emph{Creative Commons (CC)}, \emph{request} (meaning that one has to ask the author of the resource), \emph{GNU General Public License (GPL)}, \emph{GNU Lesser General Public License (LGPL)}, \emph{GNU Free Documentation License (FDL)}, \emph{MIT License}, \emph{copyright} (stating that there is a certain copyright holder), \emph{other} and \emph{unknown}.
+    The license distribution is shown in \autoref{tbl:plx-license-counts}.
 \end{itemize}
+
+\begin{table}
+\centering
+\begin{scriptsize}
+\begin{tabular}{lrclrclr}
+License          & Count &&
+License          & Count &&
+License          & Count \\
+\cline{1-2} \cline{4-5} \cline{7-8}
+\emph{unknown}   &  1886 && % 1st
+\emph{PD}        & 123 && % 2nd
+\emph{LGPL}      &     9 \\ % 3rd
+\emph{copyright} &  1212 && % 1st
+\emph{other}     &   104 && % 2nd
+\emph{request}   &     5 \\ % 3rd
+\emph{CC}        &   335 && % 1st
+\emph{MIT}       &    32 && % 2nd
+\\
+\emph{GPL}       &   172 && % 1st
+\emph{FDL}       &    24 && % 2nd
+\end{tabular}
+\end{scriptsize}
+\caption{Number of approvers using a certain license}
+\label{tbl:plx-license-counts}
+\end{table}
 
@@
 \begin{table}
-\centering
-\begin{scriptsize}
-\begin{tabular}{lr}
-\toprule
+  \centering\begin{scriptsize}
+  \begin{tabular}{lrclr}
+    Entity             & Instances  &&
     Entity             & Instances  \\
-%\midrule
+    \cline{1-2} \cline{4-5}
+    Denotations        & 50,803,243 &&  % SELECT COUNT(dn) FROM dn;
+    Language Varieties &      7,248 \\  % SELECT COUNT(lv) FROM lv;
+    Meanings           & 20,023,427 &&  % SELECT COUNT(mn) FROM mn;
     Approvers          &      3,905 \\  % SELECT COUNT(ex) FROM ex;
-Expressions & 18,580,594 \\ % SELECT COUNT(ex) FROM ex;
-%\midrule
-Languages   & 7,839 \\ % SELECT COUNT(lc) FROM lc; -- all language codes (ISO 639-3, ISO 639-2, ISO 639-5) known in the PanLex db
-% FIXME: there should not be more languages than language varieties
-%\midrule
-Language Varieties & 7,248 \\ % SELECT COUNT(lv) FROM lv;
-%\midrule
-Meanings    & 20,023,427 \\ % SELECT COUNT(mn) FROM mn;
-%\midrule
-Definitions & 2,522,605 \\ % SELECT COUNT(df) FROM df;
-%\midrule
-Denotations & 50,803,243 \\ % SELECT COUNT(dn) FROM dn;
-%\midrule
+    Expressions        & 18,580,594 && % SELECT COUNT(ex) FROM ex;
+    Licenses           &         10 \\ % SELECT COUNT(li) FROM apli;
+    Definitions        &  2,522,605 && % SELECT COUNT(df) FROM df;
     Users              &          7 \\ % SELECT COUNT(us) FROM us;
-%\midrule
-Licenses    & 10 \\ % SELECT COUNT(li) FROM apli;
-\bottomrule
+    Languages          &      7,839 && % SELECT COUNT(lc) FROM lc; -- all language codes (ISO 639-3, ISO 639-2, ISO 639-5) known in the PanLex db
   \end{tabular}
@@
 
-%\todo{Possibly also: approver editors, language variety editors, exemplar/approved characters, approver varieties}
-
-%The same expression may be mentioned in multiple information sources.
-%However, an expression's possible meanings depend on the user's interpretion of them in context of such a source.
-%For this reason, every meaning depends on an approver.
-%\emph{Denotations} capture the mappings between expressions and their possible meanings.  
-Note that \emph{approver} combines a user with an \emph{information source},
-however the \emph{information source} is not modeled as a distinct entity.
-Also, the information of whether or not two meanings with different approvers
-are the same is not being captured.
-
-%These sources are fundamental for the PanLex database design and are aggregated in the approver table.
-%According to these approvers there are special meanings that are represented by expressions.
-%%Meanings are always related to an approver, which captures how a user interprets an expression information source.
-
-%So, if there are two expressions, they can either be synonyms (if they are given in the same language) or translations of each other.
-%\\
-%Expressions are lexemes that can be seen as entries of a lexicon or dictionary of a given language.
-%They are represented in their lemma or dictionary form.
-
-%These expressions can be given in different language varieties allowing a meaning preserving translation between these varieties.
-%An expression is given in a language variety.
-%Such a variety can be a standard variety, a dialect, a script-based variety a controlled technical variety etc.
-%Examples would be English, Simple English, British English and American English.\\
-
-%  Ghidorah, the Three-Headed Monster | Ghidorah
-%  describe or draw attention to (a product, service, or event) in a public medium in order to promote sales or attendance | advertise
-%  move or surf through various files on a computer, the Internet, etc. | browse
+Note that \emph{approver} combines a user with an \emph{information source}, however the \emph{information source} is not modeled as a distinct entity.
+Also, the information of whether or not two meanings with different approvers are the same is not being captured.
 
@@
 \label{sec:vocabulary}
-%The schema described above was adopted to create RDF triple structures.
-%is a good starting point most of the relations were taken over introducing a PanLex RDF vocabulary.
-The entities and relations of the schema described in the previous section
-serve as the base for the development of the PanLex RDF vocabulary.
-In general, all PanLex RDF resources reside
-in the namespace \texttt{\small <http://ld.panlex.org/plx/>}, abbreviated with
-\texttt{\small plx}.
-An example of the resulting ontology is depicted in \autoref{fig:vocabulary} and
-summarized as follows:
-Unless otherwise noted, the URIs of instances of PanLex classes follow the pattern
-\emph{plx:\{className\}/\{id\}}, where \{className\} is spelled in lower camel case
-and the \{id\} is the primary key of the corresponding database table.
+The entities and relations of the schema described in the previous section serve as the base for the development of the PanLex RDF vocabulary.
+In general, all PanLex RDF resources reside in the namespace \texttt{\small <http://ld.panlex.org/plx/>}, abbreviated with \texttt{\small plx}.
+An example of the resulting ontology is depicted in \autoref{fig:vocabulary} and summarized as follows:
+Unless otherwise noted, the URIs of instances of PanLex classes follow the pattern \emph{plx:\{className\}/\{id\}}, where \{className\} is spelled in lower camel case and the \{id\} is the primary key of the corresponding database table.
 
-%As in the original PanLex database there are also \emph{expressions} used in the RDF dataset.
 \begin{itemize}
   \item Expressions are modeled as instances of the class \texttt{\small plx:Expression}.
-Their original and normalized textual representations become the 
-values of the properties \texttt{\small rdfs:label} and \texttt{\small plx:degradedText}, respectively.
-Their corresponding language variety is stated
-using \texttt{\small plx:languageVariety}.
-
-\item For language and language varieties the classes
-\texttt{\small plx:Language} and \texttt{\small plx:LanguageVariety} are introduced.
-\emph{ISO 639-1} and \emph{ISO 639-3 codes} become instances of the classes
-\texttt{\small plx:Iso639-1Code} and \texttt{\small plx:Iso639-3Code}.
-
+    Their original and normalized textual representations become the values of the properties \texttt{\small rdfs:label} and \texttt{\small plx:degradedText}, respectively.
+    Their corresponding language variety is stated using \texttt{\small plx:languageVariety}.
+  \item For language and language varieties the classes \texttt{\small plx:Language} and \texttt{\small plx:LanguageVariety} are introduced.
+    \emph{ISO 639-1} and \emph{ISO 639-3 codes} become instances of the classes \texttt{\small plx:Iso639-1Code} and \texttt{\small plx:Iso639-3Code}.
   \item The RDF analog of the PanLex \emph{meaning} is the \texttt{\small plx:Meaning}.
-Entities of this class may have an identifier assigned with the
-\texttt{\small plx:identifier} property pointing to an \texttt{\small xsd:string} literal.
-%This meaning identifier represents ids given to the considered meaning in other
-%data sources and may at a later stage be used to build links to the meaning
-%as represented on these sites.
-Meanings may also have \emph{definitions}, entities of the
-\texttt{\small plx:Definition} class, giving a textual representation
-(\texttt{\small rdfs:label})
-%with more than just a lexeme
-in a certain
-language variety (\texttt{\small plx:languageVariety}).
-
-%These resources have an \texttt{rdfs:label} containing the textual representation of the considered expression.
-%Additionally there is a normalized version of the label, reachable via the \texttt{plx:degradedText} property.
-%Expression entities also have an assigned language variety (\texttt{plx:languageVariety}).
-%It refers to the language variety the textual representation is written in.
+    Entities of this class may have an identifier assigned with the \texttt{\small plx:identifier} property pointing to an \texttt{\small xsd:string} literal.
+    Meanings may also have \emph{definitions}, entities of the \texttt{\small plx:Definition} class, giving a textual representation (\texttt{\small rdfs:label}) in a certain language variety (\texttt{\small plx:languageVariety}).
   \item Following the semantics of the PanLex database, meanings and expressions are linked via \emph{denotations}.
-These are entities of the \texttt{\small plx:Denotation} class pointing
-to meanings and expressions via the properties \texttt{\small plx:denotationMeaning}
-and \\ \texttt{\small plx:denotationExpression}. % manual linebreak
+    These are entities of the \texttt{\small plx:Denotation} class pointing to meanings and expressions via the properties \texttt{\small plx:denotationMeaning} and \\ \texttt{\small plx:denotationExpression}.
     Denotations may also have a word class assigned to them.
     This can be achieved with the denotation's \texttt{\small plx:wordClass} property pointing to a \texttt{\small plx:WordClass} entity.
-\item All approvers share the \texttt{\small plx:Approver} class. The characteristics of an approver are described using mainly triples with literal objects.
-These are for example %\texttt{\small plx:registrationDate} pointing to an \texttt{\small xsd:date} literal,
-\texttt{\small dc:title} to assign the title of a source,
-\texttt{\small dc:creator} to give an \texttt{\small xsd:string} containing the author's name.
+  \item All approvers share the \texttt{\small plx:Approver} class.
+    The characteristics of an approver are described using mainly triples with literal objects.
+    These are for example \texttt{\small dc:title} to assign the title of a source, \texttt{\small dc:creator} to give an \texttt{\small xsd:string} containing the author's name.
 \end{itemize}
-
-%These entities are mainly used for interlinking to the Lexvo dataset\footnote{\url{http://lexvo.org/ontology}}.
-%See \autoref{sec:linking} for more details.
-
-%Besides the approvers there are also \emph{languages} and \emph{language varieties} taken over into the RDF vocabulary.
-%They are of the classes \texttt{plx:Language} and \texttt{plx:LanguageVariety} respectively.
-%Language varieties have a name given by the \texttt{rdfs:label} property and belong to a certain language referred to via \texttt{plx:languageVarietyOf}.
-%For those languages that have an ISO 639-3 language code, this code is referenced via the \texttt{plx:iso639-3Code} property.
-%The equivalent ISO 639-1 code can be retrieved using the \texttt{plx:iso639-1Code} property.
-
-%The structure of \emph{approvers} is nearly the same as in the database schema.
-% etc.
-%All properties defined for \texttt{plx:Approver} instances are shown in \autoref{tbl:vocabulary}.
-
-%Besides the approvers there are also \emph{languages} and \emph{language varieties} taken over into the RDF vocabulary.
-%They are of the classes \texttt{plx:Language} and \texttt{plx:LanguageVariety} respectively.
-%Language varieties have a name given by the \texttt{rdfs:label} property and belong to a certain language referred to via \texttt{plx:languageVarietyOf}.
-%For those languages that have an ISO 639-3 language code, this code is referenced via the \texttt{plx:iso639-3Code} property.
-%The equivalent ISO 639-1 code can be retrieved using the \texttt{plx:iso639-1Code} property.
-
 Since the PanLex project compiled its database by extracting data from different sources, the licenses of these sources were also considered.
 At present, we support the different license categories given in the database by creating resources of the \texttt{\small plx:License} class.
-%with appropriate descriptions in their \texttt{rdfs:label}.
-%In the RDF dataset these have an additional \texttt{rdfs:label} property giving the names or descriptions 
+
 \begin{table}
   \begin{tiny}
-\begin{tabular}{ll}
+  \begin{tabular}{p{52px}p{140px}}
     Class                 & Properties \\
     \toprule
-\texttt{plx:Approver} & \texttt{plx:registrationDate}, \texttt{rdfs:label}\\
-                      & \texttt{foaf:homepage}, \texttt{plx:license}, \texttt{dc:date}\\
-                      & \texttt{dc:creator}, \texttt{dc:title}, \texttt{plx:quality} \\
-                      & \texttt{dc:publisher}, \texttt{dbpedia-owl:isbn} \\
+    \texttt{plx:Approver} & \mbox{\texttt{plx:registrationDate}, \texttt{rdfs:label}, \texttt{dc:title},}
+                            \mbox{\texttt{dc:creator}, \texttt{plx:license}, \texttt{dc:date}, \texttt{plx:quality},}
+                            \mbox{\texttt{foaf:homepage}, \texttt{dc:publisher}, \texttt{dbpedia-owl:isbn}} \\
     \midrule
-\texttt{plx:Language}
-                      & \texttt{plx:iso639-3Code} \\
-                      & \texttt{plx:iso639-1Code} \\
+    \texttt{plx:Language} & \texttt{plx:iso639-3Code}, \texttt{plx:iso639-1Code} \\
     \midrule
@@
     \texttt{plx:Expression}
-                      & \texttt{plx:languageVariety} \\
-                      & \texttt{plx:degradedText}, \texttt{rdfs:label} \\
+                          & \mbox{\texttt{plx:languageVariety}, \texttt{plx:degradedText},} \texttt{rdfs:label} \\
     \midrule
     \texttt{plx:Meaning}
-                      & \texttt{plx:approver}, \texttt{plx:identifier} \\
-                      & \texttt{plx:meaningDefinition} \\
+                          & \mbox{\texttt{plx:approver}, \texttt{plx:identifier},} \texttt{plx:meaningDefinition} \\
     \midrule
@@
     \texttt{plx:Denotation}
-                      & \texttt{plx:denotationMeaning} \\
-                      & \texttt{plx:denotationExpression} \\
-                      & \texttt{plx:wordClass} \\
+                          & \texttt{plx:denotationMeaning}, \texttt{plx:denotationExpression}, \texttt{plx:wordClass} \\
     \midrule
@@
 
-%An overview of the PanLex vocabulary is shown in \autoref{fig:vocabulary}.%
 \begin{figure}
@@
 
+\begin{table}
+  \centering\begin{scriptsize}
+  \begin{tabular}{p{48px}p{62px}p{70px}}
+    Panlex                  & lemon                       & GOLD \\
+    \midrule
+    \texttt{plx:Denotation} & --                          & \texttt{gold:LinguisticSign} \\
+    \texttt{plx:Meaning}    & \texttt{lemon:LexicalSense} & \texttt{gold:SemanticUnit} \\
+    \texttt{plx:Expression} & \texttt{lemon:LexicalEntry} & \texttt{gold:FormUnit} \\
+  \end{tabular}
+  \end{scriptsize}
+  \caption{Classes considered to be similar across the re-used vocabulary}
+  \label{tbl:sameclasses}
+\end{table}
+
+\subsection{Vocabulary Reuse}
+\label{sec:vocabulary-reuse}
+The PanLex vocabulary is based on PanLex's conceptual schema and
+enables all of PanLex's data to be directly exposed as RDF.
+%Apart from modeling a PanLex RDF vocabulary
+Additionally, we also re-use existing vocuabularies, namely
+the \emph{Lexicon Model for Ontologies} (lemon)~\cite{lemon2011} as well as the \emph{General Ontology for Linguistic Description} (GOLD)~\cite{farr2003}.
+Since these models differ from the PanLex one to some extent, we follow an
+incremental approach of aligning the PanLex data with them.
+%accordingly. 
+% one differ to some
+%extend, not the whole schemas were implemented.
+\autoref{tbl:sameclasses} shows PanLex classes with their current counterparts
+in lemon and GOLD respectively.
+The parts implemented in our RDF conversion are displayed in \autoref{fig:goldlemonimplemented}.
+\begin{figure}
+  \centering
+  \includegraphics[width=\linewidth]{images/pdf/gold_lemon_implemented.pdf}
+  \caption{Parts of the GOLD (left) and lemon model (right) re-used in PanLex (URI prefixes are omitted for brevity)}
+  \label{fig:goldlemonimplemented}
+\end{figure}
+
 \subsection{Transformation workflow}
 \label{sec:conversion}
-New sources are added to PanLex on almost a daily basis.   
-The size of the PanLex database (including indexes) is currently approximately
-18 GB, which is big enough to make recurrent RDF conversion cumbersome:
-The RDF file of a full conversion of a database is much larger than the
-database itself.
-Using conventional hardware, it takes impractically long
-to convert all the data. This makes
-testing and debugging and fixing issues in the modeling or data conversion of the
-data very time-consuming.
-As the PanLex data already resides in a relational database, the use of a virtual
-RDB2RDF\footnote{\url{http://www.w3.org/2001/sw/wiki/RDB2RDF}}  mapping solution is a natural choice.
-The \emph{Sparqlify system}\footnote{\url{https://github.com/AKSW/Sparqlify}} offers,
-besides an efficient query rewriting engine, also a very easy-to-use mapping
-language, called \emph{Sparqlification Mapping Language} (SML).
+Since new sources are added to the PanLex database on almost a daily basis and because of its current size (\textasciitilde18 GB) an approach of recurrent conversions is impractical.
+%Using conventional hardware, a full conversion takes infeasibly long.
+%This makes the testing and debugging of the modeled vocabularies and data very
+% time-consuming.
+As the PanLex data already resides in a relational database, the use of a virtual RDB2RDF\footnote{\url{http://www.w3.org/2001/sw/wiki/RDB2RDF}} mapping solution is a natural choice.
+The \emph{Sparqlify system}\footnote{\url{https://github.com/AKSW/Sparqlify}} offers, besides an efficient query rewriting engine, also a very easy-to-use mapping language, called \emph{Sparqlification Mapping Language} (SML).
 Essentially, these mappings consist of three clauses:
-\begin{itemize}
-  \item \emph{From:} Specifies the logical SQL table (i.e. table, view or query) underlying the SML view.
-  \item \emph{With:} Defines SPARQL variables by means of expressions over relational columns that yield RDF terms.
-  \item \emph{Construct:} The template (i.e. the set of triple patterns) to be generated in this view, based on the SPARQL variable definitions.
-\end{itemize}
+The \emph{From} clause specifies the logical SQL table (i.e. table, view or
+query) to be used in the SML view.
+The \emph{With} clause binds a set of SPARQL variables to expressions that
+yield RDF terms from relational columns.
+Finally, the \emph{Construct} clause holds a set of triple patterns.
+%clause making use of the defined SPARQL variable definitions.
 \autoref{fig:ex:sparqlify-ml} shows an example of an SML view for the
-languages in PanLex: From each row of the \emph{i1} table three resources are
-created from the \emph{iso3} column and 
+languages in PanLex: From each row of the table \emph{i1} three resources are
+created based on the \emph{iso3} column and 
 bound to the variable names \emph{?lang}, \emph{?iso3} and \emph{?lexvo3}.
@@
 This view-based approach also demonstrates that changing the vocabulary or
-adding support for new ones does not require an extract transform load (ETL) process, and can therefore be done with little effort.
-%\todo{Maybe we should link to lexvo directly.}
-%PanLex's languages  iso 639-3 code 
-%typed with plx:Language
-%resources are created for
-%the languages in PanLex, linked with \emph{lexvo}, and
-
-%how meanings are mapped to their definition text:
-%Two resources are generated, namely \texttt{?mn} and \texttt{?mt}, which are typed with \texttt{plx:Meaning} and \texttt{plx:MeaningText} respectively.
-%Note that we could not simply use \texttt{rdfs:label} for associating the definition text to the meaning resource directly, because
-%the ISO 639-3 language code and PanLex language varieties do not directly map to valid RDF language tags.\todo{cite the standard} 
-%Instead, we retained the variety information by attaching it to the meaning text resource.
-%The with clause specifies the creation of URIs in the panlex namespace based on the respective entities database IDs.
-%Furthermore, it defines the definition text to become a plain literal.
+adding support for new ones does not require an extract transform load (ETL) process,
+and can therefore be done with little effort.
 
@@
 \begin{lstlisting}
-Create View i1 As
-  Construct {
-    ?lang a plx:Language ;
-          a <http://schema.org/Language> ;
+Create View i1 As Construct {
+    ?lang a plx:Language, <http://schema.org/Language> ;
           plx:iso639-3Code ?iso3 .
-
     ?iso3 a plx:Iso639-3Code ;
-          owl:sameAs ?lexvo3 .
-  }
+          owl:sameAs ?lexvo3 .  }
   With
@@
       ?lexvo3 = uri('http://lexvo.org/id/iso639-3/', ?iso3)
-  From
-    i1
+  From [[SELECT iso3 FROM i1]]
 \end{lstlisting}
@@
 \label{sec:linking}
-The SML view in the previous section (\autoref{fig:ex:sparqlify-ml}) already
-established the interlinking of the PanLex languages with Lexvo.
+The SML view in the previous section (\autoref{fig:ex:sparqlify-ml}) already established the interlinking of the PanLex languages with Lexvo.
 In this section we outline the interlinking with DBpedia.
-
-%In this section we first describe how we interlinked PanLex's language codes
-%with Lexvo\footnote{\url{http://www.lexvo.org/}} and the expressions with DBpedia\footnote{\url{http://dbpedia.org}}.
-%Afterwards we give an overview of how the dataset and the related resources are published.
-
-%Since Lexvo covers all ISO 639-3 language codes, interlinking could be done with
-%the Sparqlify-ML view in Figure~\ref{fig:ex:sparqlify-ml}.
-For DBpedia we were interested in creating \emph{valid} and
-thus \emph{dereferenceable} links.
-Therefore, we iterated the \emph{titles} datasets\footnote{\url{http://wiki.dbpedia.org/Downloads38}% - Note that more languages are available in the raw download section at \url{http://downloads.dbpedia.org/3.8/}},
-},
-which map (non-localized) DBpedia URIs to their page titles in the respective language.
-For each language version we normalized the labels by applying
-Unicode NFKD\footnote{\url{http://unicode.org/reports/tr15/}} normalization and
-removal of punctuation characters.
-% PanLex now uses NFC, not NFKD, for character normalization and does not remove punctuation.
+For DBpedia, we were interested in creating \emph{valid} and thus \emph{dereferenceable} links.
+Therefore, we iterated the \emph{titles} datasets\footnote{\url{http://wiki.dbpedia.org/Downloads38}}, which map (non-localized) DBpedia URIs to their page titles in the respective language.
+For each language version we normalized the labels by applying Unicode NFKD\footnote{\url{http://unicode.org/reports/tr15/}} normalization and removal of punctuation characters.
 Each DBpedia resource was then mapped to the PanLex expression that was equal to the resource's normalized label in the respective language.
@@
 
-In total, about 2.5 million links were obtained for approx. 20 million
-expressions. This relatively low coverage can be attributed to
-frequently appearing multi-word expressions that do not match the DBpedia titles well, and
-the fact that in this work we yet only considered DBpedia datasets for
-mainstream languages, whereas PanLex focuses on low-density ones.
+In total, about 2.5 million links were obtained for approx. 20 million expressions.
+This relatively low coverage can be attributed to frequently appearing multi-word expressions that do not match the DBpedia titles well, and the fact that in this work we yet only considered DBpedia datasets for mainstream languages, whereas PanLex focuses on low-density ones.
 
-%The relatively low coverage of these links (about 2,500,000 links for about 20,000,000 expressions) can be attributed to the frequently appearing multi-word expressions that do not match well to titles and the fact that the considered DBpedia datasets are mainly given for mainstream languages but not for the low-density ones in focus of PanLex.
 \begin{table}
-\centering
-\begin{scriptsize}
-\begin{tabular}{lllr}
-\toprule
-Language & 639-1 & 639-3 & Links \\
-\midrule
-English    & en & eng & 1,415,241 \\
-German     & de & deu &   224,146 \\
-French     & fr & fra &   187,364 \\
-Italian    & it & ita &   147,485 \\
-Spanish    & sp & spa &   117,056 \\
-Portuguese & pt & por &   112,266 \\
-Polish     & pl & pol &   110,974 \\
-Russian    & ru & rus &    68,040 \\
-Czech      & cs & ces &    28,767 \\
-Catalan    & ca & cat &    27,779 \\
-Korean     & ko & kor &    24,912 \\
-Turkish    & tr & tur &    22,258 \\
-Bulgarian  & bg & bul &    19,431 \\
-Hungarian  & hu & hun &    18,203 \\
-Slovene    & sl & slv &    11,981 \\
-Greek      & el & ell &     1,112 \\
-\midrule
-Total      &    &     & 2,537,015 \\
-\bottomrule
+  \centering\begin{scriptsize}
+
+  \begin{tabular}{lrclr}
+    Language   &   Links   && Language  & Links     \\
+    \cline{1-2}\cline{4-5}
+    English    & 1,415,241 && Catalan   &    27,779 \\
+    German     &   224,146 && Korean    &    24,912 \\
+    French     &   187,364 && Turkish   &    22,258 \\
+    Italian    &   147,485 && Bulgarian &    19,431 \\
+    Spanish    &   117,056 && Hungarian &    18,203 \\
+    Portuguese &   112,266 && Slovene   &    11,981 \\
+    Polish     &   110,974 && Greek     &     1,112 \\
+    \cline{4-5}
+    Russian    &    68,040 \\
+    Czech      &    28,767 && Total     & 2,537,015 \\
   \end{tabular}
@@
 
-% // (i1)
-% // ISO 639-3 language name links to lexvo.org
-% Create View i1 As
-%     Construct {
-%         ?iso1 a plx:Iso639-1Code.
-%         ?iso3 a plx:Iso639-3Code.
-%         ?iso1 owl:sameAs ?lexvo1.
-%         ?iso3 owl:sameAs ?lexvo3.
-%         ?lang a plx:Language.
-%         ?lang a lvont:Language.
-%         ?lang a <http://schema.org/Language>.
-%         ?lang plx:iso639-1Code ?iso1.
-%         ?lang plx:iso639-3Code ?iso3.
-%     }
-%     With
-%         ?iso1 = uri(concat(plx:iso639-1, '/', ?iso1))
-%         ?iso3 = uri(concat(plx:iso639-3, '/', ?iso3))
-%         ?lexvo1 = uri(concat('http://lexvo.org/id/iso639-1/', ?iso1))
-%         ?lexvo3 = uri(concat('http://lexvo.org/id/iso639-3/', ?iso3))
-%         ?lang = uri(concat(plx:language, '/', ?iso3))
-%     From
-%         i1
-%
-
 \section{Publishing}
 \label{sec:publishing}
-% PanLex itself already provides direct access to the PostgreSQL database for scientists,
-% several user interfaces\footnote{\url{http://panlex.org/try/}} and an
-% API\footnote{\url{http://panlex.org/tech/doc/api/}}.
-As stated in~\autoref{sec:intro}, the PanLex project already provides several
-interfaces for data access.
-%The usage of these interfaces is not restricted in order to make the provided lexical
-%translations as accessible as possible.
-With our RDF conversion work, we complement these interfaces with
-Linked Data
-%\footnote{\url{http://panlex.org/plx/TODO}}
-and an SPARQL
-endpoint\footnote{\url{http://ld.panlex.org/sparql} and \url{http://ld.panlex.org/snorql}}.
+With our RDF conversion work, we complement existing
+APIs\footnote{\url{http://panlex.org/try/}} with Linked Data, powered by
+Pubby\footnote{\tiny{\url{http://wifo5-03.informatik.uni-mannheim.de/pubby/}}},
+and two SPARQL
+endpoints\footnote{\url{http://ld.panlex.org/vsparql}}
+\footnote{\url{http://ld.panlex.org/sparql}}, powered by Sparqlify and Virtuoso,
+respectively.
+% and \url{http://ld.panlex.org/snorql}}.
 An overview is shown in ~\autoref{fig:panlex-architecture}.
-Our SML views and the code for interlinking are
-hosted on GitHub\footnote{\url{https://github.com/AKSW/PanLex-2-RDF}}.
-The created linksets are hosted in the PanLex database and published together
-with the other data using Sparqlify.
-Downloads and additional information are available on the Panlex-RDF
-project page\footnote{\url{http://ld.panlex.org}}.
+The SPARQL browser
+\emph{SNORQL}\footnote{\url{https://github.com/kurtjx/SNORQL}} can be accessed
+by replacing \emph{sparql} with \emph{snorql} in the respective links. Our SML
+views and the interlinking code are hosted on GitHub\footnote{\url{https://github.com/AKSW/PanLex-2-RDF}}.
+The created linksets are hosted in the PanLex database and are published together with the other data using the Sparqlify RDB2RDF tool.
+Finally, we offer downloads tagged with timestamps of their creation\footnote{\url{http://ld.panlex.org/downloads/releases/}}.
 \begin{figure}
 \centering
-\includegraphics[width=0.4\textwidth]{images/PanLex-SW-Architecture-New}
+\includegraphics[width=0.5\textwidth]{images/pdf/sparqlify_setup02.pdf}
 \caption{PanLex architecture}
@@
 \label{sec:usage}
-There are general benefits of RDF conversions,
-such as the paradigm shift towards thinking about what one intends to express
-with the data, making this meaning explicit using ontologies,
-enabling of interlinking, easing data integration tasks and uniform
-data access via Linked Data and SPARQL.
+There are general benefits of using Semantic Web technologies, such as the
+potential for simplified data integration due to RDF and vocabulary reuse, the
+possibility of enriching data based on interlinking, drawing advantage from
+reasoning and the exploration of the data through the use of generic Semantic
+Web tools.
+Moreover, some applications, like the TeraDict translation lookup
+service\footnote{\url{http://panlex.org/teradict/?lg=eng}}, can now be realized
+using SPARQL queries and so easily integrated in other applications.
+Due to space considerations, we refer the reader to the PanLex Linked Data
+landing page\footnote{\url{http://ld.panlex.org}}, where a
+collection of SPARQL queries is maintained.
+Also, since PanLex covers a niche of providing linguistic data for
+non-mainstream languages, investigation of its fitness for use in cross
+language information retrieval, as well as annotation projects, like DBpedia
+Spotlight\footnote{\url{http://spotlight.dbpedia.org}}, seems worthwhile.
 
-With the RDF version of PanLex, a large lexical resource participates in the Web of Data,
-making it available to a broader community. Users can easily explore the PanLex data using
-the services provided by SNORQL\footnote{\url{https://github.com/kurtjx/SNORQL}} and Pubby\footnote{\url{http://wifo5-03.informatik.uni-mannheim.de/pubby/}}.
-This also increases the chance of development of
-new interesting mashups.
-For example, the TeraDict translation service\footnote{\url{http://panlex.org/teradict/?lg=eng}}
-could now be easily realized using simple SPARQL queries.
-% as shown in
-%\todo{Create the query: translations in a given target language of an expression}. 
-Another future usage scenario is to link PanLex to Wortschatz\footnote{\url{http://wortschatz.uni-leipzig.de/}} with the aim of 
-improving the recall of word-by-word translations by considering their co-occurrences.
+%exploiting data of projects like Wordnet, e.g. via the LemonWordNet
+% dataset\footnote{\url{http://datahub.io/de/dataset/lemonwordnet}} can help
+ % finding errors or derive new data and improve translations.
+%Another source that could be combined with the PanLex data is the Wortschatz project\footnote{\url{http://wortschatz.uni-leipzig.de/}}.
+%Using Wortschatz's co-occurrences statistics of words in different languages one could also improve the recall of word-by-word translations.
+
+%(see \autoref{lst:terradict} for an example).
+
+%that currently only support mainstream languages.
+%Apart from this users can easily explore the PanLex data using the
+%Linked Data and SPARQL services.
+%This also increases the chance of development of new interesting mashups.
+% vision web of data
+%% \begin{figure}
+%% \centering
+%% \begin{lstlisting}
+%% SELECT ?expr_lbl WHERE {
+%%     ?expr_lang rdfs:label "language" .
+%%     ?denotation_lang plx:denotationExpression ?expr_lang .
+%%     ?denotation_lang plx:denotationMeaning ?meaning .
+%%     ?denotation_other plx:denotationMeaning ?meaning .
+%%     ?denotation_other plx:denotationExpression ?expr_other .
+%%     ?expr_other rdfs:label ?expr_lbl .
+%%     FILTER(?expr_lang != ?expr_other)
+%% }
+%% \end{lstlisting}
+%% \caption{SPARQL query to retrieve all known lexical translations of the given English word \emph{`language'}}
+%% \label{lst:terradict}
+%% \end{figure}
 
@@
 \label{sec:related}
-PanLex is an integration project of many existing lexical resources. % with
-% the goal of eventually covering all languages in the world, especially
-% ones having few speakers.
-The extraction of information from linguistic sources, and
-techniques for automatically inferring translations, are
-relevant work discussed in \cite{panlex_probtrans}.
-
-An important initiative is the
-\emph{Global Wordnet Association}\footnote{\url{http://www.globalwordnet.org/}},
+PanLex is an integration project of many existing lexical resources.
+The extraction of information from linguistic sources, and techniques for automatically inferring translations,
+are relevant work discussed in \cite{panlex_probtrans}.
+%A similar approach of linking expressions to meanings is followed by the
+% \emph{Global Wordnet Association}.
+An important initiative is the Global Wordnet Association\footnote{\url{http://www.globalwordnet.org/}\vspace{0.7em}},
 which offers a platform for sharing wordnets and defines several goals. 
@@
 of different languages and establishing a universal index of meaning.
-%
-%It draws from the EuroWordNet~\cite{eurowordnet} and the  
-%Manually engineered dictionaries %such as EuroWordNet [38]
-% andcurrently mainly used for research purposes\todo{correct?}.
-%A list of the resources used in panlex is compiled on the website\footnote{\url{http://panlex.org/tech/plrefs.shtml}}.
-%In this section we discuss work related to the extraction and inference
-%approaches used in PanLex, and work on the Semantic Web. 
-%
-%- There has been considerable research on methods to acquire translation lexicons from either
-%MRDs (Neff and McCord, 1990; Helmreich et
-%al., 1993; Copestake et al., 1994) or from par-
-%allel text (Gale and Church, 1991; Fung, 1995;
-%Melamed, 1997; Franz et al., 2001), but this has
-%generally been limited to a small number of lan-
-%guages.
-%- Manually engineered dictionaries such as
-%EuroWordNet (Vossen, 1998) (whats the name of the new project?)
-%- TRANS GRAPH algorithm (Etzioni etal., 2007).
-%
-At the level of the Semantic Web, there is currently a trend in publishing linguistic resources
-%as Linked Data\footnote{\url{http://www.w3.org/DesignIssues/LinkedData.html}} and via SPARQL{\footnote{\url{http://www.w3.org/TR/sparql11-query/}}}.
-as Linked Data %\footnote{\url{http://www.w3.org/DesignIssues/LinkedData.html}}
-and via SPARQL.
-These efforts are referred to as the \emph{Linguistic Linked Open Data Cloud} (LLOD). 
-%Some of these datasets already make use of quasi-standard linguistic Semantic Web Ontolgies, such as  
-Several (quasi-)standard ontologies have been developed for covering different aspects of linguistic resources.
-Examples include the \emph{Ontologies of Linguistic Annotation} (OLiA)~\cite{olia2010},
-the \emph{Lexicon Model for Ontologies} (lemon)~\cite{lemon2011} as well as \emph{General Ontology for Linguistic Description} (GOLD)~\cite{farr2003} for modeling lexicon and machine-readable dictionaries,
-\emph{POWLA} for modeling linguistic corpora\cite{powla2012}
-and the \emph{Natural Language Processing Interchange Format} (NIF)\footnote{\url{http://nlp2rdf.org/nif-1-0}}.
-
-%\begin{itemize}
-%\item @author: Claus, Patrick
-%\item Research related to PanLex itself \url{http://panlex.org/research/}
-%\item Panlingual Translator: Word by word translation + back translation for verification, broken English (no grammar) ranked by
-%Do we have a comparison to wiktionary?
-
-%Difference to dict.cc\footnote{\url{http://dict.cc}}, leo.org\footnote{\url{http://leo.org}}, Wiktionary\footnote{\url{http://wiktionary.org}} and Wikipedia (DBpedia)
-
-%\item
-
-%\item Research in the same field as PanLex
-%\end{itemize}
+Wordnets are usually focused on the
+definition of \emph{synsets} and relations between them in a single language,
+whereas PanLex's focus is on capturing lexical translations between languages.
+Hence, these efforts are complementary.
+%As already mentioned in PanLex there can be several meaning entities
+% representing the same unique sense.
+In the Semantic Web context, several (quasi-)standard vocabularies and ontologies have been developed with the rise of the \emph{Linguistic Linked Open Data} movement.
+Examples include the
+\emph{Ontologies of Linguistic Annotation} (OLiA)~\cite{olia2010} for modeling
+lexicon and machine-readable dictionaries, \emph{POWLA} for modeling linguistic corpora\cite{powla2012} and the \emph{Natural Language Processing Interchange Format} (NIF)\footnote{\url{http://nlp2rdf.org/nif-1-0}}.
 
@@
 \label{sec:conclusion}
-In this dataset description we detailed the PanLex database and its conversion
-to RDF.
+In this dataset description we detailed the PanLex database and its conversion to RDF.
 Based on our URI and vocabulary design, we created appropriate view definitions
-for the virtual RDB2RDF solution \emph{Sparqlify} which carries out
-the actual RDF transformation.
-Furthermore, we interlinked the languages in PanLex with Lexvo,
-and created about 2.5 million links to DBpedia for expressions in 16 languages.
-Although not discussed explicitly due to space limitations, we
-also made first steps towards enhancing the RDB2RDF view definitions
-as to facilitate the integration of the PanLex data with the lemon and GOLD
-data models.
+for the \emph{Sparqlify} system, which carries out the actual RDF
+transformation.
+Furthermore, we interlinked the languages in PanLex with Lexvo, and created about 2.5 million links to DBpedia for expressions in 16 languages.
+With the integration of lemon and GOLD we also support the data access via external linguistic ontologies.
 
-In conclusion, we see this work as a valuable contribution to the vision of
-countering language attrition by making a vast amount of lexical
-resources available as Linked Open Data.
-
-% We see this work of making a vast amount of lexical
-% resources available as Linked Open Data as a valuable
-% contribution to the vision of countering language attrition.
-
-%We detailed the developed vocabulary and
-
-We identified some weaknesses in
-the original database design, which we intend to overcome in the future:
-The original dataset currently does not cleanly model information sources and approvers
-as distinct entities.
-Also, while it is possible to determine each meaning's and each denotation's approver, an approver can have multiple users with editorial privileges, so it is not possible to track each contributed, modified, and deleted meaning and denotation by editor.
-Retaining this information seems beneficial for translation approaches,
-as this enables for example the attribution of qualities and relevances to
-specific editors.
-
-%We made first steps towards the support of additional linguistic
-%ontology models: There
-%So apart from the PanLex data model, expression-meaning relations can also be
-%retrieved using the lemon model.
-%Another approach to reuse existing vocabularies is the integration of the
-% linguistic sign model of the GOLD standard.
-
-%\todo{Is there some quasi standard translation vocab we could re-use?}
-%\todo{lang codes for minor languages}
+There exist some shortcomings which we intend to overcome in the future:
+Modeling information sources and users as distinct entities would enable one
+to unequivocally relate meanings and denotations to them, which in turn would
+allow for a more fine grained attribution of qualities and relevances.
+This could be beneficial for translation approaches.
+Another important aspect is, that each of PanLex's information sources can be
+seen as a dataset on its own, and thus relations between them could be modeled
+with the VoID vocabulary\footnote{\url{http://rdfs.org/ns/void}}.
 

### 2014-02-18 panlex-rdf-paper/2012/MLOD_PanLex/paper_v4.tex
added paper_v4.tex

--- 2014-01-22 panlex-rdf-paper/2012/MLOD_PanLex/paper_v3.tex
@@
 \address[A]{University of Leipzig, \{pwestphal, cstadler\}@informatik.uni-leipzig.de}
-\address[B]{Long Now Foundation, San Francisco, pool@panlex.org}
+\address[B]{The Long Now Foundation, San Francisco, pool@panlex.org}
+
+%(%!TEX TS-program = xelatex)
+%(%!TEX encoding = UTF-8 Unicode)
 
@@
 At present, there are approximately 7,000 living languages in the world.
-However, some experts claim that the process of globalization may eventually lead to the world losing this linguistic diversity.
-The vision of the PanLex project is to help save these languages, especially low-density ones, by allowing them to be intertranslatable and thus to be a part of the Information Age.
-Semantic Web technologies can support achieving this goal, for reasons such as
-their capabilities of flexibly representing, interlinking and reasoning with 
-data, in our case particularly linguistic resources and annotations.
+However, some forecasters expect the world to lose most of this linguistic diversity.
+The mission of the PanLex project is to investigate whether it is feasible to make all languages intertranslatable and thus usable for world-wide communication, and, if so, whether that new situation will interfere with this expected language attrition.
+This mission entails, among other prerequisites, making the lexicons of the world's languages intertranslatable.
+To accomplish that, the project is systematically integrating and making publicly available all known lexical translations.
+Semantic Web technologies can enhance the power of PanLex by flexibly representing and reasoning with the content of its database and interlinking it with linguistic and other resources and annotations.
 %Using Semantic Web technologies can support achieving this goal, especially due
 % to inference capabilities and the interlinking of the PanLex data with other data sources.
-Conversely, an RDF version of PanLex makes a significant contribution towards
-improving the coverage of the Linguistic Web of Data, as to the best of our
-knowledge there exists no large scale Linked Data data set for panlingual translation of
-non-mainstream languages.
-In this dataset description paper we detail how we transformed the data of the PanLex project to RDF,
+Conversely, PanLex, with its unique collection of translation links between more than a billion pairs of lexemes from more than 9,000 language varieties, can improve the coverage of the Linguistic Web of Data.
+In this dataset description paper we detail how we transformed the content of the PanLex database to RDF,
 established conformance with the lemon and GOLD data models,
@@
 \begin{keyword}
-Multi-lingual Linked Open Data, PanLex, Lexcial Resource, RDF, RDB2RDF, Sparqlify
+Multilingual Linked Open Data, LLOD Cloud, PanLex, Lexical Resource, RDF, RDB2RDF, SPARQL, Sparqlify
 \end{keyword}
@@
 \label{sec:intro}
-At present, there are about 7,000 living languages in the world\footnote{See \url{http://www.sil.org/iso639-3/download.asp} for a list of registered languages}.
-Nonetheless, some experts claim that processes such as nation-state consolidation and globalization are producing language attrition so rapidly that up to 90\% of all languages alive today will be extinct within a century~\cite{endang_lang}.
-Theorists of biolinguistic diversity argue that the loss of language diversity, the loss of human biological knowledge, and the loss of species diversity are mutually supportive and thus that language preservation and revitalization are essential to the preservation of biological diversity~\cite{nettle}.
-The vision of the PanLex project is to help save these thousands of languages, especially those low-density ones that are threatened by extinction, by supporting their use in global communication.
-This requires panlingual translation: translation from any language of the world into any other.
-One of the crucial components of panlingual translation of discourses is panlingual lexical translation.
-PanLex is designed to support that component by making use of several thousand
-sources and artificial intelligence approaches.
+At present, there are about 7,000 living languages in the world\footnote{\url{http://www-01.sil.org/iso639-3/iso-639-3.tab}.
+As of 25 January 2014, this list contained 6,975 living individual languages.}.
+Nonetheless, processes such as urbanization, nation-state consolidation, and globalization appear to be producing language attrition, with, for example, the extinction or near-extinction in the last 60 years of from 10\% to over 75\% (depending on the world's region) of all living languages\cite{lang_crisis}.
+Theorists of biolinguistic diversity argue that the loss of language diversity, the loss of human biological knowledge, and the loss of species diversity are mutually supportive and thus that language preservation and revitalization are essential to the preservation of biological diversity\cite{nettle}.
+
+The mission of the PanLex project is to investigate the possibility that language attrition would be decreased or halted if utterances in any language were efficiently translatable into all other languages and, as a result, the speakers of all languages could use their languages for world-wide communication.
+Among the prerequisites of the panlingual translatablity of discourses would be the panlingual translatability of lexicons.
+In fact, lexical translation is almost the entirety of translation in some contexts (such as personal profiles, bibliographic subject headings, content tagging, product price lists, and web navigation controls).
+PanLex is designed to support that component by systematically integrating millions of lexical translations, documented in thousands of sources, into a database accessible to the research community, application developers, and the public.
+The database content can be interpreted as a graph linking millions of lexemes with one another in billions of ``is-a-translation-of'' relations, and permitting intelligent automated inference to trillions more such relations that have not been attested.
 %It documents the known lexical translations (translations of lexemes) among all
@@
 
-Apart from using PanLex's data for inferring lexical translations, a
-further direction of work is to generate added value by enriching this data with
-knowledge from other (linguistic) sources.
-Grounding on the idea of a Semantic Web, a complete stack of technologies has
-been devised and standardized by the
+Precisely because the citation forms of lexemes, which is what PanLex directly enables the translation of, are commonly used as labels in collections of data, it is reasonable to speculate that PanLex might facilitate cross-linguistic access to knowledge.
+The idea of a Semantic Web has led to a stack of technologies standardized by the
 \emph{World Wide Web Consortium}\footnote{\url{http://www.w3.org/standards/semanticweb/}}
-supporting the definition of a machine readable and interpretable Linked Data
-network, which has led to emergence of the \emph{Web of Data}.
+supporting a machine-readable and machine-interpretable Linked Data
+network. This has led to the concept and reality of a \emph{Web of Data}.
 %Technologies that support this aim are already available and in use in the so
@@
 In \autoref{sec:triplificate} we introduce the PanLex dataset, present our
-PanLex RDF vocabulary, explain how we transformed the one into the other, and
-how we established conformance with additional data models.
+PanLex RDF vocabulary, and explain how we transformed the one into the other and established conformance with additional data models.
 \autoref{sec:linking} is about how we linked to other datasets of the LLOD cloud,
-whereas \autoref{sec:publishing} is about the actual dataset publishing.
+and \autoref{sec:publishing} is about the publication of the dataset.
 Usage scenarios are given in \autoref{sec:usage}.
-In \autoref{sec:related} we discuss related work, and finally, in \autoref{sec:conclusion} we conclude our approach and give some hints to future work.
+In \autoref{sec:related} we discuss related work, and finally, in \autoref{sec:conclusion}, we conclude our approach and give some hints about future work.
 
@@
 In this section, we first provide an analysis of the PanLex dataset.
-Subsequently, we introduce our URI and vocabulary design which closely
+Subsequently, we introduce our URI and vocabulary design, which closely
 resembles PanLex's original conceptual model.
@@
 \label{sec:analysis}
-At the core of the PanLex \emph{project}, there is the PanLex \emph{database} which is created from the imports of thousands of lexical resources, such as mono- and multilingual dictionaries, glossaries, standards, and thesauri.
-The concrete list of used sources is available online\footnote{\url{http://panlex.org/tech/plrefs.shtml}}.
-The data derived from these sources comprise single- and multi-word expressions and meanings assigned to them.
-Conceptually, the PanLex database thus ``represents \emph{assertions} about the \emph{meanings} of \emph{expressions}''\footnote{\url{http://panlex.org/tech/doc/design/panlex-db-design.pdf}}.
-As of now, the database contains about 20 million meanings and 19 million expressions extracted from about 2,000 sources.
-The most important entities and relations of PanLex's conceptual model are
-depicted in \autoref{fig:db-schema} and are explained in more the detail in the following.
+At the core of the PanLex \emph{project}, there is the PanLex \emph{database}, which is created by editors who consult thousands of lexical sources, such as mono- and multilingual dictionaries, glossaries, standards, and thesauri.
+The list of sources that have been and are being consulted is available online\footnote{\url{http://panlex.org/tech/plrefs.shtml}}.
+The data based on these sources include single- and multi-word expressions, meanings assigned to them, and related information.
+Far from being mere copies of the sources, which would not make PanLex data interoperable, the data constitute editors' \emph{interpretations} of various sources' \emph{assertions} that two or more \emph{expressions} share a \emph{meaning}\footnote{\url{http://panlex.org/tech/doc/design/panlex-db-design.pdf}}.
+As of January 2014, the database contained about 20 million meanings and 20 million expressions in about 9,200 language varieties based on the consultation of about 1,500 sources by about 20 editors.
+The most important entities and relations of PanLex's conceptual model are depicted in \autoref{fig:db-schema} and are explained in more detail below.
 
@@
   \centering
-  \includegraphics[width=0.4\textwidth]{images/schema.png}
+  \includegraphics[width=0.4\textwidth]{images/schema-new.png}
   \caption{The PanLex database schema}
@@
 \begin{itemize}
-  \item The starting point of the data acquisition is the \emph{approver} entity: An editor processes the content of a certain mono- or multilingual source as mentioned above and adds it to the PanLex dataset.
-    This combination of a user and a source is referred to as an approver.
-  \item \emph{Expressions} are lemmas, i.e. dictionary entries.
-    For example, \emph{``go''} is a valid expression, whereas \emph{``went''} is not.
-    Expressions are always given in a language variety and can only be given once per language variety.
-  \item \emph{Languages} in PanLex are identified using ISO 639-3\footnote{\url{http://www.sil.org/iso639-3/codes.asp}} individual and macrolanguage codes, ISO 639-2\footnote{\url{http://www.loc.gov/standards/iso639-2/}} collective codes and ISO 639-5\footnote{\url{http://www.loc.gov/standards/iso639-5/}} codes.
-  \item \emph {Language varieties} allow one to make more fine grained distinctions within a language.
-    Their codes are composed of the language code combined with a PanLex specific identifier.
-    For example, ``eng'' is the ISO 639-3 code for the English language.
-    Panlex defines various varieties of it, including ``English'' (eng-000), ``Simple English'' (eng-001) and ``British English'' (eng-005).
-    Their labels, when possible, are autonyms, written in the native writing system.
-    So, in contrast to mechanisms like IETF BCP 47\footnote{\url{http://tools.ietf.org/html/bcp47}} there is no need for a transcription.
-  \item \emph{Meanings} in PanLex are entities of which each represents a unique possible sense of an expression.
-    Meanings are assigned by editors based on their interpretation of expressions.
-    Usually this assignment is done on a per-source basis so that identical meanings across multiple sources are not resolved.
-    This means that if there is e.g.
-a translation of the fruit \emph{``apple''} in an English to German dictionary and another translation from English to French, these do not necessarily result in a single meaning entity linking to all involved languages.
-    Instead, there could be \emph{``apple''} and \emph{``Apfel''} sharing one meaning entity and \emph{``apple''} and \emph{``pomme''} sharing another.
-  \item \emph{Definitions} are optional descriptions of a meaning.
-    They are given in a certain language variety.
-    A description of the verb \emph{browse} for example could be \emph{``move or surf through various files on a computer, the Internet, etc.''}, marked as a definition in the ``English'' language variety.
-  \item \emph{Denotations} are entities that relate expressions to meanings and may optionally carry annotations in form of sets of key value pairs.
-    For instance, an English expression \emph{pig}, when referring to \emph{police officer}, could be annotated with \emph{pragmatics=vulgar}.
-    Furthermore, denotations can be tagged with part-of-speech tags, such as word classes, selected from a closed list based on the Open Lexicon Interchange Format (OLIF) standard\footnote{\url{http://www.olif.net/}}.
+  \item The starting point of data acquisition is the \emph{source} entity: An editor consults some mono- or multilingual source as mentioned above, interprets it as making assertions about lexical translations, and adds schema-compliant representations of these assertions to the PanLex dataset.
+  \item \emph{Expressions} are lexical entities, each defined as a string of characters in some variety of some language.
+As a first approximation, expressions can be considered lemmas, i.e.
+dictionary-entry headwords.
+They differ from traditional headwords, however, in at least two ways.
+(1) Homographs, such as the verb ``hide'' (conceal) and the noun ``hide'' (animal skin) in English, are treated as a single expression in PanLex.
+(2) Multiword expressions, such as ``fall in love'', traditionally found in an entry headed by one of their words, such as ``fall'' or ``love'', are treated as independent expressions in PanLex.
+Expressions are strings of Unicode codepoints subjected to Normalization Form NFC, but this still makes it difficult to search precisely for expressions, so to assist in approximate searching PanLex also adds for each expression a degraded form permitting only lower-case letters and digits.
+    \item \emph{Languages} in PanLex are identified using ISO 639-3\footnote{\url{http://www.sil.org/iso639-3/codes.asp}} individual and macrolanguage codes, ISO 639-2\footnote{\url{http://www.loc.gov/standards/iso639-2/}} collective codes, and ISO 639-5\footnote{\url{http://www.loc.gov/standards/iso639-5/}} codes.
+    \item \emph {Language varieties} are the fundamental linguistic entities in PanLex.
+Expressions are in them, not in languages.
+The language-variety concept is fine-grained and permissive: there can be many varieties of a language, and editors can add varieties as needed.
+A language variety is identified with some language code, combined with an arbitrary integer differentiating it from other varieties of the same language.
+For example, six dialects of Ahtna are identified as ``aht-000'' through ``aht-005''.
+These symbolic 3-letter, 3-digit labels are, themselves, treated as a (controlled) language variety, so editors can clarify their referents by translating them into natural-language expressions and into expressions in other controlled language varieties, such as the IETF standard BCP 47\footnote{\url{http://tools.ietf.org/html/bcp47}}.
+Such controlled language varieties may be defined as prohibiting synonymy and/or ambiguity.
+    \item \emph{Meanings} in PanLex are entities with which expressions are semantically classified.
+An editor documents that two or more expressions are asserted to be translations or synonyms of one another by assigning a shared meaning to them.
+For example, depending on an editor's judgment, a source's translation of the German lexeme ``klingen'' into a sequence of alternative English lexemes ``ring, sound, seem'' can be documented by either (1) the assignment of all four expressions to the same meaning or (2) the assignment of two or three meanings to ``klingen'' and of one of those meanings to each of the English expressions.
+For practicality, PanLex treats expressions and meanings asymmetrically.
+Newly added translations automatically re-use existing matching expressions, but they don't re-use existing meanings, because there is no straightforward test of equivalence for meanings.
+PanLex generates new meanings as needed, and meanings are source-specific, leaving it for subsequent research to investigate which meanings should be considered equivalent and thus consolidated.
+Meanings can further have properties of three types.
+(1) \emph{Definitions} are descriptions of a meaning, consisting of arbitrary text strings annotated as being in particular language varieties.
+(2) \emph{Domains} are expressions (e.g., ``medicine'') that classify a meaning, but do not express it.
+(3) \emph{Meaning identifiers} are strings acting as references to identifiers used in a source.
+    \item \emph{Denotations} are assignments of meanings to expressions.
+Each denotation, in addition to its meaning-expression pair, may be assigned one or more \emph{word classes} (a closed set based on OLIF, the Open Lexicon Interchange Format) and/or \emph{metadata} (arbitrary strings paired as keys and values).
     %For example, \emph{fall} can be a verb or a noun for autumn.
     %Homonyms are those expressions that are connected to multiple meanings.
-  \item \emph{Users} have editorial privileges over the language varieties and the approvers that they define.
-  \item \emph{Licenses} are also considered by the PanLex project.
-    At present there are ten different license categories an approver can be annotated with.
-    They are \emph{public domain}, \emph{Creative Commons (CC)}, \emph{request} (meaning that one has to ask the author of the resource), \emph{GNU General Public License (GPL)}, \emph{GNU Lesser General Public License (LGPL)}, \emph{GNU Free Documentation License (FDL)}, \emph{MIT License}, \emph{copyright} (stating that there is a certain copyright holder), \emph{other} and \emph{unknown}.
-    The license distribution is shown in \autoref{tbl:plx-license-counts}.
+  \item \emph{Users} are the entities that may act as editors by defining sources and then defining data attributed to those sources, as well as defining language varieties.
+  \item Sources in PanLex are defined with about two dozen properties, among them being \emph{licenses}.
+As of January 2014 eleven different license categories were recognized.
+They were \emph{public domain}, \emph{Creative Commons (CC)}, \emph{request} (meaning that one has to ask the author of the resource), \emph{GNU General Public License (GPL)}, \emph{GNU Lesser General Public License (LGPL)}, \emph{GNU Free Documentation License (FDL)}, \emph{MIT License}, \emph{copyright} (meaning that there is a copyright claim without a more liberal license also being offered), \emph{PanLex Use Permission} (where an author or publisher has given a source to PanLex for its use), \emph{other}, and \emph{unknown}.
+    The distribution of licenses among sources that had been or were being consulted in January 2014 is shown in \autoref{tbl:plx-license-counts}.
 \end{itemize}
@@
 \cline{1-2} \cline{4-5} \cline{7-8}
-\emph{unknown}   &  1886 && % 1st
-\emph{PD}        & 123 && % 2nd
-\emph{LGPL}      &     9 \\ % 3rd
-\emph{copyright} &  1212 && % 1st
-\emph{other}     &   104 && % 2nd
-\emph{request}   &     5 \\ % 3rd
-\emph{CC}        &   335 && % 1st
-\emph{MIT}       &    32 && % 2nd
-\\
-\emph{GPL}       &   172 && % 1st
-\emph{FDL}       &    24 && % 2nd
+\emph{copyright} &  1343  && \emph{LGPL}      &     9 && \emph{PD}           & 149 \\
+\emph{CC}          &   387   && \emph{MIT}         &    32 && \emph{other}       &   106 \\
+\emph{FDL}        &    24    && \emph{PanLex}    &    7 && \emph{unknown}  &  1958 \\
+\emph{GPL}        &   172   && \emph{request}    &     5 \\
 \end{tabular}
 \end{scriptsize}
-\caption{Number of approvers using a certain license}
+\caption{Number of sources using a certain license}
 \label{tbl:plx-license-counts}
@@
 
-An overview of the number of instances per entity in the current PanLex database is given in~\autoref{fig:plx-entity-counts}.
+An overview of the number of instances per entity as of January 2014 is given in~\autoref{fig:plx-entity-counts}.
 \begin{table}
@@
   \begin{tabular}{lrclr}
-    Entity             & Instances  &&
     Entity             & Instances  \\
-    \cline{1-2} \cline{4-5}
-    Denotations        & 50,803,243 &&  % SELECT COUNT(dn) FROM dn;
-    Language Varieties &      7,248 \\  % SELECT COUNT(lv) FROM lv;
-    Meanings           & 20,023,427 &&  % SELECT COUNT(mn) FROM mn;
-    Approvers          &      3,905 \\  % SELECT COUNT(ex) FROM ex;
-    Expressions        & 18,580,594 && % SELECT COUNT(ex) FROM ex;
-    Licenses           &         10 \\ % SELECT COUNT(li) FROM apli;
-    Definitions        &  2,522,605 && % SELECT COUNT(df) FROM df;
-    Users              &          7 \\ % SELECT COUNT(us) FROM us;
-    Languages          &      7,839 && % SELECT COUNT(lc) FROM lc; -- all language codes (ISO 639-3, ISO 639-2, ISO 639-5) known in the PanLex db
+    \cline{1-2}
+    Denotations        & 54,278,860 \\  % SELECT COUNT(dn) FROM dn;
+    Languages          &      7,843 \\ % SELECT COUNT(lc) FROM lc; -- all language codes (ISO 639-3, ISO 639-2, ISO 639-5) known in the PanLex db
+    Language Varieties identified           &      9,310 \\  % SELECT COUNT(lv) FROM lv;
+    Language Varieties with data           &      9,239 \\  % 
+    Meanings           & 20,773,371 \\  % SELECT COUNT(mn) FROM mn;
+    Sources being consulted         &      4,190 \\  % 
+    Sources already consulted         &      1,453 \\  % 
+    Expressions        & 19,790,453 \\ % SELECT COUNT(ex) FROM ex;
+    Definitions        &  2,747,892 \\ % SELECT COUNT(df) FROM df;
+    Users              &          23 \\ % SELECT COUNT(us) FROM us;
   \end{tabular}
   \end{scriptsize}
-  \caption{Number of entities in the PanLex database}
+  \caption{Number of instances of main entities in the PanLex database}
   \label{fig:plx-entity-counts}
 \end{table}
-
-Note that \emph{approver} combines a user with an \emph{information source}, however the \emph{information source} is not modeled as a distinct entity.
-Also, the information of whether or not two meanings with different approvers are the same is not being captured.
 
@@
 In general, all PanLex RDF resources reside in the namespace \texttt{\small <http://ld.panlex.org/plx/>}, abbreviated with \texttt{\small plx}.
-An example of the resulting ontology is depicted in \autoref{fig:vocabulary} and summarized as follows:
+An example of the resulting ontology is depicted in \autoref{fig:vocabulary} and summarized as follows.
 Unless otherwise noted, the URIs of instances of PanLex classes follow the pattern \emph{plx:\{className\}/\{id\}}, where \{className\} is spelled in lower camel case and the \{id\} is the primary key of the corresponding database table.
@@
   \item Expressions are modeled as instances of the class \texttt{\small plx:Expression}.
-    Their original and normalized textual representations become the values of the properties \texttt{\small rdfs:label} and \texttt{\small plx:degradedText}, respectively.
+    Their original and degraded textual representations become the values of the properties \texttt{\small rdfs:label} and \texttt{\small plx:degradedText}, respectively.
     Their corresponding language variety is stated using \texttt{\small plx:languageVariety}.
@@
     This can be achieved with the denotation's \texttt{\small plx:wordClass} property pointing to a \texttt{\small plx:WordClass} entity.
-  \item All approvers share the \texttt{\small plx:Approver} class.
-    The characteristics of an approver are described using mainly triples with literal objects.
+  \item All sources share the \texttt{\small plx:Approver} class.
+    The characteristics of a source are described using mainly triples with literal objects.
     These are for example \texttt{\small dc:title} to assign the title of a source, \texttt{\small dc:creator} to give an \texttt{\small xsd:string} containing the author's name.
+At present, we support the different license categories recognized in the database by creating resources of the \texttt{\small plx:License} class.
 \end{itemize}
-Since the PanLex project compiled its database by extracting data from different sources, the licenses of these sources were also considered.
-At present, we support the different license categories given in the database by creating resources of the \texttt{\small plx:License} class.
 
@@
 \label{sec:conversion}
-Since new sources are added to the PanLex database on almost a daily basis and because of its current size (\textasciitilde18 GB) an approach of recurrent conversions is impractical.
+Since new sources are added to the PanLex database on almost a daily basis and because of its current size (\textasciitilde18 GB), the recurrent conversion of the database to capture changes in it is impractical.
 %Using conventional hardware, a full conversion takes infeasibly long.
@@
 % time-consuming.
-As the PanLex data already resides in a relational database, the use of a virtual RDB2RDF\footnote{\url{http://www.w3.org/2001/sw/wiki/RDB2RDF}} mapping solution is a natural choice.
+As the PanLex data already reside in a relational database, the use of a virtual RDB2RDF\footnote{\url{http://www.w3.org/2001/sw/wiki/RDB2RDF}} mapping solution is a natural choice.
 The \emph{Sparqlify system}\footnote{\url{https://github.com/AKSW/Sparqlify}} offers, besides an efficient query rewriting engine, also a very easy-to-use mapping language, called \emph{Sparqlification Mapping Language} (SML).
 Essentially, these mappings consist of three clauses:
-The \emph{From} clause specifies the logical SQL table (i.e. table, view or
+The \emph{From} clause specifies the logical SQL table (i.e. table, view, or
 query) to be used in the SML view.
@@
 \label{sec:related}
-PanLex is an integration project of many existing lexical resources.
+PanLex is a project whose editors integrate information discovered from many lexical resources.
 The extraction of information from linguistic sources, and techniques for automatically inferring translations,
@@
 % \emph{Global Wordnet Association}.
-An important initiative is the Global Wordnet Association\footnote{\url{http://www.globalwordnet.org/}\vspace{0.7em}},
-which offers a platform for sharing wordnets and defines several goals. 
-These include setting forth standards for uniformly representing wordnets
+An important related initiative is the Global Wordnet Association\footnote{\url{http://www.globalwordnet.org/}\vspace{0.7em}},
+which offers a platform for sharing wordnets and, among other goals, uniformly representing wordnets
 of different languages and establishing a universal index of meaning.
 Wordnets are usually focused on the
-definition of \emph{synsets} and relations between them in a single language,
-whereas PanLex's focus is on capturing lexical translations between languages.
-Hence, these efforts are complementary.
+definition of \emph{synsets} and relations between them in a single language; GWA is helping to transform these single-language synonym silos into a virtual multilingual translation resource.
+PanLex is approximating that, in a different way, by integrating data from numerous wordnets along with translingual sources into a single graph.
 %As already mentioned in PanLex there can be several meaning entities
 % representing the same unique sense.
-In the Semantic Web context, several (quasi-)standard vocabularies and ontologies have been developed with the rise of the \emph{Linguistic Linked Open Data} movement.
+In the Semantic Web context, several standard or quasi-standard vocabularies and ontologies have been developed with the rise of the \emph{Linguistic Linked Open Data} movement.
 Examples include the
@@
 Furthermore, we interlinked the languages in PanLex with Lexvo, and created about 2.5 million links to DBpedia for expressions in 16 languages.
-With the integration of lemon and GOLD we also support the data access via external linguistic ontologies.
+With the integration of lemon and GOLD we also support data access via external linguistic ontologies.
 
-There exist some shortcomings which we intend to overcome in the future:
-Modeling information sources and users as distinct entities would enable one
-to unequivocally relate meanings and denotations to them, which in turn would
-allow for a more fine grained attribution of qualities and relevances.
-This could be beneficial for translation approaches.
-Another important aspect is, that each of PanLex's information sources can be
-seen as a dataset on its own, and thus relations between them could be modeled
-with the VoID vocabulary\footnote{\url{http://rdfs.org/ns/void}}.
+We intend to address some limitations in the future.
+Attribution is currently limited to sources, obscuring the possible roles of multiple editors in iteratively interpreting sources and the subsidiary sources of data in sources that are themselves aggregations.
+Metadata attached to PanLex denotations are currently limited to arbitrary pairs of strings, but this sacrifices discovery possibilities when the metadata (as they often do) describe facts that can be expressed with PanLex expressions (e.g., that the expression is archaic).
+New collaborations between PanLex and related standards and technologies (e.g., language identification, language geolocation, lemmatization, transliteration, localization, etc.) are promising arenas for development.
+For example, the relations among PanLex's information sources, if treated as distinct datasets, could be modeled with the VoID vocabulary\footnote{\url{http://rdfs.org/ns/void}}.
 

### 2014-02-18 panlex-rdf-paper/2012/MLOD_PanLex/paper_v4.tex
replaced approver with source and shrunk occurences table

@@
   \centering
-  \includegraphics[width=0.4\textwidth]{images/schema-new.png}
+  \includegraphics[width=0.4\textwidth]{images/schema_new.png}
   \caption{The PanLex database schema}
@@
 An overview of the number of instances per entity as of January 2014 is given in~\autoref{fig:plx-entity-counts}.
+%\begin{table}
+%  \centering\begin{scriptsize}
+%  \begin{tabular}{lrclr}
+%    Entity             & Instances  \\
+%    \cline{1-2}
+%    Denotations        & 54,278,860 \\  % SELECT COUNT(dn) FROM dn;
+%    Languages          &      7,843 \\ % SELECT COUNT(lc) FROM lc; -- all language codes (ISO 639-3, ISO 639-2, ISO 639-5) known in the PanLex db
+%    Language Varieties identified           &      9,310 \\  % SELECT COUNT(lv) FROM lv;
+%    Language Varieties with data           &      9,239 \\  % 
+%    Meanings           & 20,773,371 \\  % SELECT COUNT(mn) FROM mn;
+%    Sources being consulted         &      4,190 \\  % 
+%    Sources already consulted         &      1,453 \\  % 
+%    Expressions        & 19,790,453 \\ % SELECT COUNT(ex) FROM ex;
+%    Definitions        &  2,747,892 \\ % SELECT COUNT(df) FROM df;
+%    Users              &          23 \\ % SELECT COUNT(us) FROM us;
+%  \end{tabular}
+%  \end{scriptsize}
+%  \caption{Number of instances of main entities in the PanLex database}
+%  \label{fig:plx-entity-counts}
+%\end{table}
+
+%    Users              &         23 \\ % SELECT COUNT(us) FROM us;
 \begin{table}
@@
   \begin{tabular}{lrclr}
-    Entity             & Instances  \\
-    \cline{1-2}
-    Denotations        & 54,278,860 \\  % SELECT COUNT(dn) FROM dn;
-    Languages          &      7,843 \\ % SELECT COUNT(lc) FROM lc; -- all language codes (ISO 639-3, ISO 639-2, ISO 639-5) known in the PanLex db
-    Language Varieties identified           &      9,310 \\  % SELECT COUNT(lv) FROM lv;
-    Language Varieties with data           &      9,239 \\  % 
-    Meanings           & 20,773,371 \\  % SELECT COUNT(mn) FROM mn;
-    Sources being consulted         &      4,190 \\  % 
-    Sources already consulted         &      1,453 \\  % 
-    Expressions        & 19,790,453 \\ % SELECT COUNT(ex) FROM ex;
-    Definitions        &  2,747,892 \\ % SELECT COUNT(df) FROM df;
-    Users              &          23 \\ % SELECT COUNT(us) FROM us;
+    Entity             & Instances  & & Entity    & Instances \\
+    \cline{1-2}\cline{4-5}
+    Denotations        & 54,278,860 & & Languages & 7,843     \\
+    Meanings           & 20,773,371 & & Sources   & 4,190     \\
+    Expressions        & 19,790,453 & & Users     &    23     \\
+    Definitions        &  2,747,892 & & Licenses  &    11     \\
+    Language Varieties &      9,310 & & & \\
   \end{tabular}
@@
   \label{fig:plx-entity-counts}
-\end{table}
+\end{table}\todo{new license `75 | pl | pup'; What does it mean?}
 
@@
     This can be achieved with the denotation's \texttt{\small plx:wordClass} property pointing to a \texttt{\small plx:WordClass} entity.
-  \item All sources share the \texttt{\small plx:Approver} class.
+  \item All sources share the \texttt{\small plx:Source} class.
     The characteristics of a source are described using mainly triples with literal objects.
@@
     \toprule
-    \texttt{plx:Approver} & \mbox{\texttt{plx:registrationDate}, \texttt{rdfs:label}, \texttt{dc:title},}
+    \texttt{plx:Source} & \mbox{\texttt{plx:registrationDate}, \texttt{rdfs:label}, \texttt{dc:title},}
                             \mbox{\texttt{dc:creator}, \texttt{plx:license}, \texttt{dc:date}, \texttt{plx:quality},}
@@
     \texttt{plx:Meaning}
-                          & \mbox{\texttt{plx:approver}, \texttt{plx:identifier},} \texttt{plx:meaningDefinition} \\
+                          & \mbox{\texttt{plx:source}, \texttt{plx:identifier},} \texttt{plx:meaningDefinition} \\
     \midrule

### 2014-02-18 panlex-rdf-paper/2012/MLOD_PanLex/paper_v4.tex
Minor update

--- 2014-02-18 panlex-rdf-paper/2012/MLOD_PanLex/paper_v4.tex
@@
 To accomplish that, the project is systematically integrating and making publicly available all known lexical translations.
-Semantic Web technologies can enhance the power of PanLex by flexibly representing and reasoning with the content of its database and interlinking it with linguistic and other resources and annotations.
+Semantic Web technologies can enhance the power of PanLex by flexibly representing and reasoning with the content of its database and interlinking it with
+%linguistic and
+other resources and annotations.
 %Using Semantic Web technologies can support achieving this goal, especially due
@@
 \label{sec:intro}
-At present, there are about 7,000 living languages in the world\footnote{\url{http://www-01.sil.org/iso639-3/iso-639-3.tab}.
-As of 25 January 2014, this list contained 6,975 living individual languages.}.
-Nonetheless, processes such as urbanization, nation-state consolidation, and globalization appear to be producing language attrition, with, for example, the extinction or near-extinction in the last 60 years of from 10\% to over 75\% (depending on the world's region) of all living languages\cite{lang_crisis}.
+At present, there are about 7,000 living languages in the world\footnote{\url{http://www-01.sil.org/iso639-3/iso-639-3.tab}}.
+%\footnote{\url{http://www-01.sil.org/iso639-3/iso-639-3.tab}. As of 25 January 2014, this list contained 6,975 living individual languages.}
+%Nonetheless, processes such as urbanization, nation-state consolidation, and globalization appear to be producing language attrition, with, %for example, the extinction or near-extinction in the last 60 years of from 10\% to over 75\% (depending on the world's region) of all living languages\cite{lang_crisis}.
+Nonetheless, processes such as urbanization, nation-state consolidation, and globalization appear to be producing language attrition: For example, in the last 60 years and depending on the world's region, from 10\% to over 75\% of all living languages have become extinct or near-extinct~\cite{lang_crisis}.
 Theorists of biolinguistic diversity argue that the loss of language diversity, the loss of human biological knowledge, and the loss of species diversity are mutually supportive and thus that language preservation and revitalization are essential to the preservation of biological diversity\cite{nettle}.
@@
 
+\todo{The combination of user + source document, which has approved any assertion, is the assertion’s source.}
 \begin{itemize}

### 2014-02-19 panlex-rdf-paper/2012/MLOD_PanLex/paper_v4.tex
Update

@@
 %Nonetheless, processes such as urbanization, nation-state consolidation, and globalization appear to be producing language attrition, with, %for example, the extinction or near-extinction in the last 60 years of from 10\% to over 75\% (depending on the world's region) of all living languages\cite{lang_crisis}.
-Nonetheless, processes such as urbanization, nation-state consolidation, and globalization appear to be producing language attrition: For example, in the last 60 years and depending on the world's region, from 10\% to over 75\% of all living languages have become extinct or near-extinct~\cite{lang_crisis}.
+Nonetheless, processes such as urbanization, nation-state consolidation, and globalization appear to be producing language attrition: For example, in the last 60 years and depending on the world's region, from 10\% to over 75\% of all living languages have become extinct or near-extinct~\cite{lang_crisis}.\todo{Jonathan: Please check statement}
 Theorists of biolinguistic diversity argue that the loss of language diversity, the loss of human biological knowledge, and the loss of species diversity are mutually supportive and thus that language preservation and revitalization are essential to the preservation of biological diversity\cite{nettle}.
 
-The mission of the PanLex project is to investigate the possibility that language attrition would be decreased or halted if utterances in any language were efficiently translatable into all other languages and, as a result, the speakers of all languages could use their languages for world-wide communication.
-Among the prerequisites of the panlingual translatablity of discourses would be the panlingual translatability of lexicons.
-In fact, lexical translation is almost the entirety of translation in some contexts (such as personal profiles, bibliographic subject headings, content tagging, product price lists, and web navigation controls).
+The mission of the PanLex project is to investigate the possibility that language attrition would be decreased or halted if utterances in any language were efficiently translatable into all other languages, and, as a result, the speakers of all languages could use their languages for world-wide communication.
+One prerequisite for panlingual translatablity of discourses is the panlingual translatability of lexicons.
+%In fact, lexical translation is almost the entirety of translation in some contexts (such as personal profiles, bibliographic subject headings, content tagging, product price lists, and web navigation controls).
 PanLex is designed to support that component by systematically integrating millions of lexical translations, documented in thousands of sources, into a database accessible to the research community, application developers, and the public.
-The database content can be interpreted as a graph linking millions of lexemes with one another in billions of ``is-a-translation-of'' relations, and permitting intelligent automated inference to trillions more such relations that have not been attested.
+%The database content can be interpreted as a graph linking millions of lexemes with one another in billions of ``is-a-translation-of'' relations, and permitting intelligent automated inference to trillions more such relations that have not been attested.
 %It documents the known lexical translations (translations of lexemes) among all
@@
 
-Precisely because the citation forms of lexemes, which is what PanLex directly enables the translation of, are commonly used as labels in collections of data, it is reasonable to speculate that PanLex might facilitate cross-linguistic access to knowledge.
-The idea of a Semantic Web has led to a stack of technologies standardized by the
+%Precisely because the citation forms of lexemes, which is what PanLex directly enables the translation of, are commonly used as labels in collections of data, it is reasonable to speculate that PanLex might facilitate cross-linguistic access to knowledge.
+
+The Semantic Web technologies
+standardized by the
 \emph{World Wide Web Consortium}\footnote{\url{http://www.w3.org/standards/semanticweb/}}
-supporting a machine-readable and machine-interpretable Linked Data
-network. This has led to the concept and reality of a \emph{Web of Data}.
+lay the foundations for the \emph{Web of Data}: an infrastructure for
+%machine-readable and
+machine-interpretable data on the Web.
+
+% the basis for the \emph{Web of Data}: a
+%: an infrastructure for
+%machine-readable
+%machine-interpretable data on the Web.
+
+%supporting a machine-readable and machine-interpretable Linked Data
+%network.
+%This has led to the concept and reality of a \emph{Web of Data}.
 %Technologies that support this aim are already available and in use in the so
 %called \emph{Web of Data}.
-Currently there is a growing community working on leveraging these Semantic Web
-technologies for linguistic knowledge and thereby building a Web of Linguistic
-Data, also known as the \emph{Linguistic Linked Open Data (LLOD) cloud}.
+Currently there is a growing community leveraging
+%Semantic Web
+these
+technologies for linguistic knowledge which is giving rise to the \emph{Linguistic Linked Open Data (LLOD) cloud}.
+
+% and thereby building a Web of Linguistic Data, also known as the \emph{Linguistic Linked Open Data (LLOD) cloud}.
 
@@
 Usage scenarios are given in \autoref{sec:usage}.
-In \autoref{sec:related} we discuss related work, and finally, in \autoref{sec:conclusion}, we conclude our approach and give some hints about future work.
+In \autoref{sec:related} we discuss related work, and finally, in \autoref{sec:conclusion}, we conclude our approach and outline future work.
 

### 2014-02-19 panlex-rdf-paper/2012/MLOD_PanLex/paper_v4.tex
changed texttt to emph and tried to compact table 3

@@
 %Nonetheless, processes such as urbanization, nation-state consolidation, and globalization appear to be producing language attrition, with, %for example, the extinction or near-extinction in the last 60 years of from 10\% to over 75\% (depending on the world's region) of all living languages\cite{lang_crisis}.
-Nonetheless, processes such as urbanization, nation-state consolidation, and globalization appear to be producing language attrition: For example, in the last 60 years and depending on the world's region, from 10\% to over 75\% of all living languages have become extinct or near-extinct~\cite{lang_crisis}.\todo{Jonathan: Please check statement}
-Theorists of biolinguistic diversity argue that the loss of language diversity, the loss of human biological knowledge, and the loss of species diversity are mutually supportive and thus that language preservation and revitalization are essential to the preservation of biological diversity\cite{nettle}.
+Nonetheless, processes such as urbanization, nation-state consolidation, and globalization appear to be producing language attrition: For example, in the last 60 years and depending on the world's region, from 10\% to over 75\% of all living languages have become extinct or near-extinct~\cite{lang_crisis}.
+Theorists of biolinguistic diversity argue that the loss of language diversity, the loss of human biological knowledge, and the loss of species diversity are mutually supportive and thus that language preservation and revitalization are essential to the preservation of biological diversity~\cite{nettle}.
 
@@
 Far from being mere copies of the sources, which would not make PanLex data interoperable, the data constitute editors' \emph{interpretations} of various sources' \emph{assertions} that two or more \emph{expressions} share a \emph{meaning}\footnote{\url{http://panlex.org/tech/doc/design/panlex-db-design.pdf}}.
-As of January 2014, the database contained about 20 million meanings and 20 million expressions in about 9,200 language varieties based on the consultation of about 1,500 sources by about 20 editors.
+% PW: I commented the following sentence out since I removed the detailed number of instances table (language varieties identified/with data; source being consulted/already consulted.
+I used the bigger numbers and in this sentence the smaller numbers are used, which might confuse
+%As of January 2014, the database contained about 20 million meanings and 20 million expressions in about 9,200 language varieties based on the consultation of about 1,500 sources by about 20 editors.
 The most important entities and relations of PanLex's conceptual model are depicted in \autoref{fig:db-schema} and are explained in more detail below.
@@
   \label{fig:plx-entity-counts}
-\end{table}\todo{new license `75 | pl | pup'; What does it mean?}
+\end{table}
 
@@
 The entities and relations of the schema described in the previous section serve as the base for the development of the PanLex RDF vocabulary.
-In general, all PanLex RDF resources reside in the namespace \texttt{\small <http://ld.panlex.org/plx/>}, abbreviated with \texttt{\small plx}.
+In general, all PanLex RDF resources reside in the namespace \emph{<http://ld.panlex.org/plx/>}, abbreviated with \emph{plx}.
 An example of the resulting ontology is depicted in \autoref{fig:vocabulary} and summarized as follows.
@@
 \begin{itemize}
-  \item Expressions are modeled as instances of the class \texttt{\small plx:Expression}.
-    Their original and degraded textual representations become the values of the properties \texttt{\small rdfs:label} and \texttt{\small plx:degradedText}, respectively.
-    Their corresponding language variety is stated using \texttt{\small plx:languageVariety}.
-  \item For language and language varieties the classes \texttt{\small plx:Language} and \texttt{\small plx:LanguageVariety} are introduced.
-    \emph{ISO 639-1} and \emph{ISO 639-3 codes} become instances of the classes \texttt{\small plx:Iso639-1Code} and \texttt{\small plx:Iso639-3Code}.
-  \item The RDF analog of the PanLex \emph{meaning} is the \texttt{\small plx:Meaning}.
-    Entities of this class may have an identifier assigned with the \texttt{\small plx:identifier} property pointing to an \texttt{\small xsd:string} literal.
-    Meanings may also have \emph{definitions}, entities of the \texttt{\small plx:Definition} class, giving a textual representation (\texttt{\small rdfs:label}) in a certain language variety (\texttt{\small plx:languageVariety}).
+  \item Expressions are modeled as instances of the class \emph{plx:Expression}.
+    Their original and degraded textual representations become the values of the properties \emph{rdfs:label} and \emph{plx:degradedText}, respectively.
+    Their corresponding language variety is stated using \emph{plx:languageVariety}.
+  \item For language and language varieties the classes \emph{plx:Language} and \emph{plx:LanguageVariety} are introduced.
+    ISO 639-1 and ISO 639-3 codes become instances of the classes \emph{plx:Iso639-1Code} and \emph{plx:Iso639-3Code}.
+  \item The RDF analog of the PanLex \emph{meaning} is the \emph{plx:Meaning}.
+    Entities of this class may have an identifier assigned with the \emph{plx:identifier} property pointing to an \emph{xsd:string} literal.
+    Meanings may also have \emph{definitions}, entities of the \emph{plx:Definition} class, giving a textual representation (\emph{rdfs:label}) in a certain language variety (\emph{plx:languageVariety}).
   \item Following the semantics of the PanLex database, meanings and expressions are linked via \emph{denotations}.
-    These are entities of the \texttt{\small plx:Denotation} class pointing to meanings and expressions via the properties \texttt{\small plx:denotationMeaning} and \\ \texttt{\small plx:denotationExpression}.
+    These are entities of the \emph{plx:Denotation} class pointing to meanings and expressions via the properties \emph{plx:denotationMeaning} and \emph{plx:\-denotation\-Expression}.
     Denotations may also have a word class assigned to them.
-    This can be achieved with the denotation's \texttt{\small plx:wordClass} property pointing to a \texttt{\small plx:WordClass} entity.
-  \item All sources share the \texttt{\small plx:Source} class.
+    This can be achieved with the denotation's \emph{plx:wordClass} property pointing to a \emph{plx:WordClass} entity.
+  \item All sources share the \emph{plx:Source} class.
     The characteristics of a source are described using mainly triples with literal objects.
-    These are for example \texttt{\small dc:title} to assign the title of a source, \texttt{\small dc:creator} to give an \texttt{\small xsd:string} containing the author's name.
-At present, we support the different license categories recognized in the database by creating resources of the \texttt{\small plx:License} class.
+    These are for example \emph{dc:title} to assign the title of a source, \emph{dc:creator} to give an \emph{xsd:string} containing the author's name.
+    At present, we support the different license categories recognized in the database by creating resources of the \emph{plx:License} class.
 \end{itemize}
@@
 \begin{table}
-  \begin{tiny}
-  \begin{tabular}{p{52px}p{140px}}
+  \centering
+  \begin{scriptsize}
+  \fontsize{5.8pt}{5pt}\selectfont
+  \begin{tabular}{p{41pt}p{150pt}}
     Class                 & Properties \\
     \toprule
-    \texttt{plx:Source} & \mbox{\texttt{plx:registrationDate}, \texttt{rdfs:label}, \texttt{dc:title},}
-                            \mbox{\texttt{dc:creator}, \texttt{plx:license}, \texttt{dc:date}, \texttt{plx:quality},}
-                            \mbox{\texttt{foaf:homepage}, \texttt{dc:publisher}, \texttt{dbpedia-owl:isbn}} \\
+%    \emph{plx:Source} & \emph{plx:registrationDate},\,\emph{rdfs:label},\,\emph{dc:title},\,\emph{dc:creator},\,\emph{dc:quality}, \emph{plx:license},\,\emph{dc:date},\,\emph{foaf:homepage},\,\emph{dc:publisher},\,\emph{dbpedia-owl:isbn} \\
+    \emph{plx:Source} & \emph{plx:registrationDate},\,\emph{rdfs:label},\,\emph{dc:title},\,\emph{dc:publisher},\,\emph{dc:quality}, \emph{plx:license},\,\emph{dc:date},\,\emph{foaf:homepage},\,\emph{dc:creator},\,\mbox{\emph{dbpedia-owl:isbn}} \\
     \midrule
-    \texttt{plx:Language} & \texttt{plx:iso639-3Code}, \texttt{plx:iso639-1Code} \\
+    \emph{plx:Language} & \emph{plx:iso639-3Code}, \emph{plx:iso639-1Code} \\
     \midrule
-    \texttt{plx:LanguageVariety}
-                          & \texttt{plx:languageVarietyOf}, \texttt{rdfs:label} \\
+    \emph{plx:LanguageVariety}
+                          & \emph{plx:languageVarietyOf}, \emph{rdfs:label} \\
     \midrule
-    \texttt{plx:Iso639-1Code} & \\
+    \mbox{\emph{plx:Iso639-1Code}} & \\
     \midrule
-    \texttt{plx:Iso639-3Code} & \\
+    \mbox{\emph{plx:Iso639-3Code}} & \\
     \midrule
-    \texttt{plx:Expression}
-                          & \mbox{\texttt{plx:languageVariety}, \texttt{plx:degradedText},} \texttt{rdfs:label} \\
+    \emph{plx:Expression}
+                          & \emph{plx:languageVariety}, \emph{plx:degradedText}, \emph{rdfs:label} \\
     \midrule
-    \texttt{plx:Meaning}
-                          & \mbox{\texttt{plx:source}, \texttt{plx:identifier},} \texttt{plx:meaningDefinition} \\
+    \emph{plx:Meaning}
+                          & \emph{plx:source}, \emph{plx:identifier}, \emph{plx:meaningDefinition} \\
     \midrule
-    \texttt{plx:Definition}
-                          & \texttt{plx:languageVariety}, \texttt{rdfs:label} \\
+    \emph{plx:Definition}
+                          & \emph{plx:languageVariety}, \emph{rdfs:label} \\
     \midrule
-    \texttt{plx:Denotation}
-                          & \texttt{plx:denotationMeaning}, \texttt{plx:denotationExpression}, \texttt{plx:wordClass} \\
+    \emph{plx:Denotation}
+                          & \emph{plx:denotationMeaning},\,\emph{plx:denotationExpression},\,\emph{plx:wordClass} \\
     \midrule
-    \texttt{plx:WordClass}  
-                          & \texttt{rdfs:label} \\
+    \emph{plx:WordClass}  
+                          & \emph{rdfs:label} \\
     \midrule
-    \texttt{plx:License}  & \texttt{rdfs:label} \\
-    \bottomrule
+    \emph{plx:License}    & \emph{rdfs:label} \\
+    \bottomrule\vspace{3pt}
   \end{tabular}
-  \end{tiny}
-  \caption{Classes and properties used in the PanLex RDF vocabulary. Note that all \texttt{rdf:type} properties are omitted for brevity.}
+  \end{scriptsize}
+  \caption{Classes and properties used in the PanLex RDF vocabulary. Note that all \emph{rdf:type} properties are omitted for brevity.}
   \label{tbl:vocabulary}
@@
   \centering\begin{scriptsize}
+  \fontsize{6.5pt}{6.4pt}\selectfont
   \begin{tabular}{p{48px}p{62px}p{70px}}
@@
     \midrule
-    \texttt{plx:Denotation} & --                          & \texttt{gold:LinguisticSign} \\
-    \texttt{plx:Meaning}    & \texttt{lemon:LexicalSense} & \texttt{gold:SemanticUnit} \\
-    \texttt{plx:Expression} & \texttt{lemon:LexicalEntry} & \texttt{gold:FormUnit} \\
+    \emph{plx:Denotation} & --                          & \emph{gold:LinguisticSign} \\
+    \emph{plx:Meaning}    & \emph{lemon:LexicalSense} & \emph{gold:SemanticUnit} \\
+    \emph{plx:Expression} & \emph{lemon:LexicalEntry} & \emph{gold:FormUnit} \\
   \end{tabular}
@@
 %Apart from modeling a PanLex RDF vocabulary
-Additionally, we also re-use existing vocuabularies, namely
+Additionally, we also re-use existing vocabularies, namely
 the \emph{Lexicon Model for Ontologies} (lemon)~\cite{lemon2011} as well as the \emph{General Ontology for Linguistic Description} (GOLD)~\cite{farr2003}.
@@
 endpoints\footnote{\url{http://ld.panlex.org/vsparql}}
-\footnote{\url{http://ld.panlex.org/sparql}}, powered by Sparqlify and Virtuoso,
+\footnote{\url{http://ld.panlex.org/sparql}}, ran by Sparqlify and Virtuoso,
 respectively.
@@
 % \emph{Global Wordnet Association}.
-An important related initiative is the Global Wordnet Association\footnote{\url{http://www.globalwordnet.org/}\vspace{0.7em}},
+An important related initiative is the Global Wordnet Association (GWA)\footnote{\url{http://www.globalwordnet.org/}},
 which offers a platform for sharing wordnets and, among other goals, uniformly representing wordnets

### 2014-02-24 panlex-rdf-paper/2012/MLOD_PanLex/paper_v5.tex
added v5 of the paper tex file

--- 2014-02-18 panlex-rdf-paper/2012/MLOD_PanLex/paper_v4.tex
@@
 \begin{abstract}
-At present, there are approximately 7,000 living languages in the world.
-However, some forecasters expect the world to lose most of this linguistic diversity.
-The mission of the PanLex project is to investigate whether it is feasible to make all languages intertranslatable and thus usable for world-wide communication, and, if so, whether that new situation will interfere with this expected language attrition.
-This mission entails, among other prerequisites, making the lexicons of the world's languages intertranslatable.
-To accomplish that, the project is systematically integrating and making publicly available all known lexical translations.
-Semantic Web technologies can enhance the power of PanLex by flexibly representing and reasoning with the content of its database and interlinking it with
-%linguistic and
-other resources and annotations.
+The world is losing some of its 7,000 languages.
+Hypothesizing that language attrition might subside if all languages were inter\-trans\-latable, the PanLex project supports panlingual lexical translation by integrating all known lexical translations.
+Semantic Web technologies can flexibly represent and reason with the content of its database and interlink it with linguistic and other resources and annotations.
 %Using Semantic Web technologies can support achieving this goal, especially due
 % to inference capabilities and the interlinking of the PanLex data with other data sources.
-Conversely, PanLex, with its unique collection of translation links between more than a billion pairs of lexemes from more than 9,000 language varieties, can improve the coverage of the Linguistic Web of Data.
-In this dataset description paper we detail how we transformed the content of the PanLex database to RDF,
-established conformance with the lemon and GOLD data models,
-interlinked it with Lexvo and DBpedia, and published it as Linked Data and via
-SPARQL.
+Conversely, PanLex, with its collection of translation links between more than a billion pairs of lexemes from more than 9,000 language varieties, can improve the coverage of the Linguistic Web of Data.
+We detail how we transformed the content of the PanLex database to RDF, established conformance with the lemon and GOLD data models,
+interlinked it with Lexvo and DBpedia, and published it as Linked Data and via SPARQL.
 \end{abstract}
@@
 \label{sec:intro}
-At present, there are about 7,000 living languages in the world\footnote{\url{http://www-01.sil.org/iso639-3/iso-639-3.tab}}.
-%\footnote{\url{http://www-01.sil.org/iso639-3/iso-639-3.tab}. As of 25 January 2014, this list contained 6,975 living individual languages.}
-%Nonetheless, processes such as urbanization, nation-state consolidation, and globalization appear to be producing language attrition, with, %for example, the extinction or near-extinction in the last 60 years of from 10\% to over 75\% (depending on the world's region) of all living languages\cite{lang_crisis}.
-Nonetheless, processes such as urbanization, nation-state consolidation, and globalization appear to be producing language attrition: For example, in the last 60 years and depending on the world's region, from 10\% to over 75\% of all living languages have become extinct or near-extinct~\cite{lang_crisis}.
-Theorists of biolinguistic diversity argue that the loss of language diversity, the loss of human biological knowledge, and the loss of species diversity are mutually supportive and thus that language preservation and revitalization are essential to the preservation of biological diversity\cite{nettle}.
+There are about 7,000 living languages\footnote{\url{http://www-01.sil.org/iso639-3/iso-639-3.tab}}, but language attrition has extinguished or threatened from 10\% to over 75\% of all languages in the last 60 years in various regions\cite{lang_crisis}.
+This attrition arguably imperils human biological knowledge and species diversity\cite{nettle}.
+Hypothetically, panlingual intertranslatability would make all languages more useful and incentivize their preservation and revitalization.
 
-The mission of the PanLex project is to investigate the possibility that language attrition would be decreased or halted if utterances in any language were efficiently translatable into all other languages and, as a result, the speakers of all languages could use their languages for world-wide communication.
-Among the prerequisites of the panlingual translatablity of discourses would be the panlingual translatability of lexicons.
-In fact, lexical translation is almost the entirety of translation in some contexts (such as personal profiles, bibliographic subject headings, content tagging, product price lists, and web navigation controls).
-PanLex is designed to support that component by systematically integrating millions of lexical translations, documented in thousands of sources, into a database accessible to the research community, application developers, and the public.
-The database content can be interpreted as a graph linking millions of lexemes with one another in billions of ``is-a-translation-of'' relations, and permitting intelligent automated inference to trillions more such relations that have not been attested.
+The PanLex project is making all languages' lexicons intertranslatable.
+In some contexts (e.g., profiles, catalogs, tags, search, and web navigation), lexical translation can be most of the translation load.
+PanLex systematically integrates lexical translations, found in diverse sources, into a database for research, applications, and public use.
+The content can be interpreted as a graph linking lexemes in ``is-a-translation-of'' relations, and permitting automated inference to additional, unattested relations.
 %It documents the known lexical translations (translations of lexemes) among all
@@
 
-Precisely because the citation forms of lexemes, which is what PanLex directly enables the translation of, are commonly used as labels in collections of data, it is reasonable to speculate that PanLex might facilitate cross-linguistic access to knowledge.
-The idea of a Semantic Web has led to a stack of technologies standardized by the
+The Semantic Web Initiative has led to a stack of technologies standardized by the
 \emph{World Wide Web Consortium}\footnote{\url{http://www.w3.org/standards/semanticweb/}}
-supporting a machine-readable and machine-interpretable Linked Data
-network. This has led to the concept and reality of a \emph{Web of Data}.
+supporting a machine-readable and \mbox{-i}nter\-pretable Linked Data
+network. This has led to a \emph{Web of Data}.
 %Technologies that support this aim are already available and in use in the so
 %called \emph{Web of Data}.
-Currently there is a growing community working on leveraging these Semantic Web
-technologies for linguistic knowledge and thereby building a Web of Linguistic
-Data, also known as the \emph{Linguistic Linked Open Data (LLOD) cloud}.
+A growing community is leveraging Semantic Web technologies for linguistic knowledge, building a Web of Linguistic
+Data, or a \emph{Linguistic Linked Open Data (LLOD) cloud}.
 
-Our steps to connect PanLex to this Linked Data network are as follows.
-In \autoref{sec:triplificate} we introduce the PanLex dataset, present our
+When linked open data are labeled in any language with lexemes in citation form (lemmas), the data are potentially linkable to, and through, PanLex.
+Here we describe how we connected PanLex to the Linked Data network.
+In \autoref{sec:triplificate} we introduce the dataset, present a
 PanLex RDF vocabulary, and explain how we transformed the one into the other and established conformance with additional data models.
@@
 Usage scenarios are given in \autoref{sec:usage}.
-In \autoref{sec:related} we discuss related work, and finally, in \autoref{sec:conclusion}, we conclude our approach and give some hints about future work.
+In \autoref{sec:related} we discuss related work, and, finally, in \autoref{sec:conclusion}, we summarize our recent and contemplated work.
 
@@
 \label{sec:triplificate}
-In this section, we first provide an analysis of the PanLex dataset.
-Subsequently, we introduce our URI and vocabulary design, which closely
-resembles PanLex's original conceptual model.
+Here we analyze the PanLex dataset; introduce our URI and vocabulary design, resembling PanLex's conceptual model;
 %, which was developed independently from other ontology engineering approaches.
-Afterwards, we briefly describe how we classified PanLex's
-instance data using additional data models. Finally, we explain the steps
-taken to transform the data to RDF.
+summarize how we classified PanLex's instance data with additional data models; and explain our transformation of the data to RDF.
 
@@
 \label{sec:analysis}
-At the core of the PanLex \emph{project}, there is the PanLex \emph{database}, which is created by editors who consult thousands of lexical sources, such as mono- and multilingual dictionaries, glossaries, standards, and thesauri.
-The list of sources that have been and are being consulted is available online\footnote{\url{http://panlex.org/tech/plrefs.shtml}}.
-The data based on these sources include single- and multi-word expressions, meanings assigned to them, and related information.
-Far from being mere copies of the sources, which would not make PanLex data interoperable, the data constitute editors' \emph{interpretations} of various sources' \emph{assertions} that two or more \emph{expressions} share a \emph{meaning}\footnote{\url{http://panlex.org/tech/doc/design/panlex-db-design.pdf}}.
-As of January 2014, the database contained about 20 million meanings and 20 million expressions in about 9,200 language varieties based on the consultation of about 1,500 sources by about 20 editors.
-The most important entities and relations of PanLex's conceptual model are depicted in \autoref{fig:db-schema} and are explained in more detail below.
+The PanLex \emph{database} is created by editors who consult sources\footnote{\url{http://panlex.org/tech/plrefs.shtml}}, such as mono- and multilingual dictionaries, glossaries, standards, and thesauri.
+The data include single- and multi-word expressions, meanings assigned to them, and related information.
+While copies of the sources would not be interoperable, PanLex data constitute editors' \emph{interpretations} of sources' \emph{assertions} that two or more \emph{expressions} share a \emph{meaning}\footnote{\url{http://panlex.org/tech/doc/design/panlex-db-design.pdf}}.
+The most important entities and relations of PanLex's conceptual model are depicted in \autoref{fig:db-schema}.
 
@@
   \centering
-  \includegraphics[width=0.4\textwidth]{images/schema_new.png}
+  \includegraphics[width=0.4\textwidth]{images/schema-new.png}
   \caption{The PanLex database schema}
@@
 
-\todo{The combination of user + source document, which has approved any assertion, is the assertion’s source.}
 \begin{itemize}
-  \item The starting point of data acquisition is the \emph{source} entity: An editor consults some mono- or multilingual source as mentioned above, interprets it as making assertions about lexical translations, and adds schema-compliant representations of these assertions to the PanLex dataset.
-  \item \emph{Expressions} are lexical entities, each defined as a string of characters in some variety of some language.
-As a first approximation, expressions can be considered lemmas, i.e.
-dictionary-entry headwords.
-They differ from traditional headwords, however, in at least two ways.
+  \item The \emph{source} entity is the authority to which an editor attributes assertions about lexical translations.
+  \item \emph{Expressions} are lexical entities, each uniquely identified with a text, i.e.
+a string of (Unicode) characters, and a variety of a language.
+Expressions resemble lemmas or dictionary-entry headwords, but differ from them in at least two ways.
 (1) Homographs, such as the verb ``hide'' (conceal) and the noun ``hide'' (animal skin) in English, are treated as a single expression in PanLex.
 (2) Multiword expressions, such as ``fall in love'', traditionally found in an entry headed by one of their words, such as ``fall'' or ``love'', are treated as independent expressions in PanLex.
-Expressions are strings of Unicode codepoints subjected to Normalization Form NFC, but this still makes it difficult to search precisely for expressions, so to assist in approximate searching PanLex also adds for each expression a degraded form permitting only lower-case letters and digits.
+Expression texts are constrained to comply with Unicode Normalization Form NFC.
     \item \emph{Languages} in PanLex are identified using ISO 639-3\footnote{\url{http://www.sil.org/iso639-3/codes.asp}} individual and macrolanguage codes, ISO 639-2\footnote{\url{http://www.loc.gov/standards/iso639-2/}} collective codes, and ISO 639-5\footnote{\url{http://www.loc.gov/standards/iso639-5/}} codes.
-    \item \emph {Language varieties} are the fundamental linguistic entities in PanLex.
-Expressions are in them, not in languages.
-The language-variety concept is fine-grained and permissive: there can be many varieties of a language, and editors can add varieties as needed.
-A language variety is identified with some language code, combined with an arbitrary integer differentiating it from other varieties of the same language.
+    \item \emph {Language varieties} are collections of expressions.
+Each has a unique identifier: a language code and a distinguishing integer.
 For example, six dialects of Ahtna are identified as ``aht-000'' through ``aht-005''.
-These symbolic 3-letter, 3-digit labels are, themselves, treated as a (controlled) language variety, so editors can clarify their referents by translating them into natural-language expressions and into expressions in other controlled language varieties, such as the IETF standard BCP 47\footnote{\url{http://tools.ietf.org/html/bcp47}}.
-Such controlled language varieties may be defined as prohibiting synonymy and/or ambiguity.
-    \item \emph{Meanings} in PanLex are entities with which expressions are semantically classified.
-An editor documents that two or more expressions are asserted to be translations or synonyms of one another by assigning a shared meaning to them.
-For example, depending on an editor's judgment, a source's translation of the German lexeme ``klingen'' into a sequence of alternative English lexemes ``ring, sound, seem'' can be documented by either (1) the assignment of all four expressions to the same meaning or (2) the assignment of two or three meanings to ``klingen'' and of one of those meanings to each of the English expressions.
-For practicality, PanLex treats expressions and meanings asymmetrically.
-Newly added translations automatically re-use existing matching expressions, but they don't re-use existing meanings, because there is no straightforward test of equivalence for meanings.
-PanLex generates new meanings as needed, and meanings are source-specific, leaving it for subsequent research to investigate which meanings should be considered equivalent and thus consolidated.
-Meanings can further have properties of three types.
-(1) \emph{Definitions} are descriptions of a meaning, consisting of arbitrary text strings annotated as being in particular language varieties.
-(2) \emph{Domains} are expressions (e.g., ``medicine'') that classify a meaning, but do not express it.
-(3) \emph{Meaning identifiers} are strings acting as references to identifiers used in a source.
+These labels are, themselves, treated as a (controlled) language variety, whose expressions (i.e.
+the labels) are translated into natural languages and other controlled languages (such as the IETF standard BCP 47\footnote{\url{http://tools.ietf.org/html/bcp47}}.
+The only properties of this entity are ambiguity and synonymy: Controlled language varieties are usually defined to prohibit both of these.
+    \item \emph{Meanings} are the entities assigned to expressions, thereby identifying expressions as translations or synonyms.
+For example, a source's translation of the German expression ``klingen'' into English as ``ring, sound, seem'' can be interpreted as (1) the assignment of a single meaning to all four expressions or (2) the assignment of two or three meanings to ``klingen'' and of one of those to each of the English expressions.
+Meanings are source-specific.
+The identification and consolidation of equivalent meanings of distinct sources is a research topic, not a database feature.
+Meanings can have properties of three types.
+(1) \emph{Definitions} are descriptions of a meaning, consisting of text strings annotated as being in particular language varieties.
+(2) \emph{Domains} are expressions (e.g., ``medicine'') that characterize a meaning, but do not express it.
+(3) \emph{Meaning identifiers} are strings acting as references to identifiers in a source.
     \item \emph{Denotations} are assignments of meanings to expressions.
-Each denotation, in addition to its meaning-expression pair, may be assigned one or more \emph{word classes} (a closed set based on OLIF, the Open Lexicon Interchange Format) and/or \emph{metadata} (arbitrary strings paired as keys and values).
+A denotation may have one or more \emph{word classes} (a closed set based on OLIF, the Open Lexicon Interchange Format) and/or \emph{metadata} (arbitrary strings paired as keys and values).
     %For example, \emph{fall} can be a verb or a noun for autumn.
     %Homonyms are those expressions that are connected to multiple meanings.
-  \item \emph{Users} are the entities that may act as editors by defining sources and then defining data attributed to those sources, as well as defining language varieties.
-  \item Sources in PanLex are defined with about two dozen properties, among them being \emph{licenses}.
-As of January 2014 eleven different license categories were recognized.
-They were \emph{public domain}, \emph{Creative Commons (CC)}, \emph{request} (meaning that one has to ask the author of the resource), \emph{GNU General Public License (GPL)}, \emph{GNU Lesser General Public License (LGPL)}, \emph{GNU Free Documentation License (FDL)}, \emph{MIT License}, \emph{copyright} (meaning that there is a copyright claim without a more liberal license also being offered), \emph{PanLex Use Permission} (where an author or publisher has given a source to PanLex for its use), \emph{other}, and \emph{unknown}.
-    The distribution of licenses among sources that had been or were being consulted in January 2014 is shown in \autoref{tbl:plx-license-counts}.
+  \item \emph{Users} may define sources, attribute data to them, and define language varieties.
+  \item Among the properties of sources are \emph{licenses}.
+Some of the license categories are \emph{public domain}, \emph{Creative Commons (CC)}, \emph{request} (author invites inquiries), \emph{GNU General Public License (GPL)}, \emph{GNU Lesser General Public License (LGPL)}, \emph{GNU Free Documentation License (FDL)}, \emph{MIT License}, \emph{copyright} (without a more liberal license), and \emph{PanLex Use Permission} (specific permission for use in PanLex).
+    The distribution of licenses as of January 2014 is shown in \autoref{tbl:plx-license-counts}.
 \end{itemize}
@@
 
-An overview of the number of instances per entity as of January 2014 is given in~\autoref{fig:plx-entity-counts}.
-%\begin{table}
-%  \centering\begin{scriptsize}
-%  \begin{tabular}{lrclr}
-%    Entity             & Instances  \\
-%    \cline{1-2}
-%    Denotations        & 54,278,860 \\  % SELECT COUNT(dn) FROM dn;
-%    Languages          &      7,843 \\ % SELECT COUNT(lc) FROM lc; -- all language codes (ISO 639-3, ISO 639-2, ISO 639-5) known in the PanLex db
-%    Language Varieties identified           &      9,310 \\  % SELECT COUNT(lv) FROM lv;
-%    Language Varieties with data           &      9,239 \\  % 
-%    Meanings           & 20,773,371 \\  % SELECT COUNT(mn) FROM mn;
-%    Sources being consulted         &      4,190 \\  % 
-%    Sources already consulted         &      1,453 \\  % 
-%    Expressions        & 19,790,453 \\ % SELECT COUNT(ex) FROM ex;
-%    Definitions        &  2,747,892 \\ % SELECT COUNT(df) FROM df;
-%    Users              &          23 \\ % SELECT COUNT(us) FROM us;
-%  \end{tabular}
-%  \end{scriptsize}
-%  \caption{Number of instances of main entities in the PanLex database}
-%  \label{fig:plx-entity-counts}
-%\end{table}
-
-%    Users              &         23 \\ % SELECT COUNT(us) FROM us;
+\autoref{fig:plx-entity-counts} gives entity counts as of January 2014.
 \begin{table}
@@
   \begin{tabular}{lrclr}
-    Entity             & Instances  & & Entity    & Instances \\
-    \cline{1-2}\cline{4-5}
-    Denotations        & 54,278,860 & & Languages & 7,843     \\
-    Meanings           & 20,773,371 & & Sources   & 4,190     \\
-    Expressions        & 19,790,453 & & Users     &    23     \\
-    Definitions        &  2,747,892 & & Licenses  &    11     \\
-    Language Varieties &      9,310 & & & \\
+    Entity             & Instances  \\
+    \cline{1-2}
+    Denotations        & 54,278,860 \\  % SELECT COUNT(dn) FROM dn;
+    Languages          &      7,843 \\ % SELECT COUNT(lc) FROM lc; -- all language codes (ISO 639-3, ISO 639-2, ISO 639-5) known in the PanLex db
+    Language Varieties identified           &      9,310 \\  % SELECT COUNT(lv) FROM lv;
+    Language Varieties with data           &      9,239 \\  % 
+    Meanings           & 20,773,371 \\  % SELECT COUNT(mn) FROM mn;
+    Sources being consulted         &      4,190 \\  % 
+    Sources already consulted         &      1,453 \\  % 
+    Expressions        & 19,790,453 \\ % SELECT COUNT(ex) FROM ex;
+    Definitions        &  2,747,892 \\ % SELECT COUNT(df) FROM df;
+    Users              &          23 \\ % SELECT COUNT(us) FROM us;
   \end{tabular}
@@
 \label{sec:vocabulary}
-The entities and relations of the schema described in the previous section serve as the base for the development of the PanLex RDF vocabulary.
+The entities and relations described above are the base for the PanLex RDF vocabulary.
 In general, all PanLex RDF resources reside in the namespace \texttt{\small <http://ld.panlex.org/plx/>}, abbreviated with \texttt{\small plx}.
@@
     This can be achieved with the denotation's \texttt{\small plx:wordClass} property pointing to a \texttt{\small plx:WordClass} entity.
-  \item All sources share the \texttt{\small plx:Source} class.
+  \item All sources share the \texttt{\small plx:Approver} class.
     The characteristics of a source are described using mainly triples with literal objects.
@@
     \toprule
-    \texttt{plx:Source} & \mbox{\texttt{plx:registrationDate}, \texttt{rdfs:label}, \texttt{dc:title},}
+    \texttt{plx:Approver} & \mbox{\texttt{plx:registrationDate}, \texttt{rdfs:label}, \texttt{dc:title},}
                             \mbox{\texttt{dc:creator}, \texttt{plx:license}, \texttt{dc:date}, \texttt{plx:quality},}
@@
     \texttt{plx:Meaning}
-                          & \mbox{\texttt{plx:source}, \texttt{plx:identifier},} \texttt{plx:meaningDefinition} \\
+                          & \mbox{\texttt{plx:approver}, \texttt{plx:identifier},} \texttt{plx:meaningDefinition} \\
     \midrule

### 2014-02-24 panlex-rdf-paper/2012/MLOD_PanLex/paper_v5.tex
minor changes
Added some of the changes made in v4, that were not considered during the revision
to v5

@@
 \label{sec:intro}
-There are about 7,000 living languages\footnote{\url{http://www-01.sil.org/iso639-3/iso-639-3.tab}}, but language attrition has extinguished or threatened from 10\% to over 75\% of all languages in the last 60 years in various regions\cite{lang_crisis}.
-This attrition arguably imperils human biological knowledge and species diversity\cite{nettle}.
+There are about 7,000 living languages\footnote{\url{http://www-01.sil.org/iso639-3/iso-639-3.tab}}, but language attrition has extinguished or threatened from 10\% to over 75\% of all languages in the last 60 years in various regions~\cite{lang_crisis}.
+This attrition arguably imperils human biological knowledge and species diversity~\cite{nettle}.
 Hypothetically, panlingual intertranslatability would make all languages more useful and incentivize their preservation and revitalization.
@@
   \centering
-  \includegraphics[width=0.4\textwidth]{images/schema-new.png}
+  \includegraphics[width=0.42\textwidth]{images/schema_new.png}
   \caption{The PanLex database schema}
@@
 These labels are, themselves, treated as a (controlled) language variety, whose expressions (i.e.
-the labels) are translated into natural languages and other controlled languages (such as the IETF standard BCP 47\footnote{\url{http://tools.ietf.org/html/bcp47}}.
-The only properties of this entity are ambiguity and synonymy: Controlled language varieties are usually defined to prohibit both of these.
+the labels) are translated into natural languages and other controlled languages (such as the IETF standard BCP 47\footnote{\url{http://tools.ietf.org/html/bcp47}}).
+%The only properties of this entity are ambiguity and synonymy: Controlled language varieties are usually defined to prohibit both of these.
     \item \emph{Meanings} are the entities assigned to expressions, thereby identifying expressions as translations or synonyms.
@@
     Denotations        & 54,278,860 \\  % SELECT COUNT(dn) FROM dn;
-    Languages          &      7,843 \\ % SELECT COUNT(lc) FROM lc; -- all language codes (ISO 639-3, ISO 639-2, ISO 639-5) known in the PanLex db
+    Meanings           & 20,773,371 \\  % SELECT COUNT(mn) FROM mn;
+    Expressions        & 19,790,453 \\ % SELECT COUNT(ex) FROM ex;
+    Definitions        &  2,747,892 \\ % SELECT COUNT(df) FROM df;
     Language Varieties identified           &      9,310 \\  % SELECT COUNT(lv) FROM lv;
     Language Varieties with data           &      9,239 \\  % 
-    Meanings           & 20,773,371 \\  % SELECT COUNT(mn) FROM mn;
+    Languages          &      7,843 \\ % SELECT COUNT(lc) FROM lc; -- all language codes (ISO 639-3, ISO 639-2, ISO 639-5) known in the PanLex db
     Sources being consulted   &      4,190 \\  % 
     Sources already consulted &      1,453 \\  % 
-    Expressions        & 19,790,453 \\ % SELECT COUNT(ex) FROM ex;
-    Definitions        &  2,747,892 \\ % SELECT COUNT(df) FROM df;
     Users              &          23 \\ % SELECT COUNT(us) FROM us;
@@
 % \emph{Global Wordnet Association}.
-An important related initiative is the Global Wordnet Association\footnote{\url{http://www.globalwordnet.org/}\vspace{0.7em}},
+An important related initiative is the Global Wordnet Association (GWA)\footnote{\url{http://www.globalwordnet.org/}\vspace{0.7em}},
 which offers a platform for sharing wordnets and, among other goals, uniformly representing wordnets

### 2014-03-08 panlex-rdf-paper/2012/MLOD_PanLex/paper_v5.tex
Update to the paper

@@
 
-The Semantic Web Initiative has led to a stack of technologies standardized by the
-\emph{World Wide Web Consortium}\footnote{\url{http://www.w3.org/standards/semanticweb/}}
+The Semantic Web Initiative has led to the development of standards and technologies
 supporting a machine-readable and \mbox{-i}nter\-pretable Linked Data
-network. This has led to a \emph{Web of Data}.
+network, known as the \emph{Web of Data}.\footnote{\url{http://www.w3.org/standards/semanticweb/}} From these efforts the \emph{Linked Open Data (LOD) cloud}\footnote{\url{http://lod-cloud.net/}} emerged.
 %Technologies that support this aim are already available and in use in the so
 %called \emph{Web of Data}.
-A growing community is leveraging Semantic Web technologies for linguistic knowledge, building a Web of Linguistic
-Data, or a \emph{Linguistic Linked Open Data (LLOD) cloud}.
+A growing community is leveraging Semantic Web technologies for linguistic knowledge, building a \emph{Linguistic LOD (LLOD) cloud}.
 
-When linked open data are labeled in any language with lexemes in citation form (lemmas), the data are potentially linkable to, and through, PanLex.
+Whenever there is Linked Data making use of descriptions, such as labels, in any language with lexemes in citation form (lemmas), the data are potentially linkable to, and through, PanLex.
 Here we describe how we connected PanLex to the Linked Data network.
@@
 \label{sec:triplificate}
-Here we analyze the PanLex dataset; introduce our URI and vocabulary design, resembling PanLex's conceptual model;
+In this section, we analyze the PanLex dataset; introduce our URI and vocabulary design, which resemble PanLex's conceptual model;
 %, which was developed independently from other ontology engineering approaches.
@@
 \label{sec:analysis}
-The PanLex \emph{database} is created by editors who consult sources\footnote{\url{http://panlex.org/tech/plrefs.shtml}}, such as mono- and multilingual dictionaries, glossaries, standards, and thesauri.
-The data include single- and multi-word expressions, meanings assigned to them, and related information.
+The PanLex \emph{database} is created by editors who consult \emph{information sources}\footnote{\url{http://panlex.org/tech/plrefs.shtml}}, such as mono- and multilingual dictionaries, glossaries, standards, and thesauri.
+The data include single- and multi-word expressions, corresponding meanings assigned to them, and related information.
 While copies of the sources would not be interoperable, PanLex data constitute editors' \emph{interpretations} of sources' \emph{assertions} that two or more \emph{expressions} share a \emph{meaning}\footnote{\url{http://panlex.org/tech/doc/design/panlex-db-design.pdf}}.
@@
 the \emph{Lexicon Model for Ontologies} (lemon)~\cite{lemon2011} as well as the \emph{General Ontology for Linguistic Description} (GOLD)~\cite{farr2003}.
-Since these models differ from the PanLex one to some extent, we follow an
+Since these models differ from the PanLex one, we follow an
 incremental approach of aligning the PanLex data with them.
@@
 Resources for \emph{?lang} become typed as a \emph{Language} in the PanLex and the \emph{schema.org} namespace.
-This view-based approach also demonstrates that changing the vocabulary or
-adding support for new ones does not require an extract transform load (ETL) process,
-and can therefore be done with little effort.
+This view-based approach makes it easy to perform future revisions of RDF mapping, such as
+adding support for new vocabularies.
+%, as this does not require an extract transform load (ETL) process.
+%and can therefore be done with little effort.
+%This view-based approach makes it easy to revise the RDF mapping, such as adding support for new vocabular
 
@@
 \label{sec:linking}
-The SML view in the previous section (\autoref{fig:ex:sparqlify-ml}) already established the interlinking of the PanLex languages with Lexvo.
-In this section we outline the interlinking with DBpedia.
-For DBpedia, we were interested in creating \emph{valid} and thus \emph{dereferenceable} links.
+The SML view in (\autoref{fig:ex:sparqlify-ml}) already establishes the interlinking of the PanLex languages with Lexvo.
+In this section, we outline the interlinking with DBpedia, where we were interested in creating \emph{valid} and \emph{dereferenceable} links.
 Therefore, we iterated the \emph{titles} datasets\footnote{\url{http://wiki.dbpedia.org/Downloads38}}, which map (non-localized) DBpedia URIs to their page titles in the respective language.
@@
 APIs\footnote{\url{http://panlex.org/try/}} with Linked Data, powered by
-Pubby\footnote{\tiny{\url{http://wifo5-03.informatik.uni-mannheim.de/pubby/}}},
+Pubby\footnote{\url{http://wifo5-03.informatik.uni-mannheim.de/pubby/}},
 and two SPARQL
 endpoints\footnote{\url{http://ld.panlex.org/vsparql}}
-\footnote{\url{http://ld.panlex.org/sparql}}, ran by Sparqlify and Virtuoso,
-respectively.
+\footnote{\url{http://ld.panlex.org/sparql}}, ran by Sparqlify and Virtuoso.
 % and \url{http://ld.panlex.org/snorql}}.
@@
 views and the interlinking code are hosted on GitHub\footnote{\url{https://github.com/AKSW/PanLex-2-RDF}}.
-The created linksets are hosted in the PanLex database and are published together with the other data using the Sparqlify RDB2RDF tool.
+The created linksets are hosted in the PanLex database and are published together with the other data using Sparqlify.
 Finally, we offer downloads tagged with timestamps of their creation\footnote{\url{http://ld.panlex.org/downloads/releases/}}.
@@
 
-We intend to address some limitations in the future.
-Attribution is currently limited to sources, obscuring the possible roles of multiple editors in iteratively interpreting sources and the subsidiary sources of data in sources that are themselves aggregations.
-Metadata attached to PanLex denotations are currently limited to arbitrary pairs of strings, but this sacrifices discovery possibilities when the metadata (as they often do) describe facts that can be expressed with PanLex expressions (e.g., that the expression is archaic).
-New collaborations between PanLex and related standards and technologies (e.g., language identification, language geolocation, lemmatization, transliteration, localization, etc.) are promising arenas for development.
-For example, the relations among PanLex's information sources, if treated as distinct datasets, could be modeled with the VoID vocabulary\footnote{\url{http://rdfs.org/ns/void}}.
+We intend to address some limitations in the future:
+The relations among PanLex's information sources, if treated as distinct datasets, could be modeled with the VoID vocabulary\footnote{\url{http://rdfs.org/ns/void}}.
+The (assertion) source entity should be refactored to model users and information sources as distinct entities.
+%Attribution is currently limited to (assertion) sources,
+% obscuring the possible roles of multiple editors in iteratively interpreting sources and the subsidiary sources of data in sources that are %themselves aggregations.
+Metadata attached to PanLex denotations are currently limited to arbitrary pairs of strings,
+but this sacrifices discovery possibilities when the metadata
+%(as they often do)
+describe facts that can again be expressed with PanLex expressions.
+% (e.g., that the expression is archaic).
+Finally, new collaborations between PanLex and related fields (e.g.
+as language identification, language geolocation, lemmatization, transliteration, localization, etc.) are promising arenas for development.
 
@@
 \bibliography{bibliography,../../bib/aksw}
+
 \end{document}

### 2014-03-08 panlex-rdf-paper/2012/MLOD_PanLex/paper_v5.tex
Updated paper; still footnotes appear in references

@@
 
-The Semantic Web Initiative has led to the development of standards and technologies
+The Semantic Web initiative has led to the development of standards and technologies
 supporting a machine-readable and \mbox{-i}nter\-pretable Linked Data
@@
 
-Whenever there is Linked Data making use of descriptions, such as labels, in any language with lexemes in citation form (lemmas), the data are potentially linkable to, and through, PanLex.
-Here we describe how we connected PanLex to the Linked Data network.
+%Whenever there is Linked Data making use of descriptions, such as labels, in any language with lexemes in citation form (lemmas), the data are potentially linkable to, and through, PanLex.
+Here we describe how we connected PanLex to this Linked Data network.
 In \autoref{sec:triplificate} we introduce the dataset, present a
 PanLex RDF vocabulary, and explain how we transformed the one into the other and established conformance with additional data models.
-\autoref{sec:linking} is about how we linked to other datasets of the LLOD cloud,
+\autoref{sec:linking} shows how we linked to other datasets of the LLOD cloud,
 and \autoref{sec:publishing} is about the publication of the dataset.
-Usage scenarios are given in \autoref{sec:usage}.
-In \autoref{sec:related} we discuss related work, and, finally, in \autoref{sec:conclusion}, we summarize our recent and contemplated work.
+Usage scenarios are given in \autoref{sec:usage}, and related work is discussed in \autoref{sec:related}.
+Finally, \autoref{sec:conclusion} concludes paper.
+%In \autoref{sec:related} we discuss related work, and, finally, in \autoref{sec:conclusion}, we
+% summarize our recent and contemplated work.
+%conclude
 
@@
 The data include single- and multi-word expressions, corresponding meanings assigned to them, and related information.
-While copies of the sources would not be interoperable, PanLex data constitute editors' \emph{interpretations} of sources' \emph{assertions} that two or more \emph{expressions} share a \emph{meaning}\footnote{\url{http://panlex.org/tech/doc/design/panlex-db-design.pdf}}.
+%While copies of the sources would not be interoperable,
+PanLex data constitute editors' \emph{interpretations} of sources' \emph{assertions} that two or more \emph{expressions} share a \emph{meaning}\footnote{\url{http://panlex.org/tech/doc/design/panlex-db-design.pdf}}.
 The most important entities and relations of PanLex's conceptual model are depicted in \autoref{fig:db-schema}.
@@
 (2) Multiword expressions, such as ``fall in love'', traditionally found in an entry headed by one of their words, such as ``fall'' or ``love'', are treated as independent expressions in PanLex.
-Expression texts are constrained to comply with Unicode Normalization Form NFC.
+%Expression texts are constrained to comply with Unicode Normalization Form NFC.
+% ~Claus: If we state this, we also have to cite NFC, which requires again more space
     \item \emph{Languages} in PanLex are identified using ISO 639-3\footnote{\url{http://www.sil.org/iso639-3/codes.asp}} individual and macrolanguage codes, ISO 639-2\footnote{\url{http://www.loc.gov/standards/iso639-2/}} collective codes, and ISO 639-5\footnote{\url{http://www.loc.gov/standards/iso639-5/}} codes.
@@
 %The only properties of this entity are ambiguity and synonymy: Controlled language varieties are usually defined to prohibit both of these.
-    \item \emph{Meanings} are the entities assigned to expressions, thereby identifying expressions as translations or synonyms.
+    \item \emph{Meanings} are entities assigned to expressions, thereby identifying expressions as translations or synonyms.
 For example, a source's translation of the German expression ``klingen'' into English as ``ring, sound, seem'' can be interpreted as (1) the assignment of a single meaning to all four expressions or (2) the assignment of two or three meanings to ``klingen'' and of one of those to each of the English expressions.
@@
   \item \emph{Users} may define sources, attribute data to them, and define language varieties.
-  \item Among the properties of sources are \emph{licenses}.
-Some of the license categories are \emph{public domain}, \emph{Creative Commons (CC)}, \emph{request} (author invites inquiries), \emph{GNU General Public License (GPL)}, \emph{GNU Lesser General Public License (LGPL)}, \emph{GNU Free Documentation License (FDL)}, \emph{MIT License}, \emph{copyright} (without a more liberal license), and \emph{PanLex Use Permission} (specific permission for use in PanLex).
-    The distribution of licenses as of January 2014 is shown in \autoref{tbl:plx-license-counts}.
+  \item Among the properties of sources are \emph{licenses}. Some of the license categories are \emph{public domain},
+%\emph{Creative Commons (CC)},
+\emph{request} (author invites inquiries),
+%\emph{GNU General Public License (GPL)},
+%\emph{GNU Lesser General Public License (LGPL)},
+\emph{GNU Free Documentation License (FDL)},
+%\emph{MIT License}, \emph{copyright} (without a more liberal license),
+and \emph{PanLex Use Permission} (specific permission for use in PanLex).
+    The distribution of licenses is shown in \autoref{tbl:plx-license-counts}.
+%as of January 2014
 \end{itemize}
@@
 
-\subsection{The PanLex vocabulary}
+\subsection{The PanLex Vocabulary}
 \label{sec:vocabulary}
 The entities and relations described above are the base for the PanLex RDF vocabulary.
-In general, all PanLex RDF resources reside in the namespace \texttt{\small <http://ld.panlex.org/plx/>}, abbreviated with \texttt{\small plx}.
+In general, all PanLex RDF resources reside in the namespace \texttt{\small{http://ld.panlex.org/plx/}}, abbreviated with \texttt{\small{plx}}.
 An example of the resulting ontology is depicted in \autoref{fig:vocabulary} and summarized as follows.
@@
     Meanings may also have \emph{definitions}, entities of the \texttt{\small plx:Definition} class, giving a textual representation (\texttt{\small rdfs:label}) in a certain language variety (\texttt{\small plx:languageVariety}).
-  \item Following the semantics of the PanLex database, meanings and expressions are linked via \emph{denotations}.
+  \item Meanings and expressions are linked via \emph{denotations}.
+%Following the semantics of the PanLex database,
     These are entities of the \texttt{\small plx:Denotation} class pointing to meanings and expressions via the properties \texttt{\small plx:denotationMeaning} and \\ \texttt{\small plx:denotationExpression}.
@@
 \includegraphics[width=0.46\textwidth]{images/pdf/ontology.pdf}
-\caption{Overview of the PanLex RDF vocabulary}
+\caption{Overview of the PanLex RDF Vocabulary}
 \label{fig:vocabulary}
@@
 
-\subsection{Transformation workflow}
+\subsection{RDF Transformation Workflow}
 \label{sec:conversion}
@@
 The relations among PanLex's information sources, if treated as distinct datasets, could be modeled with the VoID vocabulary\footnote{\url{http://rdfs.org/ns/void}}.
-The (assertion) source entity should be refactored to model users and information sources as distinct entities.
+The source entity should be refactored to reference users and information sources as distinct entities.
 %Attribution is currently limited to (assertion) sources,

### 2014-03-08 panlex-rdf-paper/2012/MLOD_PanLex/paper_v5.tex
Except for the reference section which is too close to the footnotes, I guess my part is done.

@@
 \label{sec:linking}
-The SML view in (\autoref{fig:ex:sparqlify-ml}) already establishes the interlinking of the PanLex languages with Lexvo.
-In this section, we outline the interlinking with DBpedia, where we were interested in creating \emph{valid} and \emph{dereferenceable} links.
+The SML view in (\autoref{fig:ex:sparqlify-ml}) establishes the interlinking of languages in PanLex with Lexvo.
+Here we outline the interlinking with DBpedia, where we were interested in creating \emph{valid} and \emph{dereferenceable} links.
 Therefore, we iterated the \emph{titles} datasets\footnote{\url{http://wiki.dbpedia.org/Downloads38}}, which map (non-localized) DBpedia URIs to their page titles in the respective language.
@@
 and two SPARQL
-endpoints\footnote{\url{http://ld.panlex.org/vsparql}}
-\footnote{\url{http://ld.panlex.org/sparql}}, ran by Sparqlify and Virtuoso.
+endpoints\footnote{\url{http://ld.panlex.org/vsparql}}\textsuperscript{,}\footnote{\url{http://ld.panlex.org/sparql}}, ran by Sparqlify and Virtuoso.
 % and \url{http://ld.panlex.org/snorql}}.
@@
 % \emph{Global Wordnet Association}.
-An important related initiative is the Global Wordnet Association (GWA)\footnote{\url{http://www.globalwordnet.org/}\vspace{0.7em}},
+An important related initiative is the Global Wordnet Association (GWA)\footnote{\url{http://www.globalwordnet.org/}},
 which offers a platform for sharing wordnets and, among other goals, uniformly representing wordnets
@@
 % representing the same unique sense.
-In the Semantic Web context, several standard or quasi-standard vocabularies and ontologies have been developed with the rise of the \emph{Linguistic Linked Open Data} movement.
+In the Semantic Web context, several standard or quasi-standard vocabularies and ontologies have been developed with the rise of the Linguistic LOD movement.
 Examples include the
@@
 
+\vfill
+%\columnbreak
+
 We intend to address some limitations in the future:
@@
 Finally, new collaborations between PanLex and related fields (e.g.
-as language identification, language geolocation, lemmatization, transliteration, localization, etc.) are promising arenas for development.
+as language identification, language geolocation, lemmatization, transliteration, localization, etc.) are promising areas for development.
+
+\vspace{-10pt}
 

### 2014-03-09 panlex-rdf-paper/2012/MLOD_PanLex/paper_v5.tex
updated figure 1
- redrew png image as svg
- added bigger arrow labels
updated figure 2
- added real world data

@@
   \centering
-  \includegraphics[width=0.42\textwidth]{images/schema_new.png}
+  \includegraphics[width=0.42\textwidth]{images/pdf/db_schema.pdf}
   \caption{The PanLex database schema}
@@
 \begin{figure}
-\includegraphics[width=0.46\textwidth]{images/pdf/ontology.pdf}
-\caption{Overview of the PanLex RDF Vocabulary}
+\includegraphics[width=0.46\textwidth]{images/pdf/ontology_new.pdf}
+% Query for ontology_new.pdf
+%SELECT * WHERE {
+%  ?den plx:denotationMeaning ?mea .
+%  ?den plx:denotationExpression ?expr .
+%  ?expr rdfs:label ?exprlbl .
+%  ?mea plx:meaningDefinition ?def .
+%  ?mea plx:approver ?app.
+%  ?def plx:languageVariety <http://ld.panlex.org/plx/languageVariety/187> .
+%  ?def rdfs:label ?dtx .
+%  ?den plx:denotationExpression <http://ld.panlex.org/plx/expression/355621> .
+%  ?app dc:title ?dctitle .
+%  ?app plx:quality ?plxquality .
+%  ?app dc:publisher ?dcpublisher .
+%  ?app dc:date ?dcdate .
+%  ?app rdfs:label ?rdfslabel .
+%  ?app foaf:homepage ?foafhomepage .
+%  ?app dc:creator ?dccreator .
+%  ?app plx:registrationDate ?regdate .
+%  ?app plx:license <http://ld.panlex.org/plx/license/mi> .
+%  <http://ld.panlex.org/plx/license/mi> rdfs:label ?liclbl .
+%  OPTIONAL { ?app dbpedia-owl:isbn ?dboisbn . }
+%  OPTIONAL { ?den plx:wordClass ?wordclass . }
+%}
+\caption{Example of the PanLex RDF vocabulary showing one meaning of the expression `between' and the corresponding source and definition}
 \label{fig:vocabulary}

### 2014-03-09 panlex-rdf-paper/2012/MLOD_PanLex/paper_v5.tex
proof read +  alternative fig 2

@@
 \label{sec:triplificate}
-In this section, we analyze the PanLex dataset; introduce our URI and vocabulary design, which resemble PanLex's conceptual model;
+In this section, we analyze the PanLex dataset, introduce our URI and vocabulary design, which resemble PanLex's conceptual model,
 %, which was developed independently from other ontology engineering approaches.
-summarize how we classified PanLex's instance data with additional data models; and explain our transformation of the data to RDF.
+summarize how we classified PanLex's instance data with additional data models and explain our transformation of the data to RDF.
 

### 2014-03-10 panlex-rdf-paper/2012/MLOD_PanLex/paper_v5.tex
updated fig 2
- switched to figure version without fake data
- used bigger foaf:homepage URL with ellipsis

@@
 \begin{figure}
-\includegraphics[width=0.46\textwidth]{images/pdf/ontology_new.pdf}
+\includegraphics[width=0.46\textwidth]{images/pdf/ontology_new3.pdf}
 % Query for ontology_new.pdf
@@
 \label{sec:linking}
-The SML view in (\autoref{fig:ex:sparqlify-ml}) establishes the interlinking of languages in PanLex with Lexvo.
-Here we outline the interlinking with DBpedia, where we were interested in creating \emph{valid} and \emph{dereferenceable} links.
+The SML view in (\autoref{fig:ex:sparqlify-ml}) establishes the interlinking of languages in PanLex with Lexvo~\cite{deMeloWeikum2008c}.
+Here we outline the interlinking with DBpedia~\cite{dbpedia-swj}, where we were interested in creating \emph{valid} and \emph{dereferenceable} links.
 Therefore, we iterated the \emph{titles} datasets\footnote{\url{http://wiki.dbpedia.org/Downloads38}}, which map (non-localized) DBpedia URIs to their page titles in the respective language.
@@
 language information retrieval, as well as annotation projects, like DBpedia
-Spotlight\footnote{\url{http://spotlight.dbpedia.org}}, seems worthwhile.
+%Spotlight\footnote{\url{http://spotlight.dbpedia.org}},
+Spotlight~\cite{isem2011mendesetal}
+ seems worthwhile.
 
@@
 The extraction of information from linguistic sources, and techniques for automatically inferring translations,
-are relevant work discussed in \cite{panlex_probtrans}.
+are relevant work discussed in~\cite{panlex_probtrans}.
 %A similar approach of linking expressions to meanings is followed by the
@@
 \emph{Ontologies of Linguistic Annotation} (OLiA)~\cite{olia2010} for modeling
-lexicon and machine-readable dictionaries, \emph{POWLA} for modeling linguistic corpora\cite{powla2012} and the \emph{Natural Language Processing Interchange Format} (NIF)\footnote{\url{http://nlp2rdf.org/nif-1-0}}.
+lexicon and machine-readable dictionaries, \emph{POWLA} for modeling linguistic corpora~\cite{powla2012} and the \emph{Natural Language Processing Interchange Format} (NIF)\footnote{\url{http://nlp2rdf.org/nif-1-0}}.
 
